### 2020-07-07 manuscript.Rmd
code, data, and manuscript files to reproduce all analyses and figures

---
title: Repeated loss of variation in insect ovary morphology highlights the role of developmental constraint in life-history evolution
output:
  bookdown::pdf_document2:
  keep_tex: yes
  latex_engine: pdflatex
  #bookdown::html_document2: default
toc: false
bibliography: manuscript_supplement.bib
csl: nature.csl
link-citations: yes
header-includes:
 \usepackage{xr} \externaldocument{manuscript_supplement}
 \usepackage{float}
 \floatplacement{figure}{H}
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(echo=FALSE, include = FALSE, warning=FALSE, message=FALSE, cache=FALSE)

load("manuscript_supplement.RData") 
library(dplyr)
```

Samuel H Church^1^\*, Bruno A. S. de Medeiros^1,2^, Seth Donoughe^1,3^, Nicole L. Márquez Reyes^4^, Cassandra G. Extavour^1,5^\*

^1^ Department of Organismic and Evolutionary Biology, Harvard University, Cambridge, MA 02138, USA

^2^ Smithsonian Tropical Research Institute, Panama City, Panama

^3^ Department of Molecular Genetics and Cell Biology, University of Chicago, Chicago, IL 60637, USA

^4^ Department of Biology, Universidad de Puerto Rico en Cayey, Cayey 00736, PR

^5^ Department of Molecular & Cellular Biology, Harvard University, Cambridge, MA 02138, USA

\* Corresponding author

# Abstract {-}

The number of offspring an organism can produce is a key component of its evolutionary fitness and life-history. Here we perform a test of the hypothesized trade off between the number and size of offspring using thousands of descriptions of the number of egg-producing compartments in the insect ovary (ovarioles), a common proxy for potential offspring number in insects. In contrast to prior claims, we find that ovariole number is not generally negatively correlated with the size of insect eggs, and we highlight several factors that may have contributed to this size-number trade off being strongly asserted in previous studies. We reconstruct the evolutionary history of the nurse cell arrangement within the ovariole, and show that the diversification of ovariole number and egg size have both been largely independent of nurse cell presence or position within the ovariole. Instead we show that ovariole number evolution has been shaped by a series of transitions between variable and invariant states, with multiple independent lineages evolving to have almost no variation in ovariole number. We highlight the implications of these invariant lineages on our understanding of the specification of ovariole number during development, as well as the importance of considering developmental processes in theories of life-history evolution.

# Introduction {-}

Offspring number is a fundamental parameter in the study of life-history[@stearns1992evolution]. This number differs widely between organisms[@stearns1992evolution], and its variation is the foundation for several hypotheses about life-history evolution, including the prediction that there is an evolutionary trade off between the number of offspring and their size (e.g. egg size)[@smith1974optimal;@lack1947significance;@stearns1992evolution]. In insects, the number of egg-producing compartments in the ovary, called ovarioles, has been used as a proxy for potential offspring number in the study of life-history[@honvek1993intraspecific;@berrigan1991allometry;@hodin2009she]. However, without an understanding of the phylogenetic distribution of ovariole number, this hypothesized relationship cannot be assessed across insects. Here we tested for the presence of a general trade off between ovariole number and egg size by collecting thousands of records of ovariole number from the published literature, placing them in a phylogenetic context, and comparing them to other datasets of insect reproductive morphology.

The insect female reproductive system includes a pair of ovaries, each of which contains a number of ovarioles[@buning1994insect]  (Fig 1a). Each ovariole consists of an anterior germarium containing the stem cell niche or resting oogonia, developing oocytes arranged in an ontogenic series from anterior to posterior, and a posterior connection to a common oviduct. The number of ovarioles varies across species[@hodin2009she], and can vary across individuals in a population[@honvek1993intraspecific], as well as between the left and right ovary within a single individual[@iwata1955comparative]. Therefore total ovariole number may be an even or odd integer for an individual female insect. In addition to variation in the number of ovarioles, the tissue morphology within ovarioles varies across insects, and has been classified into several modes of oogenesis based on the presence and position of special nutritive cells called nurse cells[@buning1994insect].

Here we compiled `r nrow(raw_ON)` records of ovariole number from across `r ON %>% distinct(order) %>% nrow()` orders, `r ON %>% distinct(family) %>% nrow()` families, and `r ON %>% filter(species != "") %>% distinct(genus,species) %>% nrow()` species of insects. We combined these data with published datasets of egg size[@church2019dataset], fecundity[@Gilbert2007], body size[@rainford2016phylogenetic], and mode of oogenesis[@buning1994insect] to test hypotheses about reproductive evolution. In these analyses we used an existing phylogeny of insects[@church2019insect] to analyze evolutionary patterns in ovariole number, and found that hypotheses about reproductive trait evolution do not hold generally true across insects. Instead we observe that the phylogenetic distribution of ovariole number suggests a model where the developmental mechanisms that govern ovariole number have shifted between variable and invariant states several times over the course of insect evolution. We propose that the developmental mechanisms used to establish ovariole number in well-studied insects such as _Drosophila melanogaster_ are unlikely to regulate ovariole number in all insects.

# Methods {-}

## Gathering trait data {-}

We searched the published literature for references to insect ovariole number using a predetermined set of 131 search terms, entered into Google Scholar (scholar.google.com) between June and October of 2019. Each search term consisted of an insect taxonomic group and the words "ovariole number". This list was created to include all insect orders, many large insect families, and groups well-represented in the insect egg dataset[@church2019dataset]. For each search term, we evaluated all publications in the first page of results (ten publications). For 61 search terms that had a large number of informative hits, significant representation in the egg dataset, or that corresponded to very speciose groups, we evaluated an additional 20 publications. The list of search terms is available in the supplementary file 'ovariole_number_search_terms.tsv'.

Using this approach, we gathered `r nrow(raw_ON)` records for ovariole number from `r raw_ON %>% distinct(publication) %>% nrow()` publications, listed in the supplementary files 'ovariole_number_bibliography.pdf' and 'ovariole_number_bibliography.bib'. We matched these records to additional taxonomic information using the software TaxReformer[@TaxReformer]. For all subsequent analyses, we excluded observations made in non-reproductive individuals from eusocial species (workers), as well as two observations that represented significant outliers and could not be validated using additional sources[@su2015testicular;@hernandez2017biology]. See Supplemental Methods Section \ref{sec:gathering} for details.

We combined the data we collected on total ovariole number with existing datasets of egg size and shape[@church2019dataset], insect lifetime fecundity and dry adult body weight[@Gilbert2007], average adult body length per insect family[@rainford2016phylogenetic], several lineage-specific measures of adult body size[@iwata1966large;@Iwata1966a;@waloff1954number;@starmer2003phylogenetic;@reinhardt2005low], and the mode of oogensis[@buning1994insect]. See Supplemental Methods Section \ref{sec:combining} for details.

All continuous traits (ovariole number, egg volume, lifetime fecundity, and all measures of body size) were log~10~ transformed for subsequent analyses.

## Phylogenetic analyses {-}

The analyses in this manuscript were performed using the insect phylogeny published in Church et al., 2019[@church2019insect], unless otherwise specified.  Analyses of insect family-level ovariole number, egg size, and body size were performed using the insect phylogeny published in Rainford et. al, 2014[@rainford2014phylogenetic]. Analyses of Drosophilidae ovariole number, egg size, and body size were performed using a phylogeny newly assembled for this study. See Supplemental Methods Section \ref{sec:trees} for details.

To evaluate the robustness of our results to uncertainty in the phylogenetic relationships, all Phylogenetic Generalized Least Squares (PGLS) analyses were performed 1000 times over a posterior distribution of trees, using a Brownian Motion based covariance matrix in the R package ape (version `r packageVersion('ape')`)[@paradis2004ape] and nlme (version `r packageVersion('nlme')`)[@pinheiro2014r]. For regressions at the species and genus level, we reshuffled and matched records for each iteration to account for variation across records for the same taxon. For regressions at the family level we recalculated the average ovariole number per insect family, randomly downsampling the representation for each family by half. To weight traits by body size, we calculated the phylogenetic residuals[@revell2010phylogenetic] of each trait to body size, and then compared the evolution of these residuals using a PGLS regression. See Supplemental Methods Section \ref{sec:PGLS} for details.

Ancestral state reconstruction of oogenesis mode was performed with the R package corHMM[@beaulieu2013identifying], version `r packageVersion('corHMM')`, and models of trait evolution were compared using the R package Ouwie[@beaulieu2012modeling], version `r packageVersion('OUwie')`. Ancestral state reconstruction and model comparison were repeated 100 times over a posterior distribution of trees and resampling data to account for variation across records for the same taxon. See Supplemental Methods Section \ref{sec:oogenesis-model}.

Other comparisons of model fit were performed using the R package geiger[@harmon2007geiger], version `r packageVersion('geiger')`, and validated using a parametric bootstrap with the R package arbutus[@pennell2015model], version `r packageVersion('arbutus')`. See Supplemental Methods Section \ref{sec:bootstrap}.

Analyses of evolutionary rate were performed using BAMM[@rabosky2014automatic], version 2.5.0. For this analysis, we calculated the average ovariole number (log~10~ transformed) for each genus present in the phylogeny (507 taxa). We used the R package BAMMtools (version `r packageVersion('BAMMtools')`)[@rabosky2014bamm] to select appropriate priors, and ran BAMM for the maximum number of generations ($2*10^{-9}$), sampling every 10^6^ generations. Convergence was evaluated both visually (Fig. \ref{fig:BAMM-analysis-convergence}) and numerically. Running BAMM for the maximum possible number of generations and selecting the optimum burn-in\ref{fig:BAMM-analysis-burnin} resulted in an effective size for the number of shifts of `r round(es_nshifts,2)`, and for log-likelihood of `r round(es_logl,2)`. Repeated BAMM analyses showed similar distributions of high and low rate regimes, indicating the implications for ovariole number evolution are robust to uncertainty in rate estimates. See Supplemental Methods Section \ref{sec:BAMM} for details.

We visualized the results from the BAMM analysis to establish a threshold (`r lowrate`) for assigning a binary rate regime to each node in the phylogeny, categorizing them as above (variable) or below (invariant) a threshold that separates these two peaks.

## Data availability {-}

The dataset of insect ovariole number is available in the supplementary file 'ovariole_number_data.tsv' and has been uploaded to Dryad (doi:10.5061/dryad.59zw3r253). The code and phylogenetic trees required to reproduce all the analyses, figures, and generate the manuscript files are provided at 'https://github.com/shchurch/insect_ovariole_number_evolution_2020'.

## Statistical significance {-}

All regressions were performed over a posterior distribution of trees and accounting for phenotypic uncertainty; therefore these analyses produced a range of p-values. We considered a relationship significant when the maximum p-value observed from this range was below the threshold 0.01. We report the maximum p-value throughout this manuscript, and have included the full range in the supplementary information (see Table \ref{tab:PGLS-posterior}).

Model comparisons of trait evolution were also performed over a posterior distribution and accounting for phenotypic uncertainty. For these analyses, we considered a model to have significantly better fit the data than other models when the difference in the corrected Akaike Information Criterion (AICc) was greater than two in every analysis iteration.

# Results {-}

## Ovariole number diversity {-}

Ovariole number varies by at least four orders of magnitude across insects (Fig. 1b). We identify seven insect families with species that have been reported to have more than 1,000 total ovarioles, including several eusocial insects (e.g. queens of the termite species _Hypotermes obscuriceps_, Blattodea: Termitidae[@bugnion1912anatomie], and several ant species, Hymenoptera: Formicidae)[@robertson1995sperm;@schneirla1957comparison] and non-eusocial insects (e.g. the blister beetle _Meloe proscarabaeus_, Coleoptera: Meloidae)[@buning1979trophic]. We also find two independent lineages that have evolved to have only one functional ovariole: dung beetles in the tribe Scarabaeinae (Coleoptera: Scarabaeidae)[@richter1974ovariole], and grass flies in the genus _Pachylophus_ (Diptera: Chloropidae)[@meier1999ovoviviparity;@pollock1996viviparous]. In these insects one of the two ovaries presumably established during embryogenesis is reported to atrophy during development[@pollock1996viviparous;@pluot1979evolution], resulting in an asymmetric adult reproductive system.

![**The diversity of ovariole number across insects.** a, Schematic of a generalized insect female reproductive system, showing a pair of ovaries, each with four ovarioles. b, The range of total adult ovariole number, log~10~ scale, across nine groups of insects, arranged with random jitter on the y-axis within each group. Groups are, from top to bottom: Apterygota, Palaeoptera, Polyneoptera, Condylognatha, Psocodea, Amphiesmenoptera, Antliophora, Neuropteroidea, and Hymenoptera.](manuscript_figures/Fig_Diversity_V3-01.png)

## Ovariole number, egg size, and body size {-}

Ovariole number has been hypothesized to be negatively correlated with egg size[@montague1981reproductive;@berrigan1991allometry;@starmer2003phylogenetic]. This hypothesis is based on the predictions that (1) female reproduction is resource-limited, therefore egg size should trade off with egg number, and (2) ovariole number can serve as a proxy for egg number[@montague1981reproductive;@smith1974optimal]. Our data do not support the hypothesis of a negative relationship between egg size and ovariole number across insects (p-value `r all_table[2,]$'pval_max'`; Figs. 2a, \ref{fig:PGLS-egg-ON-genus}, Table \ref{tab:PGLS-posterior}).

Given that this prediction is often conditioned on taking into account adult body size[@stewart1991reproductive;@starmer2003phylogenetic], we combined data on ovariole number and egg size with several measurements of body size, and applied a size correction that accounts for phylogenetic relationships [@revell2009size]. Across insects we did not observe a significant relationship between ovariole number and egg size when controlling for body size (p-value `r all_table[6,]$'pval_max'`; Fig. 2b, \ref{fig:PGLS-Gilbert-name}, \ref{fig:PGLS-egg-ON-family}, Table \ref{tab:PGLS-posterior})), though we do note the presence of a non-significant negative trend, primarily driven by a few extreme values (e.g. wasps in the family Trigonalidae, which have many ovarioles and lay minute eggs, Fig. \ref{fig:PGLS-egg-ON-family}).

We assessed the relationship between these traits for four sublineages of insects and found that across Drosophilidae species, egg size is indeed strongly negatively correlated with ovariole number when accounting for body size (p-value <0.001; Fig. 2c, Table \ref{tab:PGLS-posterior}). However this relationship did not hold for any of the three other lineages we analyzed (e.g. Orthoptera p-value `r all_table[9,]$'pval_max'`; Fig. 2d, \ref{fig:PGLS-Coleoptera-Hymenoptera}, Table \ref{tab:PGLS-posterior}).

We additionally tested whether ovariole number is positively correlated with adult body size, and in contrast to previous studies[@honvek1993intraspecific], we found no correlation between ovariole number and adult body weight or length across insects (p-values `r body_all_table[3,]$'pval_max'` and `r body_all_table[1,]$'pval_max'`, respectively; Fig. \ref{fig:PGLS-body-plot}, Table \ref{tab:PGLS-body-posterior}).

![**Tests of the hypothesized trade off between egg size and ovariole number.** a, Egg volume (mm^3^) and ovariole number, both log~10~ scale, n=`r nrow(egg_ON_name_pgls)` species. See section _Modeling ovariole number evolution_ for discussion of the enrichment of certain low values of ovariole number (points appearing vertically arranged) b, Egg volume and ovariole number, residuals to adult dry body weight, n=`r length(which(egg_ON_gilbert_genus_pgls$rank %in% tree$tip.label))` genera. c, Drosophilidae egg volume and ovariole number, residuals to thorax length, n=`r length(which(DIP_pgls$rank %in% dros_tree$tip.label))` species. d, Orthoptera egg volume and ovariole number, residuals to body length, n=`r length(which(ORT_pgls$rank %in% tree$tip.label))` genera.](manuscript_figures/Fig_Egg_V2-01.png)

## Ovariole number and fecundity {-}

If the hypothesized trade off between the number and size of offspring is true for insects, then one explanation for the lack of a consistent negative relationship between ovariole number and egg size is that ovariole number may not be a reasonable proxy for offspring number. Previous research has shown that, across individuals within the same species, ovariole number is correlated with certain measurements of fecundity, such as maximum daily rate of egg production for _Drosophila_ [@david1970nombre;@bouletreau1982reproductive], but not others, such as lifetime fecundity[@schmidt2005geographic] or fitness in competition assays[@wayne1997quantitative]. Few studies have compared fecundity and ovariole number across species[@stewart1991reproductive], likely due to the difficulties of measuring fecundity consistently across insects, many of which lay eggs singly and continuously rather than in distinct clutches.

Using a previously reported dataset of lifetime fecundity measurements across insects[@Gilbert2007], we assessed the relationship between lifetime fecundity and ovariole number, and observed a positive, but not statistically significant, correlation (p-value `r all_table[8,]$'pval_max'`; Figs. 3, \ref{fig:PGLS-fecundity-name}). We interpret our results, in conjunction with those previously reported, to suggest that ovariole number, when variable across insects in a lineage, may be one factor among many influencing the number of eggs produced. However, we caution against using ovariole number as a direct mathematical proxy for offspring number.

![**The relationship between lifetime fecundity and ovariole number.** Both values are shown on a log~10~ scale. n=`r length(na.omit(fecundity_genus_pgls$fecundity))` genera, colored according to the groups shown in Fig. 1b.](manuscript_figures/Fig_Fecundity_V1-01.png)

## Evolution of nurse cells {-}

In addition to the number of ovarioles, insect ovary morphology has been classified into several modes of oogenesis based on the presence and position of nutritive cells called nurse cells[@buning1994insect], (Fig. 4a). The ovary of the well-studied species _D. melanogaster_ is an example of a meroistic oogenesis mode, meaning that its ovarioles contain nurse cells of germ line origin that are connected to developing oocytes via cytoplasmic bridges[@KingRobertC1970OdiD]. In insects with a polytrophic arrangement, these nurse cells are clonally related and immediately adjacent to each oocyte. An alternative arrangement is seen in telotrophic meroistic ovaries, where oocytes in each ovariole are connected to a common pool of nurse cells located in the germarium[@buning1994insect]. Meroistic ovaries are thought to have evolved from a panoistic ancestral state, meaning they lack nurse cells[@buning1994insect]. Using a previously published set of descriptions of these oogenesis modes across insects[@buning1994insect], we reconstructed the evolutionary transitions between these states. Consistent with previous analyses[@buning1994insect], we found that the ancestral insect likely had panoistic ovaries (without nurse cells), with several independent shifts to both telotrophic and polytrophic meroistic modes, and at least two reversals from meroistic back to panoistic (Figs. 4b, \ref{fig:reconstruct-ovary}).

Using this ancestral state reconstruction, we then compared models of trait evolution to test whether evolutionary transitions in oogenesis mode helped explain the diversification of ovariole number and egg morphology. We found that, for the traits studied here, models that take into account evolutionary changes in mode of oogenesis do not consistently demonstrate a significant improvement over models that do not take these changes into account ($\Delta$AIC < 2, Table \ref{tab:ovary-model-comparison}). In other words, the evolution of nurse cells and their position within the ovary do not explain the diversification of egg size, egg shape, or ovariole number.

To analyze the robustness of these results to uncertainty in the tree topology and in the inference of ancestral states, we repeated each analysis over a posterior distribution of trees. For egg asymmetry and curvature, but not for volume or aspect ratio, we observed a few iterations where a model that takes into account oogenesis mode evolution was significantly favored over models that did not ($\Delta$AIC > 2, Table \ref{tab:ovary-model-comparison}). However, this result was infrequent over 100 repetitions of the analysis. We therefore interpret these results as suggestive of a possible relationship between mode of oogenesis and egg asymmetry and curvature, but one which cannot be confirmed given the current data available.

![**The evolution of the presence and position of nurse cells.** a, Insect oogenesis was categorized into several modes by Büning[@buning1994insect] based on the presence and position of nurse cells. b, Phylogenetic reconstruction of mode of oogenesis. Scale bar indicates 100 million years (Myr). Gray = panoistic ovaries, without nurse cells, cyan=polytrophic meroistic ovaries, with nurse cells adjacent to maturing oocytes, red = telotrophic meroistic ovaries, with nurse cells located in germaria, black = unique meroistic ovary type observed in Strepsiptera. Insect taxonomic groups are, from top to bottom: Apterygota, Palaeoptera, Polyneoptera, Condylognatha, Psocodea, Hymenoptera, Neuropteroidea, Amphiesmenoptera, and Antliophora. ](manuscript_figures/Fig_Ovary_V8-01.png)

## Modeling ovariole number evolution { - }

Using the dataset compiled here and a previously published phylogeny of insects (Fig. 5a)[@church2019insect], we modeled the rate of evolutionary change in ovariole number (Figs. \ref{fig:arbutus}, \ref{fig:BAMM-analysis-convergence}, \ref{fig:BAMM-analysis-burnin}, \ref{fig:BAMM-analysis-results}). We observed substantial rate heterogeneity in the evolution of ovariole number (Fig. \ref{fig:BAMM-analysis-results}), meaning that for some lineages ovariole number has evolved rapidly where in others, ovariole number has evolved very slowly or not at all. The most striking example of this are the multiple lineages which have independently evolved invariant or near-invariant ovariole number across taxa (e.g. nearly all Lepidoptera have exactly eight ovarioles, Fig. 5b, cyan), from an ancestral variable state. These invariant lineages were objectively identified by finding regions of the phylogeny that experience extremely low rates of ovariole number diversification (Figs. \ref{fig:BAMM-analysis-results}, \ref{fig:BAMM-anlaysis-mean-rates}). Using this approach, we found that invariant ovariole numbers have evolved at least nine times independently across insects, with several subsequent reversals from invariant to variable states (Fig. 5a).

We find that the rate of evolutionary change in ovariole number is correlated with the number of ovarioles: lineages with relatively low ovariole number also experience relatively low degrees of change (Fig. \ref{fig:arbutus}). This is evidenced by the fact that, of the nine invariant lineages, none have greater than seven ovarioles per ovary (Fig. 5c). However we note that not all insects with low ovariole counts are in invariant lineages; many insects with fewer than 14 total ovarioles are in lineages with relatively high rates of intra- and interspecific ovariole number variation (Fig. 5)

Additionally, the distribution of ovariole numbers across insects is enriched for even numbers of total ovarioles (Fig. 5c). While many insects show asymmetries in the number of ovarioles between the left and right ovaries, all of the invariant lineages are symmetric (at 4, 6, 8, 10, 12, and 14 total ovarioles). Therefore, invariant lineages have near-zero variation when comparing between species, between individuals within a species, and between the left and right ovary within an individual.

Using these results, we propose a multi-rate model, where the rate of ovariole number evolution differs based on the evolution of a discrete trait representing invariant or variable status. We propose that the evolution of this discrete trait is governed by a model where the likelihood of transitions from a variable to an invariant state is negatively correlated with the current number of ovarioles. Here we demonstrate that a multi-rate Brownian motion model far outperforms a single rate model in fitting the data ($\Delta$ AICc `r round(BM1$AICc - BMS$AICc,2)`).

![**The evolutionary distribution of ovariole number across insects.** a, Phylogeny of insect genera, colored according to the inferred rate regime of ovariole number evolution, variable in gray and invariant in black (see Supplementary Methods). b, Total ovariole numbers, shown on a log~10~ scale and arranged by insect genus according to the phylogeny. Tips with more than one point represent genera with multiple records for total ovariole number in the dataset. c, The distribution of values shown in (b), showing enrichment for even values in the left tail of the distribution.](manuscript_figures/Fig_Phylogeny_V3-01.png)

# Discussion {-}

A frequently invoked life-history prediction is that, given a finite set of metabolic resources, organisms can either produce few offspring, each with high fitness, or many low-fitness offspring[@smith1974optimal;@lack1947significance;@stearns1992evolution]. In insects, egg size and ovariole number are often used as proxies for offspring fitness[@koch2014mother] and number[@david1970nombre;@bouletreau1982reproductive], respectively, and therefore it has been predicted that insects with more ovarioles lay smaller eggs than insects with fewer ovarioles[@hodin2009she;@montague1981reproductive;@berrigan1991allometry;@starmer2003phylogenetic]. Our results, using a dataset that spans `r nrow(raw_ON)` observations and that takes into account phylogenetic relationships, indicate that a general trade-off between insect egg size and ovariole number does not exist (Fig. 2).

Lineages of insects with invariant ovariole number illustrate this point. Despite having the same ovariole number, these lineages contain a range of egg sizes that is comparable to the four orders of magnitude observed across all insects (Fig. 2a). Furthermore, we observed no relationship between the evolutionary rates of change for ovariole number and egg size (Fig. \ref{fig:combine-rates}). Therefore, if a trade-off between egg size and fecundity exists, factors beyond variation in ovariole number must contribute to fecundity. These factors might include variation in the rate of egg production per ovariole[@drummond2001stem;@ables2012control;@mirth2019turning;@wcislo1987roles], among others[@partridge1986examination;@parker1984models].

We suggest that considering the evolution of developmental processes that govern ovariole number specification may be more useful in explaining patterns of diversity than predictions based on metabolic trade-offs. The regulation of ovariole number is best understood from research on _Drosophila melanogaster_, where the number of ovarioles in the adult is determined by cell proliferation and rearrangement during larval development[@godt1995mechanisms;@king1968development]. In this species, the number of ovarioles can vary between the left and right ovaries within an individual, as well as across individuals within a population[@telonis2005genetic;@bergland2008quantitative]. This variation is derived primarily from variation in the number of “terminal filament precursor cells”[@green2012convergent;@sarikaya2019reproductive], as well as from variation in the number of those precursor cells that group together to form the structure that initiates ovariole formation, known as a “terminal filament”[@sarikaya2012roles]. Across species of _Drosophila_, variation in average adult ovariole number results primarily from variation in the average number of terminal filament precursor cells[@sarikaya2019reproductive].

When considering the developmental processes that could give rise to invariant ovariole number, we propose that the major determinants of ovariole number known from _Drosophila_ may not apply. To achieve an invariant ovariole number, these processes might instead include mechanisms for strict counting of individual cells or discrete cell subpopulations. In the former, if the cells that ultimately comprised a terminal filament were derived by mitotic division from a single progenitor, rather than by cellular rearrangements as is the case in _Drosophila_[@godt1995mechanisms], then an invariant ovariole number could be achieved via strict control of the number of precursor cells. Alternatively, an invariant ovariole number could be achieved by partitioning the starting population of precursor cells into a tightly regulated number of subpopulations. This would again be a departure from known mechanisms in _Drosophila_, in which a variable number of precursor cells are gathered into terminal filaments until the population is depleted[@godt1995mechanisms;@sarikaya2012roles]. The determining factor for partitioning the precursor pool could be, for example, a spatially variable morphogen emanating from adjacent tissues [@lawrence1996morphogens] or a reaction-diffusion patterning process [@kondo2010reaction] within the developing ovary, as has been shown to generate fixed numbers of multicellular structures in other developmental contexts[@salazar2008tooth;@hatini2001divide;@clark2019arthropod]. These predictions could be tested by characterizing the dynamics of cell number and position across invariant lineages, and making comparisons to corresponding data from their variable relatives.

The evolutionary transitions between variable and invariant ovariole number are reminiscent of other quantitative traits across multicellular life, including patterns of variability and invariance in arthropod segment number[@arthur1999pattern;@vedel2008temperature], vertebrate digit number[@holder1983developmental;@saxena2017origins], or the number of angiosperm floral organs[@AmbroseBarbaraA2012Teop;@kitazawa2014developmental]. Across these systems, the evolutionary history of morphogenetic counting mechanisms is poorly understood. We suggest that insect ovariole number presents an ideal case to study this phenomenon. In particular, we note the evidence that invariance has evolved convergently at least nine times, as well as the evidence of several reversals back to variability from an invariant ancestral state (Fig. 5). These convergent lineages provide an opportunity to test the predictability of evolutionary changes to counting mechanisms, by asking whether convergent evolution of invariance involves convergent canalization of shared molecular mechanisms.

# Acknowledgements {-}

This work was supported by NSF GRFP DGE1745303 to SHC and funds from Harvard University to support SHC and CGE. NM was supported by the E3 REU program at Harvard University. We thank members of the Extavour Lab for discussion.

# Author Contributions {-}

SHC led the data collection, analysis, and writing of the manuscript. NM performed the initial literature search, data collection, and analysis. BASdM assembled all phylogenies for analysis. SD, BASdM, and CGE contributed to data analysis, visualization, discussion, and writing.

# Competing Interests {-}

The authors declare no competing interests.

# References {-}

### 2020-07-07 manuscript_supplement.Rmd
code, data, and manuscript files to reproduce all analyses and figures

---
title: Repeated loss of variation in insect ovary morphology highlights the role of developmental constraint in life-history evolution - Supplementary Methods
output: 
  bookdown::pdf_document2:
    keep_tex: yes
    latex_engine: pdflatex
    toc: no
  #bookdown::html_document2: default
bibliography: manuscript_supplement.bib
csl: nature.csl
link-citations: yes
header-includes:
 \usepackage{float}
 \floatplacement{figure}{H}
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(echo=FALSE, include = FALSE, warning=FALSE, message=FALSE, cache=TRUE)
```

```{r preliminaries,cache=FALSE}
library(BAMMtools)
library(phytools)
library(coda)
library(OUwie)
library(corHMM)
library(grid)
library(gridExtra)
library(gtable)
library(geiger)
library(arbutus)
library(dplyr)
library(ggtree)
library(ape)
library(ggplot2)
library(RColorBrewer)
theme_set(theme_classic())
options(tinytex.clean = FALSE)
```

\makeatletter
\renewcommand{\thefigure}{S\@arabic\c@figure}
\renewcommand{\thetable}{S\@arabic\c@table}
\makeatother

```{r load_data, cache=FALSE}
# This script reads in and formats the ovariole number dataset
source("analysis/get_ovariole_number_data.R")
```

Samuel H Church^1^\*, Bruno A. S. De Medeiros^1,2^, Seth Donoughe^1,3^, Nicole L. Márquez Reyes^4^, Cassandra G. Extavour^1,5^\*

^1^ Department of Organismic and Evolutionary Biology, Harvard University, Cambridge, MA 02138, USA

^2^ Smithsonian Tropical Research Institute, Panama City, Panama

^3^ Department of Molecular Genetics and Cell Biology, University of Chicago, Chicago, IL 60637, USA

^4^ Department of Biology, Universidad de Puerto Rico en Cayey, Cayey 00736, PR

^5^ Department of Molecular & Cellular Biology, Harvard University, Cambridge, MA 02138, USA

\* Corresponding author

```{=latex}
% Trigger ToC creation in LaTeX
\tableofcontents
```

# Gathering ovariole number records {#sec:gathering}

We searched the published literature for references to insect ovariole number using a predetermined set of 131 search terms, entered into Google Scholar (scholar.google.com) between June and October of 2019. Each search term consisted of an insect taxonomic group and the words "ovariole number". This list was created to include all insect orders, many large insect families, and groups well-represented in the insect egg dataset. The list of search terms is available in the supplementary file 'ovariole_number_search_terms.tsv'.

For each search term, we evaluated all publications in the first page of results (ten publications). For 61 search terms that had a large number of informative hits, significant representation in the egg dataset, or that corresponded to very speciose groups, we evaluated an additional 20 publications. If a publication reported ovariole number for one or more insect species, we recorded the following information: (1) genus, (2) species name, when available, (3) taxonomic order, (4) sample size, when available, (5) ovariole number, and (6) additional notes (e.g. for eusocial insects, whether the observation was made in a reproductive or non-reproductive individual). This dataset is made publically available at Dryad (doi:10.5061/dryad.59zw3r253).

Ovariole number was recorded as either an average with deviations, a range, or a single total value, with priority for recording given in that order when multiple data types were available. Ovariole number was recorded as the total number of ovarioles per female, summing over both the left and right adult ovaries. When authors reported ovariole number from a single ovary, the total value was calculated by doubling the reported value. When authors described differences between the two ovaries, this information was recorded in an additional column.

Using this approach, we gathered `r nrow(raw_ON)` records for ovariole number from 460 publications. A full list of publications is provided in the supplementary files 'ovariole_number_bibliography.pdf'. We matched the scientific names to additional taxonomic information using the software TaxReformer[@TaxReformer] and found additional taxonomic data for `r nrow(matched)` of the `r nrow(raw_ON)` records. We verified that TaxReformer had found a valid match by comparing the originally recorded taxonomic order to the order populated by online databases, and removed `r nrow(raw_ON %>% filter(problem == "taxonomy"))` taxonomic records for which these values did not match. For all subsequent analyses, we also excluded observations made in non-reproductive individuals from eusocial species (workers), as well as two observations which represented significant outliers and could not be validated using additional sources or figures[@su2015testicular;@hernandez2017biology].

# Phylogenetic trees {#sec:trees}

The analyses herein were performed using the insect phylogeny published in Church et al, 2019[@church2019insect], unless otherwise specified. This phylogeny was constructed by combining ribosomal genetic data from `r length(tree$tip.label)` insect genera, published originally in the SILVA database[@Yilmaz2014], with constrained, time-calibrated nodes for each insect order, published originally in Misof et al, 2014[@Misof:2014jv]. This phylogeny is enriched for insect genera with records in the egg trait dataset, and also has considerable overlap with the genera included in this ovariole number dataset (`r ON %>% filter(genus %in% tree$tip.label) %>% distinct(genus) %>% nrow()` genera). For generalized least squares analyses and trait model comparisons, analyses were performed over a posterior distribution of trees associated with this published phylogeny[@church2019insect].

Analyses of insect family-level ovariole number, egg size, and body size were performed using the insect phylogeny published in Rainford et. al, 2014[@rainford2014phylogenetic].

Analyses of Drosophilidae ovariole number, egg size, and body size were performed using a phylogeny newly assembled for this study. Published genetic data for 317 Drosophilidae species were retrieved from NCBI in June of 2019[@katoh2017multiple;@magnacca2015rapid;@lapoint2014phylogenetics;@o2011phylogenetic;@lapoint2011phylogenetic;@bonacum2005phylogeny;@o2004phylogenetic;@lapoint2013diversification;@baker1997multiple]. These data encompassed 41 gene regions including mitochondrial, nuclear, and ribosomal genes. When multiple sequences for a gene region were available from the same species, the one with the least amount of missing data was selected. Each gene region was aligned using the program MAFFT[@Katoh2013], model auto selected). Alignments were concatenated and trimmed to 3% occupancy across species using the program phyutlity[@smith2008phyutility]. Documentation including accession numbers, sequence files, and alignments are available in the supplementary directory 'https://github.com/shchurch/insect_ovariole_number_evolution_2020/phylogeny/Drosophilidae_sequences/'.

To the extent possible, sequence data were not curated beyond what was downloaded from NCBI, with the following exceptions: [1] two sequences labeled as 16S that did not align to other 16S sequences were removed manually. [2] COI sequences were trimmed to remove regions with large quantities of missing sites prior to alignment. [3] One species name (_D. albovittata_) was corrected for typographical error. [4] Sequences identified as _Drosophila crassifemur_ were taxonomically corrected to _Scaptomyza crassifemur_[@o2003placement].

Phylogenetic estimation of the Drosophilidae data were performed using RAxML (model GTRGAMMA), setting the split between Hawaiian _Drosophila_ and _Scaptomyza_ as the root of the tree[@katoh2017multiple;@o2011phylogenetic]. The final tree was pruned to remove undescribed species (e.g. _Drosophila_ nr _dorsigera_), and was time-calibrated using the R package ape, function chronos (default parameters, version `r packageVersion('ape')`)[@paradis2004ape]. This tree is available in the supplementary file 'https://github.com/shchurch/insect_ovariole_number_evolution_2020/phylogeny/Drosophilidae_time_calibrated.tre'.

# Phylogenetic regressions

## Combining datasets {#sec:combining}

We combined the data we collected on total ovariole number with existing datasets of egg size and shape[@church2019dataset], insect lifetime fecundity and adult dry body weight[@Gilbert2007], average adult body length per insect family[@rainford2016phylogenetic], and several lineage-specific measures of adult body size[@iwata1966large;@Iwata1966a;@waloff1954number;@starmer2003phylogenetic;@reinhardt2005low].

Ovariole number and egg size[@church2019insect] data were combined by matching records at the species level (Fig 2a). When multiple records existed for a given species, the dataset was randomly shuffled and a single matching record was selected. This variation across records for the same species was accounted for in regressions by reshuffling and matching records at each iteration of the analysis. In order to test whether results were robust with a larger sample size, we also matched records at the genus level following the same reshuffling method (Fig. \ref{fig:PGLS-egg-ON-genus}).

Average adult body length per insect family[@rainford2016phylogenetic] was matched to the average ovariole number for the corresponding families (Fig. \ref{fig:PGLS-egg-ON-family}). The Rainford et al, 2016[@rainford2016phylogenetic] dataset contains a small number of average adult body lengths at the order level (e.g. Strepsiptera), which were matched to their equivalent group in the ovariole number dataset. To test the effect of uncertainty in the estimated average ovariole number on our results, the dataset for each family was downsampled by half at each iteration of the regression analysis.

Ovariole number, egg volume, lifetime fecundity, and adult body weight[@Gilbert2007] were combined by matching records at the species level and genus level, using the same method as described above (Figs. 2b, 3, and \ref{fig:PGLS-Gilbert-name}). We excluded one value from this dataset which appeared to include a typographical error for lifetime fecundity (Hymenoptera: Trichogrammatidae, lifetime fecundity recorded as 0.1[@Gilbert2007]).

Several lineage-specific measurements for body size were matched to the ovariole number and egg size datasets, as follows: Drosophilidae thorax length[@starmer2003phylogenetic] was matched at the species level (Fig. 2c), Orthoptera body length[@waloff1954number] was matched at the genus level (Fig. 2d), Hymenoptera mesosoma width[@Iwata1966a] was matched at the genus level, and Curculonoidea elytra length[@iwata1966large] was matched at the genus level (Fig. \ref{fig:PGLS-Coleoptera-Hymenoptera}).

## Phylogenetic Generalized Least Squares (PGLS) analyses {#sec:PGLS}

```{r PGLS-posterior, include=TRUE, cache=FALSE}
#source("analysis/run_pgls_over_posterior_distribution.R")
load("analysis/pgls_results_June2020/run_pgls_over_posterior_distribution.RData")

# analysis names in table 
pgls_analysis_names <- c("ovariole number vs egg volume",
  "ovariole number vs egg volume",
  "ovariole number vs egg volume, residuals to body length",
  "ovariole number vs egg volume, residuals to body length, excluding Trigonalidae",
  "ovariole number vs egg volume, residuals to body weight",
  "ovariole number vs egg volume, residuals to body weight",
  "ovariole number vs lifetime fecundity",
  "ovariole number vs lifetime fecundity",
  "Curculionoidea ovariole number vs egg volume, residual to elytra length",
  "Orthoptera ovariole number vs egg volume, residuals to body length",
  "Hymenoptera ovariole number vs egg volume, residuals to mesosoma width",
  "Drosophilidae ovariole number vs egg volume, residuals to thorax length")
# level at which each PGLS was performed 
taxonomic_levels <- c("species","genus","family","family","species","genus","species","genus","genus","genus","genus","species")

sp <- " -- " # separator in tables

# write table
all_table <- as.data.frame(do.call(rbind, allometry_tables)) %>% round(.,3) %>% mutate(analysis = pgls_analysis_names, "taxonomic level" = taxonomic_levels,slope = paste(sprintf("%.3f",slope_min),sprintf("%.3f",slope_max),sep=sp),
      intercept = paste(sprintf("%.3f",int_min),sprintf("%.3f",int_max),sep=sp),
      "p-value" = paste(sprintf("%.3f",pval_min),gsub("0.000","<0.000",sprintf("%.3f",pval_max)),sep=sp),
      "taxa" = taxa) 
knitr::kable(all_table  %>% select(c("analysis","taxonomic level","slope","intercept","p-value","taxa")), "latex",caption="Results of PGLS analysis of ovariole number, egg size, and fecundity across a posterior distribution", booktabs = T, linesep = "") %>%
  kableExtra::kable_styling(latex_options = "striped") %>%
  kableExtra::column_spec(1, width = "16em") %>%
  kableExtra::kable_styling(latex_options = "hold_position")  %>%
  kableExtra::kable_styling(font_size = 8) 
```

```{r PGLS-body-posterior, include=TRUE, cache=FALSE}
#source("analysis/run_pgls_body_ovariole_number.R")
load("analysis/pgls_results_June2020/run_pgls_body_ovariole_number.RData")

# analysis names in table 
pgls_analysis_names <- c("ovariole number vs body length",
  "ovariole number vs body weight",
  "ovariole number vs body weight",
  "Curculionoidea ovariole number vs elytra length",
  "Orthoptera ovariole number vs body length",
  "Hymenoptera ovariole number vs mesosoma width",
  "Drosophilidae ovariole number vs thorax length")
# level at which each PGLS was performed 
taxonomic_levels <- c("family","species","genus","genus","genus","genus","species")

# write table 
body_all_table <- as.data.frame(do.call(rbind, body_allometry_tables)) %>% round(.,3) %>% mutate(analysis = pgls_analysis_names, "taxonomic level" = taxonomic_levels,slope = paste(sprintf("%.3f",slope_min),sprintf("%.3f",slope_max),sep=sp),
       intercept = paste(sprintf("%.3f",int_min),sprintf("%.3f",int_max),sep=sp),
       "p-value" = paste(sprintf("%.3f",pval_min),gsub("0.000","<0.000",sprintf("%.3f",pval_max)),sep=sp),
       "taxa" = taxa)
knitr::kable(body_all_table %>% select(c("analysis","taxonomic level","slope","intercept","p-value","taxa")), "latex",caption="Results of PGLS analysis of ovariole number and body size across a posterior distribution", booktabs = T, linesep = "") %>%
  kableExtra::kable_styling(latex_options = "striped") %>%
  kableExtra::column_spec(1, width = "16em") %>%
  kableExtra::kable_styling(latex_options = "hold_position")  %>%
  kableExtra::kable_styling(font_size = 8) 
```

For each regression, we repeated a Phylogenetic Generalized Least Squares (PGLS) analysis 1000 times, accounting for phylogenetic and phenotypic uncertainty, using the R packages ape (version `r packageVersion('ape')`)[@paradis2004ape] and nlme (version `r packageVersion('nlme')`)[@pinheiro2014r]. In these analyses we used a Brownian Motion based covariance matrix for traits.

For regressions at the species and genus level, we reshuffled and matched records at each iteration to account for variation across records for the same taxon. For regressions at the family level we recalculated the average ovariole number per insect family, downsampling the representation for each family by half. No posterior distribution was made available with the published family level phylogeny[@rainford2014phylogenetic].

To account for body size, we calculated the phylogenetic residuals[@revell2010phylogenetic] of each trait to body size, and then compared the evolution of these residuals using a PGLS regression.

```{r PGLS-setup}
# build body size dataset
# set factor for resampling the body size dataframe, here no downsampling used
downsample_factor <- 1.0
fam_count_threshold <- 1
source("analysis/combine_body_egg_datasets.R")
source("analysis/pgls_functions.R") # get functions for plotting PGLS results
```

```{r PGLS-egg-ON-name}
### PGLS egg volume vs ON, species level pgls
egg_ON_name_pgls <- egg_ON_name %>% select(genus,vol,on,group) %>% rename(rank = genus, trait1 = on, trait2 = vol)

# this function helps format scientific notation correctly
scientific_10 <- function(x) {
  parse(text=gsub("1e", "10^", scales::scientific_format()(x)))
}

pdf("manuscript_figures/panels/FigEgg_ON_name.pdf",height=5,width=7.5,useDingbats = F)
  g1 <- ggplot(egg_ON_name_pgls,aes(x = trait1, y = trait2, color = group)) +
      geom_point() +
      scale_color_manual(values = mrk) +
      scale_x_log10(breaks=c(1,10,100,1000,10000)) +
      scale_y_log10(label=scientific_10,breaks=c(0.00001,0.0001,0.001,0.01,0.1,1,10,100,100)) +
      theme(legend.position = "none") +
      xlab('ovariole number') +
      ylab(bquote('egg volume '(~mm^3)))
  print(g1)
invisible(dev.off())

### Gilbert egg volume vs ON, independent body weight, genus level pgls
egg_ON_gilbert_genus_pgls <- egg_ON_gilbert_genus %>% select(genus,logvol,logon,logweight,group) %>% mutate(rank = genus, trait1 = logon, trait2= logvol, indep = logweight)
pdf("manuscript_figures/panels/FigEgg_gilbert_weight_residuals.pdf",height=2.5,width=2.5,useDingbats = F)
run_plotting_resid_pgls(egg_ON_gilbert_genus_pgls,tree,"gilbert_genus") + xlab("residual of ovariole number") +
  ylab("residual of egg volume")
invisible(dev.off())

### Drosophilidae pgls
DIP_pgls <- DIP_body_egg_ON_name %>% select(name,logvol,logon,logbody,group) %>% rename(rank = name, trait1 = logon, trait2= logvol, indep = logbody)
pdf("manuscript_figures/panels/FigEgg_Drosophila_residuals.pdf",height=2.5,width=2.5,useDingbats = F)
run_plotting_resid_pgls(DIP_pgls,dros_tree,"DIP") + xlab("residual of ovariole number") +
  ylab("residual of egg volume")
invisible(dev.off())

### Orthoptera pgls
ORT_pgls <- ORT_body_egg_ON_genus %>% select(genus,logvol,logon,logbody,group) %>% rename(rank = genus, trait1 = logon, trait2= logvol, indep = logbody)
pdf("manuscript_figures/panels/FigEgg_Orthoptera_residuals.pdf",height=2.5,width=2.5,useDingbats = F)
run_plotting_resid_pgls(ORT_pgls,tree,"ORT") + xlab("residual of ovariole number") +
  ylab("residual of egg volume")
invisible(dev.off())
```

\clearpage

```{r PGLS-egg-ON-genus, fig.cap="**Egg volume vs ovariole number, matching records at the genus level.**", fig.width=7.5, fig.height=5, include=TRUE}
### PGLS egg volume vs ON, species level pgls
egg_ON_genus_pgls <- egg_ON_genus %>% select(genus,logvol,logon,group) %>% rename(rank = genus, trait1 = logon, trait2 = logvol)
run_plotting_pgls(egg_ON_genus_pgls,tree,"egg_ON_genus") +
  xlab(bquote('ovariole number,'~log[10])) +
  ylab(bquote('egg volume,'~mm^3~','~log[10]))
```

```{r PGLS-Gilbert-name, fig.cap="**Egg volume vs ovariole number, phylogenetic residuals to dry adult body weight, matching records at the species level.**", fig.width=3, fig.height=3, include=TRUE}
### Gilbert body weight, species level pgls
egg_ON_gilbert_name_pgls <- egg_ON_gilbert_name %>% select(genus,logvol,logon,logweight,group) %>% mutate(rank = genus, trait1 = logon, trait2= logvol, indep = logweight)
run_plotting_resid_pgls(egg_ON_gilbert_name_pgls,tree,"gilbert_name") + xlab("residual of ovariole number") +
  ylab("residual of egg volume")
```

```{r PGLS-egg-ON-family, fig.cap="**Family average egg volume vs ovariole number, phylogenetic residuals to adult body length.** Point outlined in black = Hymenoptera: Trigonalidae, an outlier that drives the relationship.",  fig.width=3, fig.height=3, include=TRUE}
### PGLS egg volume vs ON, family averages
body_egg_ON_family_pgls <- body_egg_ON_full %>% select(rank,logvol,logon,logbody,group) %>% rename(trait1 = logon, trait2= logvol, indep = logbody)
egfam1 <- run_plotting_resid_pgls(body_egg_ON_family_pgls,fam_tree,"body_egg_ON_family")

egfam1 + xlab("residual of average ovariole number") +
  ylab("residual of average egg volume") +
  geom_point(data = (egfam1$data %>% filter(rank == "Trigonalidae")),aes(fill=group),color="black",stroke=1,pch=21)
```

```{r PGLS-Coleoptera-Hymenoptera, fig.cap="**Additional lineage-specific comparisons of egg volume vs ovariole number, phylogenetic residuals to body size, matching records at the genus level.** Weevils (Curculionoidea, left) were measured using elytra length and wasps (Hymenoptera, right) were measured using mesosoma width.", fig.width=6, fig.height=3, include=TRUE}
### Curculionoidea pgls
COL1_pgls <- COL1_body_egg_ON_genus %>% select(genus,logvol,logon,logbody,group) %>% rename(rank = genus, trait1 = logon, trait2 = logvol, indep = logbody)
theme_update(plot.title = element_text(hjust = 0.5))
col1 <- run_plotting_resid_pgls(COL1_pgls,tree,"COL1") + xlab("residual of ovariole number") +
  ylab("residual of egg volume") + ggtitle("Coleoptera: Curculionoidea")

### Apoidea pgls
HYM_pgls <- HYM_body_egg_ON_genus %>% select(genus,logvol,logon,logbody,group) %>% rename(rank = genus, trait1 = logon, trait2 = logvol, indep = logbody)
hym1 <- run_plotting_resid_pgls(HYM_pgls,tree,"HYM") + xlab("residual of ovariole number") +
  ylab(element_blank()) + ggtitle("Hymenoptera")

grid.arrange(col1,hym1,ncol=2)
```

```{r PGLS-body-plot, fig.cap="**Adult body size vs ovariole number.** Adult body weight compared at the genus level (left), and average adult body length compared at the family level (right).", fig.width=6, fig.height=3, include=TRUE}
### Gilbert body weight vs ON pgls
ON_body_gilbert_genus_pgls <- egg_ON_gilbert_genus %>% select(genus,logon,logweight,group) %>% mutate(rank = genus, trait1 = logon, trait2 = logweight)
gilbert_body <- run_plotting_pgls(ON_body_gilbert_genus_pgls,tree,"body_gilbert") + xlab(bquote('ovariole number,'~log[10])) +
  ylab(bquote('adult dry body weight (g),'~log[10]))

### PGLS body length vs ON, family averages
ON_body_family_pgls <- body_egg_ON_full %>% select(rank,logon,logbody,group) %>% rename(trait1 = logon, trait2 = logbody)
family_body <- run_plotting_pgls(ON_body_family_pgls,fam_tree,"body_egg_ON_family") + xlab(bquote('average ovariole number,'~log[10])) +
  ylab(bquote('average adult body length (mm),'~log[10]))

grid.arrange(gilbert_body,family_body,ncol=2) 
```

```{r PGLS-fecundity-genus}
### Gilbert fecundity vs ON, genus level
fecundity_genus_pgls <-  ON_gilbert_genus %>% select(genus,fecundity,on,group) %>% mutate(rank = genus, trait1 = on, trait2= fecundity) 
pdf("manuscript_figures/panels/FigFecundity.pdf",height=4,width=4,useDingbats = F)
  g1 <- ggplot(fecundity_genus_pgls,aes(x = trait1, y = trait2, color = group)) +
      geom_point() +
      scale_color_manual(values = mrk) +
      scale_x_log10() +
      scale_y_log10() +
      theme(legend.position = "none") +
      xlab('ovariole number') +
      ylab('lifetime fecundity')
  print(g1)
invisible(dev.off())
```

```{r PGLS-fecundity-name, fig.cap="**Lifetime fecundity vs ovariole number, matching records at the species level.**", fig.width=3, fig.height=3, include=TRUE}
### Gilbert fecundity vs ON, species level
fecundity_name_pgls <-  ON_gilbert_name %>% select(genus,logfecundity,logon,group) %>% mutate(rank = genus, trait1 = logon, trait2= logfecundity)
run_plotting_pgls(fecundity_name_pgls,tree,"fecundity_name") + xlab(bquote("ovariole number,"~log[10])) +
  ylab(bquote("lifetime fecundity,"~log[10]))
```

# Evolution of nurse cells

## Combining datasets

We used the descriptions of the mode of oogenesis recorded by Büning[@buning1994insect]. This author catalogued the ovary morphology for 136 insect genera, categorizing them into four modes: those without nurse cells (panoistic), with nurse cells adjacent to each clonally related, developing oocyte (polytrophic meroistic), with all nurse cells located in the germarium (telotrophic meroistic), and a unique mode of oogenesis reported only in Strepsiptera (reduced polytrophic meroistic ovaries).

Of the 136 genera observed by Büning, 70 are represented in the phylogeny used here[@church2019insect]. Another 36 come from families or orders that have representative genera in the phylogeny, and within which all observations have the same recorded mode of oogenesis, when more than one was recorded. Therefore, we used a substitute genus as the phylogenetic tip for these 36 groups, bringing the total overlap between dataset and phylogeny to 106 taxa.

## Reconstructing evolutionary shifts in oogenesis mode

```{r reconstruct-ovary, include=TRUE, fig.cap = "**Ancestral state reconstruction of oogenesis mode, full phylogeny.** Scale bar indicates 100 million years. Gray = panoistic, red = telotrophic meroistic, cyan = polytrophic meroistic, black = unique meroistic mode found in Strepsiptera.", fig.width = 5, fig.height=9}
#source("analysis/reconstruct_ancestral_ovary_morphology_type.R")
load("analysis/mode_of_oogenesis_results_June2020/reconstruct_ancestral_ovary_morphology_type.RData")

# Extract estimated tip states
ovary_nums <- data.frame(num=as.numeric(as.factor(unique(ovary_table$ovary))),ovary=unique(ovary_table$ovary),stringsAsFactors=F)

est_tips <- data.frame("state1" = round(pp$tip.states,0)[,1],
            "state2" = round(pp$tip.states,0)[,2],
            "state3" = round(pp$tip.states,0)[,3],
            "state4" = round(pp$tip.states,0)[,4]) %>% 
      mutate(discrete = ifelse(state1 == 1,1,
                          ifelse(state2 == 1,2,
                            ifelse(state3 == 1,3,4))),
          genus = pp$phy$tip.label) %>% 
      select(genus,discrete) %>% 
      mutate("ovary_regime"=plyr::mapvalues(discrete,ovary_nums$num,ovary_nums$ovary))

ovary_cols <- c("black","#949494","#00c2c2","red")
names(ovary_cols) <- c("Mstar","P","PM","TM")

### Plot the full oogenesis mode ASR on the phylogeny
branch_cols <- c(as.integer(plyr::mapvalues(pp$phy$tip.label,est_tips$genus,est_tips$discrete)),pp$phy$node.label)
plot(pp$phy,show.tip.label=F,edge.width=1,edge.col=plyr::mapvalues(branch_cols[pp$phy$edge[,2]],seq(1:length(ovary_cols)),ovary_cols))
add.scale.bar()

### Plot just the reconstruction for the originally recorded 106 tips
pp_original_records <- drop.tip(pp$phy,tree_dataset %>% filter(!(genus %in% new_ovary_table$genus)) %>% pull(genus))
branch_cols <- c(as.integer(plyr::mapvalues(pp_original_records$tip.label,est_tips$genus,est_tips$discrete,warn_missing=F)),pp_original_records$node.label)
pdf("manuscript_figures/panels/FigBuning_Reconstruction.pdf",height=4.75,width=2.5,useDingbats = F)
  plot(pp_original_records,show.tip.label=F,edge.width=1,edge.col=plyr::mapvalues(branch_cols[pp_original_records$edge[,2]],seq(1:length(ovary_cols)),ovary_cols))
  add.scale.bar()
invisible(dev.off())
```

Using these 106 records for ovary type, we reconstructed the ancestral state at each node of the published phylogeny of `r length(pp$phy$tip.label)` insect genera using an equal-rates model that allows for missing data (Fig. \ref{fig:reconstruct-ovary}, R package corHMM,[@beaulieu2013identifying] version `r packageVersion('corHMM')`, function rayDISC, node.states = 'marginal').

## Oogenesis mode model comparison {#sec:oogenesis-model}

```{r ovary-model-comparison, include=TRUE}
# get corrected AIC values from results
get_aicc <- function(ou_res) {
  BM1 <- sapply(ou_res,function(x){x[[1]]$AICc})
  BMS <- sapply(ou_res,function(x){x[[2]]$AICc})
  OU1 <- sapply(ou_res,function(x){x[[3]]$AICc})
  OUM <- sapply(ou_res,function(x){x[[4]]$AICc})
  ouwie_res <- data.frame(BM1 = BM1, BMS = BMS, OU1 = OU1, OUM = OUM) %>%
        rowwise() %>%
        mutate(min = min(BM1,BMS,OU1,OUM)) %>%
        ungroup() %>%
        mutate(dBM1 = BM1 - min,
          dBMS = BMS - min,
          dOU1 = OU1 - min,
          dOUM = OUM - min,
          bm1_bms = (dBM1 - dBMS > 2),
          bm1_ou1 = (dBM1 - dOU1 > 2),
          bm1_oum = (dBM1 - dOUM > 2),
          ou1_oum = (dOU1 - dOUM > 2))
  return(ouwie_res)
}

# calculate the mean AICc value over the posterior distribution
get_mean_aicc <- function(res) {
  mean_ouwie_res <- res %>% select(BM1,BMS,OU1,OUM,dBM1,dBMS,dOU1,dOUM) %>%       
    group_by() %>%
    summarize_all("mean")
  return(mean_ouwie_res)
}

# count the number of AICc > 2
get_count_aicc <- function(res) {
  count_ouwie_res <- data.frame(bm1_bms = sum(res$bm1_bms),
          bm1_ou1 = sum(res$bm1_ou1),
          bm1_oum = sum(res$bm1_oum),
          ou1_oum = sum(res$ou1_oum))
  return(count_ouwie_res)
}

# use those functions to evaluate results for each trait of interest
load("analysis/mode_of_oogenesis_results_June2020/asym_ovary_model_comparison.RData") # egg asymmetry
asym_results <- ouwie_results
load("analysis/mode_of_oogenesis_results_June2020/curv_ovary_model_comparison.RData") # egg curvature
curv_results <- ouwie_results
load("analysis/mode_of_oogenesis_results_June2020/egg_ovary_model_comparison.RData") # egg volume
egg_results <- ouwie_results
load("analysis/mode_of_oogenesis_results_June2020/ar_ovary_model_comparison.RData") # egg aspect ratio
ar_results <- ouwie_results
load("analysis/mode_of_oogenesis_results_June2020/on_ovary_model_comparison.RData") # ovariole number
on_results <- ouwie_results

asym_res <- get_aicc(asym_results)
asym_mean_res <- get_mean_aicc(asym_res)
asym_count_res <- get_count_aicc(asym_res)

egg_res <- get_aicc(egg_results)
egg_mean_res <- get_mean_aicc(egg_res)
egg_count_res <- get_count_aicc(egg_res)

ar_res <- get_aicc(ar_results)
ar_mean_res <- get_mean_aicc(ar_res)
ar_count_res <- get_count_aicc(ar_res)

curv_res <- get_aicc(curv_results)
curv_mean_res <- get_mean_aicc(curv_res)
curv_count_res <- get_count_aicc(curv_res)

on_res <- get_aicc(on_results)
on_mean_res <- get_mean_aicc(on_res)
on_count_res <- get_count_aicc(on_res)

# count the number of taxa in each analysis
sample_size <- tibble("analysis name" = c("ovariole number","egg volume","egg aspect ratio","egg asymmetry","egg curvature"),
            "taxa" = c(nrow(on_results[[1]][[1]]$data),
              nrow(egg_results[[1]][[1]]$data),
              nrow(ar_results[[1]][[1]]$data),
              nrow(asym_results[[1]][[1]]$data),
              nrow(curv_results[[1]][[1]]$data)))

# write table
mean_table <- rbind(on_mean_res,egg_mean_res,ar_mean_res,asym_mean_res,curv_mean_res) %>%
  mutate("analysis name" = c("ovariole number","egg volume","egg aspect ratio","egg asymmetry","egg curvature")) %>%
  select("analysis name",BM1,BMS,OU1,OUM)
knitr::kable(mean_table, "latex",caption="Average corrected AIC (AICc) value from model comparison", booktabs = T, linesep = "") %>%
  kableExtra::kable_styling(latex_options = "striped") %>%
  kableExtra::column_spec(1, width = "10em") %>%
  kableExtra::kable_styling(latex_options = "hold_position")

```

```{r ovary-model-comparison-count, include=TRUE}
count_table <- rbind(on_count_res,egg_count_res,ar_count_res,asym_count_res,curv_count_res)  %>%
  mutate("analysis name" = c("ovariole number","egg volume","egg aspect ratio","egg asymmetry","egg curvature")) %>%
  select("analysis name",bm1_bms,bm1_ou1,bm1_oum,ou1_oum) %>%
  rename("BMS vs. BM1" = bm1_bms,
        "OU1 vs. BM1" = bm1_ou1,
        "OUM vs. BM1" = bm1_oum,
        "OUM vs. OU1" = ou1_oum) %>% 
  left_join(.,sample_size,by="analysis name")
knitr::kable(count_table, "latex",caption="Results of model comparison analysis over posterior distribution, showing the number of iterations out of 100 where the difference in model fit ($\\Delta$AICc) was greater than 2.", booktabs = T, linesep = "") %>%
  kableExtra::kable_styling(latex_options = "striped") %>%
  kableExtra::column_spec(1, width = "10em") %>%
  kableExtra::kable_styling(latex_options = "hold_position")
```

Using the ancestral state reconstruction of evolutionary shifts in the mode of oogenesis, we inferred the most likely mode of oogenesis for all nodes and unobserved extant tips in the phylogeny (Fig. \ref{fig:reconstruct-ovary}). We then compared the fit of models of trait evolution that take into account these shifts in oogenesis mode against those that do not. These comparisons were performed with the R package OUwie[@beaulieu2012modeling]  (version `r packageVersion('OUwie')`).

Each analysis compared four models of evolution: single-rate Brownian Motion (BM1), multi-rate Brownian Motion (BMS), single-optimum Ornstein-Uhlenbeck (OU1), and an Ornstein-Uhlenbeck model with different optima for each mode of oogenesis (OUM).

These comparisons were repeated 100 times over a posterior distribution of trees. At each iteration we selected a random representative trait record for each genus in the phylogeny, when multiple records were available.

# Modeling rate of ovariole number change {#sec:arbutus}

## Parametric bootstrap of Brownian Motion model {#sec:bootstrap}

```{r arbutus, include=TRUE, fig.cap="**Bootstrap analysis of Brownian Motion model for ovariole number evolution, using the R package arbutus.** In each panel the red line represents the observed value and the black distribution represents the bootstrap simulation. See Section \\ref{sec:arbutus} for details on each parameter.",fig.width=9,fig.height=6}

# set up trait and tree for arbutus analysis
arbutus_tree <- drop.tip(tree,tree$tip.label[!(tree$tip.label %in% ON$genus)])
arbutus_trait <- ON %>% filter(genus %in% arbutus_tree$tip.label) %>%
  group_by(genus) %>% slice(1L)
arbutus_trait_vec <- arbutus_trait %>% pull(logon)
names(arbutus_trait_vec) <- arbutus_trait$genus

# fit a Brownian Motion model
fit_bm <- arbutus_trait_vec[order(match(names(arbutus_trait_vec),arbutus_tree$tip.label))] %>%
  fitContinuous(dat = ., phy =arbutus_tree, model="BM")

# run a parametric bootstrap of the model with arbutus
arbutus_fit_bm <- arbutus(fit_bm)

# plot the results
arbutus_names <- c("Mean of squared contrasts","Variation of contrast absolute values","Slope of contrast absolute values against variances","Slope of contrast absolute value against trait value","Slope of contrast absolute value against node age","D-statistic, difference from normality")
names(arbutus_fit_bm$obs) <- arbutus_names
names(arbutus_fit_bm$sim) <- arbutus_names
plot(arbutus_fit_bm,legend=F)
```

We evaluated the fit of a Brownian Motion (BM) model for ovariole number evolution using the R package arbutus[@pennell2015model], version `r packageVersion('arbutus')`  (Fig. \ref{fig:arbutus}). In this approach, a BM model is fit to the data (R package geiger[@harmon2007geiger], version `r packageVersion('geiger')`), and the resulting parameters of the model are used to simulate 1000 new datasets. Six statistical parameters are used to compare the phylogenetic contrasts of the observed data to the simulated data, and their interpretations are as follows[@pennell2015model]:

1. _Mean of squared contrasts_. The rate of evolution of ovariole number can be well estimated by the Brownian Motion model (the observed value falls within the null distribution).
2. _Coefficient of variation of the absolute value of the contrasts._ There is substantially more variation in contrasts than expected by chance, indicating heterogeneity in the rate of evolution beyond what a single-rate Brownian Motion model predicts (the observed value falls well outside the null distribution).
3. _Slope of a linear model fitted to the absolute value of the contrasts against their expected variances._ Contrasts are larger than expected on short branches in the phylogenetic tree, resulting in a negative slope. This could be explained by error in estimation of branch lengths.
4. _Slope of a linear model fitted to the absolute value of the contrasts against the ancestral state at the corresponding node._ The number of ovarioles is more correlated with contrast values than would be expected by chance. Phylogenetic nodes with a low ovariole number experience lower rates of evolution.
5. _Slope of a linear model fitted to the absolute value of the contrasts against node depth._ Contrast values are not correlated with time, falling within the null distribution. Therefore the rate of ovariole number change is not increasing or decreasing over time.
6. _The D statistic from a Kolmogorov-Smirnov test comparing the distribution of contrasts to an expected normal distribution._ The data do not fit a normal distribution of contrasts well, suggesting there are likely non-Brownian motion based processes at play (e.g. jump-diffusion processes).

## Assessing rate heterogeneity {#sec:BAMM}

```{r BAMM-analysis-setup}
### Select a continuous trait to analyze
trait_of_interest <- "on"
dat_of_interest <- ON

### name the analysis
name <- paste("BAMM",trait_of_interest,sep="_")

### Set up the data frame for analysis
dat <- dat_of_interest[sample(nrow(dat_of_interest)),] %>% # shuffle the dataset
  mutate(raw_trait = eval(parse(text = trait_of_interest))) %>%  # pull out genus and trait
  filter(!(is.na(raw_trait))) %>% select(genus,raw_trait) %>% # remove missing data
  group_by(genus) %>% summarize(mean_trait = mean(raw_trait)) %>% # calculate mean per genus
  mutate(trait = log10(mean_trait)) # transform the data

### Format the data frame for BAMM
bmm_red <- dat %>% mutate(taxa = genus) %>%
    filter(genus %in% tree$tip.label) %>%
    select(genus,trait)

### Use only taxa with phenotypic data
bmm <- drop.tip(tree,setdiff(tree$tip.label,bmm_red$genus))

### Write the files for BAMM
#write.table(bmm_red,file="bamm_trait.txt",sep="\t",row.names=F,col.names=F,quote=F)
#write.tree(bmm,file="bamm_tree.tre")

# Select appropriate priors
#priors <- setBAMMpriors(phy = bmm, traits = "bamm_trait.txt", outfile=NULL)
#priors[1] <- 10

#params <- c(treefile = 'bamm_tree.tre',
#        traitfile = 'bamm_trait.txt',
#        numberOfGenerations = '2000000000', # use 10^9 generations
#        mcmcWriteFreq = '50000',
#        eventDataWriteFreq = '1000000', # sample every 10^6
#        outName = 'bamm',
#        overwrite = '1',
#        printFreq = '1000000',
#        priors)

#generateControlFile('bamm.config', type = 'trait', params = as.list(params))

#### OUTSIDE OF R, RUN BAMM
#### bamm -c bamm.config
```

Given the result that our dataset contains substantial rate heterogeneity, we identified regions of the tree with high and low rates of ovariole number evolution using the software BAMM[@rabosky2014automatic], version 2.5.0. For this analysis, we calculated the average ovariole number for each genus in the insect phylogeny[@church2019insect]. Average ovariole number was log~10~ transformed, and the tree was filtered to include only tips for which there were corresponding ovariole number data (sample size = `r length(bmm$tip.label)`). We used the R package BAMMtools (version `r packageVersion('BAMMtools')`)[@rabosky2014bamm] to select priors, and ran BAMM for the maximum number of generations ($2*10^{9}$), sampling every 10^6^ generations.

```{r BAMM-analysis-convergence, include=TRUE, fig.cap="Convergence of trait diversification rate analysis", fig.width=10, fig.height=5}
### Assess convergence
mcmcout <- read.csv("analysis/BAMM_results_June2020/bamm_mcmc_out.txt", header=T)

#pdf(file=paste(name,"_convergence.pdf",sep=""),width=10,height=5)
  plot(mcmcout$logLik ~ mcmcout$generation,pch=16,cex=0.5,xlab="MCMC generation",ylab="Log-likelihood")
#dev.off()
```

```{r BAMM-analysis-burnin, include=TRUE, fig.cap="**Comparison of burn-in proportions.** The burn-in proportion that maximized the effective size was used in subsequent analyses.", fig.width=10, fig.height=5}
# calculate the effective size of the logL using different burn in values
bi <- seq(0.01,0.99,by=0.001) 
es_logl <- sapply(bi,function(x){effectiveSize(mcmcout[(x * nrow(mcmcout)):nrow(mcmcout),]$logLik)})

burnin <- bi[which.max(es_logl)]

#pdf(file=paste(name,"_effectiveSize.pdf",sep=""),width=10,height=5)
  plot(y=es_logl,x=bi,pch=16,cex=0.5,xlab="Proportion designated as burn-in",ylab="Effective sample size of log-likelihood")
#dev.off()

burnstart <- floor(burnin * nrow(mcmcout))
postburn <- mcmcout[burnstart:nrow(mcmcout), ]

es_nshifts <- effectiveSize(postburn$N_shifts) # this should be >200
es_logl <- effectiveSize(postburn$logLik) # this should be >200, though it has never gotten close
#es_nshifts
#es_logl
```

Convergence was evaluated both visually (Fig. \ref{fig:BAMM-analysis-convergence}) and numerically by comparing the effective sample size for number of shifts and log-likelihood to the standard recommended by the software (>200). We determined the most appropriate burn-in proportion to use by finding the maximum effective sample size of the log-likelihood across an array of possible burn-in proportions (Fig. \ref{fig:BAMM-analysis-burnin}). Running BAMM for the maximum possible number of generations and selecting the optimum burn-in\ref{fig:BAMM-analysis-burnin} resulted in an effective size for the number of shifts of `r round(es_nshifts,2)`, and for log-likelihood of `r round(es_logl,2)`. Repeated BAMM analyses showed similar distributions of high and low rate regimes, indicating the implications for ovariole number evolution are robust to uncertainty in rate estimates. See Supplemental Methods Section \ref{sec:BAMM} for details.

```{r BAMM-anlaysis-event-data}
### Read in the BAMM results and plot the shifts on the tree
bamm_tree <- ladderize(read.tree(file="analysis/BAMM_results_June2020/bamm_tree.tre"))

# get event data
edata <- getEventData(bamm_tree, eventdata = paste("analysis/BAMM_results_June2020/bamm_event_data.txt",sep=""), burnin=burnin, type="trait")
```

```{r BAMM-analysis-results, include=TRUE, fig.cap="**Best rate shift configuration from BAMM trait diversification analysis on ovariole number.** Purple = low rate of evolution, yellow = high rate of evolution.", fig.width=5, fig.height=10}
# plot the best tree as a phylorate plot
#pdf(file=paste(name,"_ratesplot.pdf",sep=""),width=10,height=12)
  pb_legend <- png::readPNG("analysis/datafiles/BAMM_legend.png")
  grid.arrange(ggplotify::as.grob(~plot(edata,lwd=2,legend=F,labels=F,cex=0.2,logcolor=T,breaksmethod="jenks",pal=c("#0D0887", "#CC4678", "#F0F921"))),rasterGrob(pb_legend),ncol=2,widths=c(10,1)) 
#dev.off()
```

```{r BAMM-anlaysis-mean-rates, include=TRUE, fig.cap="**Distribution of trait diversification rates.** Dotted line shows threshold used to assigned rate regimes.", fig.width=10, fig.height=5}
### Identify fixed lineages
# build a tree with branch lengths encoded as mean rate
brlen <- getMeanBranchLengthTree(edata)

# choose a low rate threshold 
lowrate <- 1e-04

# plot rates
#pdf(file=paste(name,"_meanrates.pdf",sep=""),width=10,height=5)
  hist(log10(brlen$phy$edge.length),breaks=200,col="black",xlab=bquote('Mean branch rate of trait diversification,'~log[10]),main="")
  abline(v = log10(lowrate), col="black", lwd=3, lty=2)
#dev.off()
```

The best configuration of rate shift regimes shows multiple independent clades with very low rates of evolution (Fig. \ref{fig:BAMM-analysis-results}). Visualizing the distribution of mean rates along branches revealed a discontinuous distribution, with one peak at a moderate rate of evolution and another at an extremely low rate, separated from the first peak by over six orders of magnitude (Fig. \ref{fig:BAMM-anlaysis-mean-rates}). We used this visualization to establish a threshold (`r lowrate`) for assigning a binary rate regime to each node in the phylogeny, categorizing them as above (variable) or below (invariant) a threshold that separates these two peaks.

```{r BAMM-analysis-mainfig}
# bin the rates into above or below the threshold
regimes <- data.frame(node=brlen$phy$edge[,2], # use the terminal nodes of each edge as the identifier
          brlen=brlen$phy$edge.length) %>% # pull the branch length = mean rate
      mutate(regime = ifelse(brlen<lowrate,"below","above")) # bin rates
mrca_node <- data.frame(node=getMRCA(bamm_tree,bamm_tree$tip.label),brlen=NA,regime="above") # set the root node as above
regimes <- rbind(regimes,mrca_node) %>% arrange(node) # add root

### Plot the tree with binned fixed lineages
# start with the bamm trees 
node_tree <- bamm_tree

# label the internal nodes to match the regime data
node_tree$node.label <- regimes %>% filter(node > length(node_tree$tip.label)) %>% pull(regime)

# build vector of branch colors, mapped to the order of terminal nodes in the tree
cols <- c("dark gray","black")
branch_cols <- plyr::mapvalues(regimes$regime[node_tree$edge[,2]],c("above","below"),cols)

### Assess trait evolution using the rate regimes
on_reg <- regimes[1:length(node_tree$tip.label),] %>% mutate(genus = node_tree$tip.label) %>% # pull out tips
  left_join(.,(ON %>% select(genus,on,group)),by="genus") %>% # add in trait data
  rename(rate = "brlen") %>% # recode for geiger
  select(genus,on,regime,rate,group)

# plot panels for figure:Diversity
size <- 0.8

# phylogeny
p0 <- ggtree(node_tree,size=0.35,color="dark gray") + scale_color_manual(values=mrk) + scale_fill_manual(values=mrk) + scale_y_reverse()  + theme_tree2()
p02 <- facet_plot(p0,panel="ovariole number",data=as.data.frame(on_reg),geom=geom_point,aes(x = log10(on),color = group),size=size,pch=16) + theme(legend.position = "none")

# add regimes
p <- ggtree(node_tree,size=0.35) %<+% (regimes) + aes(color=regime) + scale_color_manual(values=mrk) + scale_fill_manual(values=mrk) + scale_y_reverse()  + theme_tree2()
p2 <- facet_plot(p,panel="ovariole number",data=as.data.frame(on_reg),geom=geom_point,aes(x = log10(on),color = group),size=size,pch=16) + theme(legend.position = "none")

# change the widths of the phylogeny plot
p02t <- ggplot_gtable(ggplot_build(p02))
p02t$widths[7] = 2*p02t$widths[7] # in this case it was colmun 7 - reduce the width by a half

p2t <- ggplot_gtable(ggplot_build(p2))
p2t$widths[7] = 2*p2t$widths[7] # in this case it was colmun 7 - reduce the width by a half

# jitter plot
g1 <- ggplot(ON,aes(y = factor(group,levels=rev(group_levels)), x= on, color = group)) + geom_jitter(size=size,alpha=0.8,pch=16) + scale_color_manual(values=mrk) + scale_x_log10() + theme(legend.position = "none",axis.text.y=element_blank()) + xlab('ovariole number')
# histogram
g2 <- ggplot(ON,aes(x = on, fill=group)) + geom_histogram(bins=125) + scale_fill_manual(values=mrk) + scale_x_log10() + theme(legend.position = "none") + xlab('ovariole number')
g3 <- ggplot(ON %>% filter(genus %in% node_tree$tip.label),aes(x = on, fill=group)) + geom_histogram(bins=125) + scale_fill_manual(values=mrk) + scale_x_log10() + theme(legend.position = "none") + xlab('ovariole number')

pdf("manuscript_figures/panels/FigDiversity_Phylogeny.pdf",height=7,width=3.5,useDingbats = F)
  print(p2)
invisible(dev.off())
pdf("manuscript_figures/panels/FigDiversity_Phylogeny_Wide.pdf",height=7,width=3.5,useDingbats = F)
  grid.draw(p2t)
invisible(dev.off())
pdf("manuscript_figures/panels/FigDiversity_Phylogeny_NoRegime.pdf",height=7,width=3.5,useDingbats = F)
  grid.draw(p02t) 
invisible(dev.off())
pdf("manuscript_figures/panels/FigDiversity_Jitter.pdf",height=3.25,width=3.25,useDingbats = F)
  print(g1)
invisible(dev.off())
pdf("manuscript_figures/panels/FigDiversity_Histogram.pdf",height=3.25,width=4,useDingbats = F)
  print(g2)
invisible(dev.off())
pdf("manuscript_figures/panels/FigDiversity_Histogram_Short.pdf",height=1.5,width=2.75,useDingbats = F)
  print(g3)
invisible(dev.off())
```

## Rate model comparison

```{r BM-model-comparison}
# set up the data frame to compare models
ouwie_on_reg <- on_reg[sample(nrow(on_reg)),] %>% # shuffle the dataset
    group_by(genus) %>% slice(1L) %>% ungroup() %>% # choose one representative per genus
    mutate(species = genus,discrete = regime,continuous = log10(on)) %>% #transform the data
    select(species,discrete,continuous) %>% 
    as.data.frame()

### Compare BM models of evolution with and without regimes
BMS <- OUwie(node_tree,ouwie_on_reg,model="BMS",simmap.tree=FALSE,diagn=FALSE,root.station=F) #rates per regime
BM1 <- OUwie(node_tree,ouwie_on_reg,model="BM1",simmap.tree=FALSE,diagn=FALSE,root.station=F) #one rate
```

We tested whether a BM model of evolution that incorporates the binary state (variable or invariant) as independent rate regimes can better explain the distribution of ovariole numbers than a single rate BM model, by comparing model fit using the R package OUwie (version `r packageVersion('OUwie')`)[@beaulieu2012modeling]. We find that a multi-rate model is significantly favored over a single-rate model ($\Delta$ AICc `r round(BM1$AICc - BMS$AICc,2)`).

## Comparing rates of trait diversification

```{r BAMM-egg-setup}
tree <- genus_mcc_tree
trait_of_interest <- "vol"
dat_of_interest <- egg_database
dat <- dat_of_interest[sample(nrow(dat_of_interest)),] %>% # shuffle the dataset
  mutate(raw_trait = eval(parse(text = trait_of_interest))) %>%  # pull out genus and trait
  filter(!(is.na(raw_trait))) %>% select(genus,raw_trait) %>% # remove missing data
  group_by(genus) %>% summarize(mean_trait = mean(raw_trait)) %>% # calculate mean per genus
  mutate(trait = log10(mean_trait)) # transform the data

bmm_red <- dat %>% mutate(taxa = genus) %>%
    filter(genus %in% tree$tip.label) %>%
    select(genus,trait)
bmm <- drop.tip(tree,setdiff(tree$tip.label,bmm_red$genus))

#write.table(bmm_red,file="egg_bamm_trait.txt",sep="\t",row.names=F,col.names=F,quote=F)
#write.tree(bmm,file="egg_bamm_tree.tre")

#priors <- setBAMMpriors(phy = bmm, traits = "egg_bamm_trait.txt", outfile=NULL)
#priors[1] <- 10
#params <- c(treefile = 'egg_bamm_tree.tre',
#        traitfile = 'egg_bamm_trait.txt',
#        numberOfGenerations = '2000000000', # use 10^9 generations
#        mcmcWriteFreq = '50000',
#        eventDataWriteFreq = '1000000', # sample every 10^6
#        outName = 'egg_bamm',
#        overwrite = '1',
#        printFreq = '1000000',
#        priors)

#generateControlFile('egg_bamm.config', type = 'trait', params = as.list(params))

#### OUTSIDE OF R, RUN BAMM
#### bamm -c egg_bamm.config
```

```{r BAMM-egg-read}
egg_mcmcout <- read.csv("analysis/BAMM_results_June2020/egg_bamm_mcmc_out.txt", header=T)

egg_bi <- seq(0.01,0.99,by=0.001)
egg_es_logl <- sapply(egg_bi,function(x){effectiveSize(egg_mcmcout[(x * nrow(egg_mcmcout)):nrow(egg_mcmcout),]$logLik)})

egg_burnin <- egg_bi[which.max(egg_es_logl)]
egg_burnstart <- floor(egg_burnin * nrow(egg_mcmcout))
egg_postburn <- egg_mcmcout[egg_burnstart:nrow(egg_mcmcout), ]
egg_es_nshifts <- effectiveSize(egg_postburn$N_shifts)
egg_es_logl <- effectiveSize(egg_postburn$logLik)

egg_bamm_tree <- ladderize(read.tree(file="analysis/BAMM_results_June2020/egg_bamm_tree.tre"))
egg_edata <- getEventData(egg_bamm_tree, eventdata = paste("analysis/BAMM_results_June2020/egg_bamm_event_data.txt",sep=""), burnin=egg_burnin, type="trait")
```

```{r BAMM-egg-plot, include=TRUE, fig.cap="**Best rate shift configuration from BAMM trait diversification analysis on egg volume.** Purple = low rate of evolution, yellow = high rate of evolution.", fig.width=5, fig.height=10}
  pb_legend <- png::readPNG("analysis/datafiles/BAMM_legend.png")
  grid.arrange(ggplotify::as.grob(~plot(egg_edata,lwd=2,legend=F,labels=F,cex=0.2,logcolor=T,breaksmethod="jenks",pal=c("#0D0887", "#CC4678", "#F0F921"))),rasterGrob(pb_legend),ncol=2,widths=c(10,1))
```

```{r combine-rates, include=TRUE, fig.cap="**Rates of trait diversification of egg volume and ovariole number.** Points are colored by phylogenetic groups shown in Fig 1a. Dotted line shows threshold used to assigned rate regimes.", fig.width = 7, fig.height = 3.5}
egg_brlen <- getMeanBranchLengthTree(egg_edata)
egg_regimes <- data.frame(node=egg_brlen$phy$edge[,2],brlen=egg_brlen$phy$edge.length)

egg_reg <- egg_regimes[1:length(egg_bamm_tree$tip.label),] %>% mutate(genus = egg_bamm_tree$tip.label) %>% # pull out tips
  left_join(.,(egg_database %>% select(genus,vol,group)),by="genus") %>% # add in trait data
  rename(egg_rate = "brlen") %>% # recode for geiger
  select(genus,vol,egg_rate) %>% na.omit()

on_egg_reg <- on_reg[sample(nrow(on_reg)),] %>%
          group_by(genus) %>%
          slice(1L) %>%
          left_join(.,egg_reg[sample(nrow(egg_reg)),] %>%
              group_by(genus) %>%
              slice(1L),by="genus")

ggplot(on_egg_reg,aes(x = rate, y = egg_rate, color= group)) + geom_point() +
  scale_color_manual(values = mrk) +
  scale_x_log10(label=scientific_10) +
  scale_y_log10(label=scientific_10) +
  theme(legend.position = "none") +
  xlab("rate of ovariole number diversification") +
  ylab("rate of egg volume diversification") +
  geom_vline(xintercept=1e-04,linetype="dashed")
```

We assessed the rate of egg volume diversification across insects using the same method of assessing rate heterogeneity as described in Section \ref{sec:BAMM} (Fig. \ref{fig:BAMM-egg-plot}). This analysis converged in $20*10^8$ generations (the effective size for the number of shifts and log-likelihood were >200, `r round(egg_es_nshifts,2)` and `r round(egg_es_logl,2)` respectively). We compared the correlation between the rates of trait diversification for ovariole number and egg volume by matching the mean rate predicted for each insect genus (the tips of the phylogeny from the BAMM analyses, Fig. \ref{fig:combine-rates}).

\pagebreak
# References

```{r save, cache=FALSE}
save.image("manuscript_supplement.RData")
```

### 2020-07-07 analysis/combine_body_egg_datasets.R

@@ -1,0 +1,216 @@
+# This script was written by SHC in 2019.
+# Its main purpose is to combine the ovariole number dataset with other published data
+
+# read in insect egg size from Church et al 2019 Scientific Data
+egg_database <- read.delim("analysis/datafiles/egg_database_final_values_July2019.tsv",header=T,stringsAsFactors=F)
+
+# read in body size data for specific groups with ovariole number
+raw_bodysize <- read.delim(file="analysis/datafiles/body_size.tsv",header=TRUE,stringsAsFactors=FALSE)
+matched_body_taxonomy <- read.delim(file="analysis/datafiles/body_matched_taxonomy_Sep2019.tsv",header=T,stringsAsFactors=F) %>% filter(problem != "no_taxonomy")
+
+matched_body <- matched_body_taxonomy %>% select(order,family,subfamily,suborder,superfamily,tribe,updated_genus,updated_species,updated_subspecies,ID,updated_fullname,ncbi_id) %>% 
+		rename(updated_order = order)
+
+body <- raw_bodysize %>% 
+		mutate(median_body = ((maximum_body - minimum_body)/2) + minimum_body) %>% 
+		mutate(body = ifelse( !(is.na(mean_body)), mean_body,
+              ifelse ( !(is.na(median_body)), median_body, single_body))) %>% 
+		mutate(logbody = log10(body)) %>% 
+		left_join(.,matched_body,by="ID") %>%
+		mutate(taxonomy = ifelse((updated_fullname == "" | is.na(updated_fullname)),"original","updated")) %>%
+		mutate(genus = ifelse(taxonomy == "updated",updated_genus,genus),
+			species = ifelse(taxonomy == "updated",updated_species,species),
+			subspecies = ifelse(taxonomy == "updated",updated_subspecies,subspecies),
+			order = ifelse(taxonomy == "updated",updated_order,order)) %>% 
+		mutate(name = ifelse(species != "",paste(genus,species,sep="_"),genus)) %>%
+		mutate(group = plyr::mapvalues(order,groups_orders$order,groups_orders$group,warn_missing=F)) 
+
+# build data frames combining egg size and ovariole number
+egg_name <- egg_database[sample(1:nrow(egg_database)),] %>% # shuffle the dataset
+			filter(species != "") %>% # choose only egg datapoints with full name
[…]

### 2020-07-07 analysis/get_ovariole_number_data.R

@@ -1,0 +1,84 @@
+# This script was written by SHC in 2019.
+# Its main purpose is to read in and format the ovariole number dataset
+library(dplyr)
+library(ggtree)
+library(ape)
+library(ggplot2)
+library(RColorBrewer)
+library(phytools)
+
+# read in data from google doc
+raw_ON <- read.delim(file="analysis/datafiles/ovariole_number.tsv",header=TRUE,stringsAsFactors=FALSE)
+
+# read in list of orders and groups = monophyletic large lineages
+groups_orders <- read.delim(file="analysis/datafiles/groups_orders.txt",header=T,stringsAsFactors=F)
+group_levels = c("Apterygota","Palaeoptera","Polyneoptera","Condylognatha","Psocodea","Amphiesmenoptera","Antliophora","Neuropteroidea","Hymenoptera")
+
+# set colors for figures
+mrk <- c("#9c9c9c","#005487","#7570B3","#679c40","#A6761D","#52bcd4","#E6AB02","#1B9E77","#e37870","#9c9c9c","#2a2a2a")
+names(mrk) <- c(group_levels,"above","below")
+
+# read in TaxReformer updated taxonomy
+matched_taxonomy <- read.delim(file="analysis/datafiles/ovariole_names_matched_June2020.tsv",header=T,stringsAsFactors=F)
+
+# select only relevant taxonomic names
+matched <- matched_taxonomy %>% select(order,family,subfamily,suborder,superfamily,superorder,tribe,subtribe,subsection,updated_genus,updated_species,updated_subspecies,ID,updated_fullname,ncbi_id) %>% 
+	rename(updated_order = order) %>%
+	filter(!(ID %in% (raw_ON %>% filter(problem == "taxonomy") %>% pull(ID)))) # remove the taxonomic information for entries where the information was found to be wrong (e.g. Diplura entries, where the order is incorrect from global names)
+
+# add a name, combining genus and species
+ON <- raw_ON %>% 
+	mutate(median_on = ((maximum_on - minimum_on)/2) + minimum_on) %>% # if a range, calculate median
+	# choose a value per record, with priority given to average, range, then single record
+	mutate(on = ifelse(!(is.na(mean_on)), mean_on,
+	       ifelse (!(is.na(median_on)), median_on, single_on))) %>% 
+	mutate(logon = log10(on),sqrton = sqrt(on)) %>% # transform values
[…]

### 2020-07-07 analysis/pgls_functions.R

@@ -1,0 +1,244 @@
+### These functions are used to execute phylogenetic regressions on traits of interest
+library(nlme)
+
+### This function runs a pgls analysis on a set of traits with a tree
+pgls.trees <- function(data,tree,form){
+  # The tree is pruned to include tips in the dataset
+  pruned <- drop.tip(tree,setdiff(tree$tip.label,rownames(data)))
+  
+  # The dataset is trimmed to include only those in the tree
+  red_dat <- data[!rownames(data) %in% setdiff(rownames(data),pruned$tip.label),]
+
+  # build a Brownian Motion correlation struction
+  lm_str <- corBrownian(phy=pruned)
+
+  # run regression with model
+  outskis <- gls(as.formula(form),correlation = lm_str,data = red_dat,method="ML")
+  
+  #outskis <- lm(form,dat)
+  return(summary(outskis))
+}
+
+### This function builds a dataframe for running a pgls analysis
+### It expects a data frame with rank, trait1, and trait2
+build_pgls <- function(data,tree) {
+  # The tree is pruned to include tips in the dataset
+  pruned <- drop.tip(tree,setdiff(tree$tip.label,data$rank))
+  # The dataset is trimmed to include only those in the tree
+  red_dat <- subset(data,!(rank %in% setdiff(data$rank,pruned$tip.label)))
+
+  # The dataframe for pgls is built        
+  pgls_dat <- data.frame(trait1 = red_dat$trait1, trait2 = red_dat$trait2 ,rank = red_dat$rank)
+
+  # Rows with NA and infinite values are removed
+  is.na(pgls_dat) <- sapply(pgls_dat, is.infinite)
+  pgls_dat <- na.omit(pgls_dat)
+
+  # The dataset is grouped according to the taxonomic rank and a random entry is chosen
+  pgls_dat <- data.frame(pgls_dat[sample(nrow(pgls_dat)),] %>% group_by(rank) %>% slice(1L))
+
+  row.names(pgls_dat) <- pgls_dat$rank
+
+  return(pgls_dat)
+}
+
+### This function builds a dataframe for running a pgls analysis against residuals
+### It expects a data frame with rank, trait1, trait2, and indep (independent variable)
+build_resid_pgls <- function(data,tree) {
+  # The dataset is trimmed to include only those in the tree 
[…]

### 2020-07-07 analysis/reconstruct_ancestral_ovary_morphology_type.R

@@ -1,0 +1,71 @@
+# This script was written by SHC in 2019.
+# Its main purpose is to reconstruct the ancestral mode of oogenesis
+library(OUwie)
+library(corHMM)
+library(parallel)
+library(geiger)
+source("analysis/get_ovariole_number_data")
+egg_database <- read.delim("analysis/datafiles/egg_database_final_values_July17.tsv",header=T,stringsAsFactors=F)
+
+### These commands are used to estimate the evolutionary shifts in ovary type 
+# read in table of taxonomic groups, labeled by ovary type, as recorded in Buning 1994
+ovary_table <- read.delim("analysis/datafiles/mode_of_oogensis_Buning.tsv",header=T,stringsAsFactors=F)
+
+# The Buning tables are listed by genus, but not all genera are in the final dataset
+# However, ovary type is consistent across families within Bunings set
+# So when a genus is not present, for phylogenetic purposes we can choose an alternate genus from the same family
+
+tree_dataset <- egg_database[sample(nrow(egg_database)),] %>% #start with the largest dataset of genera, families, and orders
+	filter(genus %in% tree$tip.label) #use only tips in the tree
+
+# Find one-to-one genus matches
+ovary_table_genera_match <- ovary_table %>% filter(genus %in% tree_dataset$genus)
+# Find families without one-to-one genus match
+ovary_table_family_match <- ovary_table %>% filter(!(family %in% ovary_table_genera_match$family)) %>% 
+	filter(family %in% tree_dataset$family)
+# Find orders without genus or family match
+ovary_table_order_match <- ovary_table %>% 	filter(!(order %in% ovary_table_genera_match$order)) %>% 
+	filter(!(order %in% ovary_table_family_match$order)) %>% 
+	filter(order %in% tree_dataset$order)
+
+# Build a new ovary table with these substitutions
+# First for families
+new_ovary_table_family <- ovary_table_family_match %>% 
+	group_by(family) %>% 
+	slice(1L) %>% 
+	select(-genus) %>% 
+	left_join(tree_dataset %>% 
+		group_by(family) %>% 
+		slice(1L) %>% 
+		select(family,genus),by="family") 
+# Then for orders
+new_ovary_table_order <- ovary_table_order_match %>% 
+	group_by(order) %>% 
+	slice(1L) %>% 
+	select(-genus) %>% 
+	left_join(tree_dataset %>% 
[…]

### 2020-07-07 analysis/run_mode_of_oogenesis_model_comparison.R

@@ -1,0 +1,144 @@
+# This script was written by SHC in 2019.
+# Its main purpose is to compare models of evolution, taking into account the evolution of oogenesis mode
+library(OUwie)
+library(corHMM)
+library(parallel)
+library(geiger)
+source("analysis/get_ovariole_number_data")
+g_trees <- genus_trees[1:100] # select 100 trees from the posterior distribution
+downsample_factor <- 1 # do not downsample the data
+fam_count_threshold <- 1 # use all families
+source("analysis/combine_body_egg_datasets.R")
+
+# read in the table from Buning, with records for ovary type by genus
+ovary_table <- read.delim("analysis/datafiles/mode_of_oogensis_Buning.tsv",header=T,stringsAsFactors=F)
+
+# set number of cores
+cores <- 50
+
+args = commandArgs(trailingOnly=TRUE)
+
+
+# choose the trait of interest, argument 1
+if(args[1] == "egg") { #egg volume
+	analysis_name <- "egg"
+	trait_dataset <- egg_database[sample(nrow(egg_database)),] %>% mutate(trait = logvol) %>%  #set the trait
+		filter(!(is.na(trait))) #remove missing data from the trait dataset
+} else if(args[1] == "ar") { #egg aspect ratio
+	analysis_name <- "ar"
+	trait_dataset <- egg_database[sample(nrow(egg_database)),] %>% mutate(trait = logar) %>% filter(!(is.na(trait)))
+} else if(args[1] == "asym") { #egg asymmetry
+	analysis_name <- "asym"
+	trait_dataset <- egg_database[sample(nrow(egg_database)),] %>% mutate(trait = sqasym) %>% filter(!(is.na(trait))) 
+} else if(args[1] == "curv") { #egg curvature
+	analysis_name <- "curv"
+	trait_dataset <- egg_database[sample(nrow(egg_database)),] %>% mutate(trait = sqcurv) %>% filter(!(is.na(trait))) 
+} else {
+	analysis_name <- "on" #ovariole number
+	trait_dataset <- ON[sample(nrow(ON)),] %>% mutate(trait = logon) %>% filter(!(is.na(trait))) 
+}
+
+# The Buning tables are listed by genus, but not all genera are in the final dataset
+# However, ovary type is consistent across families within Bunings set
+# So when a genus is not present, for phylogenetic purposes we can choose an alternate genus from the same family
+
[…]

### 2020-07-07 analysis/run_pgls_body_ovariole_number.R

@@ -1,0 +1,103 @@
+# This script was written by SHC in 2019.
+# Its main purpose is to run phylogenetic regressions on ovariole number and body size
+library(phytools)
+source("analysis/pgls_functions.R")
+
+# Set factor for resampling the body size dataframe
+downsample_factor <- 0.5
+fam_count_threshold <- 1
+
+body_analysis_names <- c("allometry_body_egg_ON_family",
+	"allometry_gilbert_name",
+	"allometry_gilbert_genus",
+	"allometry_COL1",
+	"allometry_ORT",
+	"allometry_HYM",
+	"allometry_DIP")
+
+# run PGLS over one tree from the posterior distribution
+run_all_allometry_pgls <- function(tree){
+	### build family body size dataset, repeat each iteration to account for phenotypic uncertainty
+	source("analysis/combine_body_egg_datasets.R")
+
+
+	### PGLS body length vs ON
+	allometry_body_egg_ON_family <- run_all_taxa_pgls(body_egg_ON_full %>% 
+							select(rank,logvol,logon,logbody,group) %>% 
+							rename(trait1 = logon, trait2 = logbody),
+								fam_tree)
+
+
+	### Gilbert body weight vs ON, species level
+	allometry_gilbert_name <- run_all_taxa_pgls(egg_ON_gilbert_name %>% 
+							select(genus,logvol,logon,logweight,group) %>% 
+							mutate(rank = genus, trait1 = logon, trait2 = logweight),
+								tree)
+
+	### Gilbert body weight, genus level
+	allometry_gilbert_genus <- run_all_taxa_pgls(egg_ON_gilbert_genus %>% 
+							select(genus,logvol,logon,logweight,group) %>% 
+							mutate(rank = genus, trait1 = logon, trait2 = logweight),
+								tree)
+
+	### Curculionoidea body vs ON
+	allometry_COL1 <- run_all_taxa_pgls(COL1_body_egg_ON_genus %>% 
+							select(genus,logvol,logon,logbody,group) %>% 
+							rename(rank = genus, trait1 = logon, trait2 = logbody),
+								tree)
+
+	### Orthoptera body vs ON
+	allometry_ORT <- run_all_taxa_pgls(ORT_body_egg_ON_genus %>% 
+							select(genus,logvol,logon,logbody,group) %>% 
+							rename(rank = genus, trait1 = logon, trait2 = logbody),
[…]

### 2020-07-07 analysis/run_pgls_over_posterior_distribution.R

@@ -1,0 +1,141 @@
+# This script was written by SHC in 2019.
+# Its main purpose is to run phylogenetic regressions on ovariole number and egg size / number
+library(phytools)
+source("analysis/pgls_functions.R")
+# Set factor for resampling the body size dataframe
+downsample_factor <- 0.5
+fam_count_threshold <- 1
+
+analysis_names <- c("allometry_egg_ON_name",
+	"allometry_egg_ON_genus",
+	"allometry_body_egg_ON_family",
+	"allometry_body_egg_ON_TRIG_family",
+	"allometry_gilbert_name",
+	"allometry_gilbert_genus",
+	"allometry_fecundity_name",
+	"allometry_fecundity_genus",
+	"allometry_COL1",
+	"allometry_ORT",
+	"allometry_HYM",
+	"allometry_DIP")
+
+# run PGLS over one tree from the posterior distribution
+run_all_allometry_pgls <- function(tree){
+	### build family body size dataset
+	source("analysis/combine_body_egg_datasets.R")
+
+	### PGLS on vs egg volume by insect name
+	allometry_egg_ON_name <- run_all_taxa_pgls(egg_ON_name %>% 
+							select(genus,logvol,logon,group) %>% 
+							rename(rank = genus, trait1 = logon, trait2 = logvol),
+								tree)
+
+	### PGLS on vs egg volume by insect genus
+	allometry_egg_ON_genus <- run_all_taxa_pgls(egg_ON_genus %>% 
+							select(genus,logvol,logon,group) %>% 
+							rename(rank = genus, trait1 = logon, trait2 = logvol),
+								tree)
+
+	### PGLS egg vs ON, independent body size
+	allometry_body_egg_ON_family <- run_all_taxa_resid_pgls(body_egg_ON_full %>% 
+							select(rank,logvol,logon,logbody,group) %>% 
+							rename(trait1 = logon, trait2= logvol, indep = logbody),
+								fam_tree)
+
+	### PGLS egg vs ON, ind body size, removing family Trigonalidae
+	allometry_body_egg_ON_TRIG_family <- run_all_taxa_resid_pgls(body_egg_ON_full %>% 
+							filter(!(family == "Trigonalidae")) %>% 
+							select(rank,logvol,logon,logbody,group) %>% 
+							rename(trait1 = logon, trait2= logvol, indep = logbody),
+								fam_tree)
+
[…]

### 2020-07-07 manuscript.Rmd
updating commit number

@@
 
-Samuel H Church^1^\*, Bruno A. S. de Medeiros^1,2^, Seth Donoughe^1,3^, Nicole L. Márquez Reyes^4^, Cassandra G. Extavour^1,5^\*
+Samuel H. Church^1^\*, Bruno A. S. de Medeiros^1,2^, Seth Donoughe^1,3^, Nicole L. Márquez Reyes^4^, Cassandra G. Extavour^1,5^\*
 
@@
 The dataset of insect ovariole number is available in the supplementary file 'ovariole_number_data.tsv' and has been uploaded to Dryad (doi:10.5061/dryad.59zw3r253).
-The code and phylogenetic trees required to reproduce all the analyses, figures, and generate the manuscript files are provided at 'https://github.com/shchurch/insect_ovariole_number_evolution_2020'.
+The code and phylogenetic trees required to reproduce all the analyses, figures, and generate the manuscript files are provided at 'https://github.com/shchurch/insect_ovariole_number_evolution_2020', commit 6cf446a.
 
@@
 Scale bar indicates 100 million years (Myr).
-Gray = panoistic ovaries, without nurse cells, cyan=polytrophic meroistic ovaries, with nurse cells adjacent to maturing oocytes, red = telotrophic meroistic ovaries, with nurse cells located in germaria, black = unique meroistic ovary type observed in Strepsiptera.
+Gray = panoistic ovaries, without nurse cells, cyan = polytrophic meroistic ovaries, with nurse cells adjacent to maturing oocytes, red = telotrophic meroistic ovaries, with nurse cells located in germaria, black = unique meroistic ovary type observed in Strepsiptera.
 Insect taxonomic groups are, from top to bottom: Apterygota, Palaeoptera, Polyneoptera, Condylognatha, Psocodea, Hymenoptera, Neuropteroidea, Amphiesmenoptera, and Antliophora.

### 2020-07-07 manuscript.Rmd
quick correction in manuscript file

@@
 
-Using this approach, we gathered `r nrow(raw_ON)` records for ovariole number from `r raw_ON %>% distinct(publication) %>% nrow()` publications, listed in the supplementary files 'ovariole_number_bibliography.pdf' and 'ovariole_number_bibliography.bib'.
+Using this approach, we gathered `r nrow(raw_ON)` records for ovariole number from `r raw_ON %>% distinct(publication) %>% nrow()` publications, listed in the supplementary file 'ovariole_number_bibliography.pdf'.
 We matched these records to additional taxonomic information using the software TaxReformer[@TaxReformer].
 For all subsequent analyses, we excluded observations made in non-reproductive individuals from eusocial species (workers), as well as two observations that represented significant outliers and could not be validated using additional sources[@su2015testicular;@hernandez2017biology].
-See Supplemental Methods Section \ref{sec:gathering} for details.
+See supplementary methods section \ref{sec:gathering} for details.
 
 We combined the data we collected on total ovariole number with existing datasets of egg size and shape[@church2019dataset], insect lifetime fecundity and dry adult body weight[@Gilbert2007], average adult body length per insect family[@rainford2016phylogenetic], several lineage-specific measures of adult body size[@iwata1966large;@Iwata1966a;@waloff1954number;@starmer2003phylogenetic;@reinhardt2005low], and the mode of oogensis[@buning1994insect].
-See Supplemental Methods Section \ref{sec:combining} for details.
+See supplementary methods section \ref{sec:combining} for details.
 
@@
 Analyses of Drosophilidae ovariole number, egg size, and body size were performed using a phylogeny newly assembled for this study.
-See Supplemental Methods Section \ref{sec:trees} for details.
+See supplementary methods section \ref{sec:trees} for details.
 
@@
 To weight traits by body size, we calculated the phylogenetic residuals[@revell2010phylogenetic] of each trait to body size, and then compared the evolution of these residuals using a PGLS regression.
-See Supplemental Methods Section \ref{sec:PGLS} for details.
+See supplementary methods section \ref{sec:PGLS} for details.
 
@@
 Ancestral state reconstruction and model comparison were repeated 100 times over a posterior distribution of trees and resampling data to account for variation across records for the same taxon.
-See Supplemental Methods Section \ref{sec:oogenesis-model}.
+See supplementary methods section \ref{sec:oogenesis-model}.
 
 Other comparisons of model fit were performed using the R package geiger[@harmon2007geiger], version `r packageVersion('geiger')`, and validated using a parametric bootstrap with the R package arbutus[@pennell2015model], version `r packageVersion('arbutus')`.
-See Supplemental Methods Section \ref{sec:bootstrap}.
+See supplementary methods section \ref{sec:bootstrap}.
 
@@
 Repeated BAMM analyses showed similar distributions of high and low rate regimes, indicating the implications for ovariole number evolution are robust to uncertainty in rate estimates.
-See Supplemental Methods Section \ref{sec:BAMM} for details.
+See supplementary methods section \ref{sec:BAMM} for details.
 
@@
 
-The dataset of insect ovariole number is available in the supplementary file 'ovariole_number_data.tsv' and has been uploaded to Dryad (doi:10.5061/dryad.59zw3r253).
+The dataset of insect ovariole number is available at Dryad (doi:10.5061/dryad.59zw3r253).
 The code and phylogenetic trees required to reproduce all the analyses, figures, and generate the manuscript files are provided at 'https://github.com/shchurch/insect_ovariole_number_evolution_2020', commit 6cf446a.
@@
 Across insects we did not observe a significant relationship between ovariole number and egg size when controlling for body size (p-value `r all_table[6,]$'pval_max'`; Fig.
-2b, \ref{fig:PGLS-Gilbert-name}, \ref{fig:PGLS-egg-ON-family}, Table \ref{tab:PGLS-posterior})), though we do note the presence of a non-significant negative trend, primarily driven by a few extreme values (e.g.
+2b, \ref{fig:PGLS-Gilbert-name}, \ref{fig:PGLS-egg-ON-family}, Table \ref{tab:PGLS-posterior}), though we do note the presence of a non-significant negative trend, primarily driven by a few extreme values (e.g.
 wasps in the family Trigonalidae, which have many ovarioles and lay minute eggs, Fig.

### 2020-12-06 analysis/combine_body_egg_datasets.R
updated analysis files for new p-value interpretation. PGLS now performed over MCC trees, then evaluated over 1000 trees from posterior distribution. Also added simulation analysis. Fixed typos and updated functions to eliminate error messages.

@@ -98,11 +98,11 @@ ON_family <- ON[sample(nrow(ON)),] %>% filter(!(is.na(on))) %>% group_by(family)
 ### fam_count_threshold <- 1
 ### The fam_count_threshold is going to be set by the parent R command, when this file is run from source()
 
-egg_fam_count <- egg_family %>% group_by(family) %>% summarize(n = n())
+egg_fam_count <- egg_family %>% group_by(family) %>% summarize(n = n(),.groups="drop")
 egg_fam_count_n <- egg_fam_count %>% filter(n > fam_count_threshold)
 egg_family <- egg_family %>% filter(family %in% egg_fam_count_n$family)
 
-ON_fam_count <- ON_family %>% group_by(family) %>% summarize(n = n())
+ON_fam_count <- ON_family %>% group_by(family) %>% summarize(n = n(),.groups="drop")
 ON_fam_count_n <- ON_fam_count %>% filter(n > fam_count_threshold)
 ON_family <- ON_family %>% filter(family %in% ON_fam_count_n$family)
 

### 2020-12-06 analysis/pgls_functions.R

@@ -10,7 +10,7 @@ pgls.trees <- function(data,tree,form){
   red_dat <- data[!rownames(data) %in% setdiff(rownames(data),pruned$tip.label),]
 
   # build a Brownian Motion correlation struction
-  lm_str <- corBrownian(phy=pruned)
+  lm_str <- corBrownian(phy=pruned,form=~rank)
 
   # run regression with model
   outskis <- gls(as.formula(form),correlation = lm_str,data = red_dat,method="ML")
@@ -30,8 +30,7 @@ build_pgls <- function(data,tree) {
   # The dataframe for pgls is built        
   pgls_dat <- data.frame(trait1 = red_dat$trait1, trait2 = red_dat$trait2 ,rank = red_dat$rank)
 
-  # Rows with NA and infinite values are removed
-  is.na(pgls_dat) <- sapply(pgls_dat, is.infinite)
+  # Rows with NA values are removed
   pgls_dat <- na.omit(pgls_dat)
 
   # The dataset is grouped according to the taxonomic rank and a random entry is chosen
@@ -51,8 +50,7 @@ build_resid_pgls <- function(data,tree) {
   # The dataframe for the pgls is built        
   pgls_dat <- data.frame(trait1 = red_dat$trait1, trait2 = red_dat$trait2, indep = red_dat$indep, rank=red_dat$rank)
 
-  # Rows with NA and infinite values are removed
-  is.na(pgls_dat) <- sapply(pgls_dat,is.infinite)
+  # Rows with NA values are removed
   pgls_dat <-na.omit(pgls_dat)
 
   # The dataset is grouped according to the taxonomic rank and a random entry is chosen
@@ -82,7 +80,7 @@ build_resid_pgls <- function(data,tree) {
 run_all_taxa_pgls <- function(data,tree){
   # Build the pgls dataset across all taxa
   all_dat <- build_pgls(data,tree)
-  # Runt the pgls across all taxa
+  # Run the pgls across all taxa
   all_pgls_out <- pgls.trees(all_dat,tree,"trait2 ~ trait1")
 
   return(all_pgls_out)
@@ -116,7 +114,7 @@ run_all_taxa_and_by_group_pgls <- function(data,tree,group_list){
 run_all_taxa_resid_pgls <- function(data,tree){
   # Build the pgls residual dataset across all taxa
   all_dat <- build_resid_pgls(data,tree)
-  # Runt the pgls residual across all taxa
[…]

### 2020-12-06 analysis/run_pgls_body_ovariole_number.R

@@ -1,103 +1,0 @@
-# This script was written by SHC in 2019.
-# Its main purpose is to run phylogenetic regressions on ovariole number and body size
-library(phytools)
-source("analysis/pgls_functions.R")
-
-# Set factor for resampling the body size dataframe
-downsample_factor <- 0.5
-fam_count_threshold <- 1
-
-body_analysis_names <- c("allometry_body_egg_ON_family",
-	"allometry_gilbert_name",
-	"allometry_gilbert_genus",
-	"allometry_COL1",
-	"allometry_ORT",
-	"allometry_HYM",
-	"allometry_DIP")
-
-# run PGLS over one tree from the posterior distribution
-run_all_allometry_pgls <- function(tree){
-	### build family body size dataset, repeat each iteration to account for phenotypic uncertainty
-	source("analysis/combine_body_egg_datasets.R")
-
-
-	### PGLS body length vs ON
-	allometry_body_egg_ON_family <- run_all_taxa_pgls(body_egg_ON_full %>% 
-							select(rank,logvol,logon,logbody,group) %>% 
-							rename(trait1 = logon, trait2 = logbody),
-								fam_tree)
-
-
-	### Gilbert body weight vs ON, species level
-	allometry_gilbert_name <- run_all_taxa_pgls(egg_ON_gilbert_name %>% 
-							select(genus,logvol,logon,logweight,group) %>% 
-							mutate(rank = genus, trait1 = logon, trait2 = logweight),
-								tree)
-
-	### Gilbert body weight, genus level
-	allometry_gilbert_genus <- run_all_taxa_pgls(egg_ON_gilbert_genus %>% 
-							select(genus,logvol,logon,logweight,group) %>% 
-							mutate(rank = genus, trait1 = logon, trait2 = logweight),
-								tree)
-
-	### Curculionoidea body vs ON
-	allometry_COL1 <- run_all_taxa_pgls(COL1_body_egg_ON_genus %>% 
-							select(genus,logvol,logon,logbody,group) %>% 
-							rename(rank = genus, trait1 = logon, trait2 = logbody),
-								tree)
-
-	### Orthoptera body vs ON
-	allometry_ORT <- run_all_taxa_pgls(ORT_body_egg_ON_genus %>% 
-							select(genus,logvol,logon,logbody,group) %>% 
-							rename(rank = genus, trait1 = logon, trait2 = logbody),
[…]

### 2020-12-06 analysis/run_pgls_over_posterior_distribution.R

@@ -1,141 +1,0 @@
-# This script was written by SHC in 2019.
-# Its main purpose is to run phylogenetic regressions on ovariole number and egg size / number
-library(phytools)
-source("analysis/pgls_functions.R")
-# Set factor for resampling the body size dataframe
-downsample_factor <- 0.5
-fam_count_threshold <- 1
-
-analysis_names <- c("allometry_egg_ON_name",
-	"allometry_egg_ON_genus",
-	"allometry_body_egg_ON_family",
-	"allometry_body_egg_ON_TRIG_family",
-	"allometry_gilbert_name",
-	"allometry_gilbert_genus",
-	"allometry_fecundity_name",
-	"allometry_fecundity_genus",
-	"allometry_COL1",
-	"allometry_ORT",
-	"allometry_HYM",
-	"allometry_DIP")
-
-# run PGLS over one tree from the posterior distribution
-run_all_allometry_pgls <- function(tree){
-	### build family body size dataset
-	source("analysis/combine_body_egg_datasets.R")
-
-	### PGLS on vs egg volume by insect name
-	allometry_egg_ON_name <- run_all_taxa_pgls(egg_ON_name %>% 
-							select(genus,logvol,logon,group) %>% 
-							rename(rank = genus, trait1 = logon, trait2 = logvol),
-								tree)
-
-	### PGLS on vs egg volume by insect genus
-	allometry_egg_ON_genus <- run_all_taxa_pgls(egg_ON_genus %>% 
-							select(genus,logvol,logon,group) %>% 
-							rename(rank = genus, trait1 = logon, trait2 = logvol),
-								tree)
-
-	### PGLS egg vs ON, independent body size
-	allometry_body_egg_ON_family <- run_all_taxa_resid_pgls(body_egg_ON_full %>% 
-							select(rank,logvol,logon,logbody,group) %>% 
-							rename(trait1 = logon, trait2= logvol, indep = logbody),
-								fam_tree)
-
-	### PGLS egg vs ON, ind body size, removing family Trigonalidae
-	allometry_body_egg_ON_TRIG_family <- run_all_taxa_resid_pgls(body_egg_ON_full %>% 
-							filter(!(family == "Trigonalidae")) %>% 
-							select(rank,logvol,logon,logbody,group) %>% 
-							rename(trait1 = logon, trait2= logvol, indep = logbody),
-								fam_tree)
-
[…]

### 2020-12-06 analysis/run_pgls_simulated.R

@@ -1,0 +1,102 @@
+source("analysis/get_ovariole_number_data.R")
+source("analysis/pgls_functions.R") # functions for building a PGLS dataframe
+library(geiger)
+library(phylolm)
+downsample_factor <- 1;	fam_count_threshold <- 1 # don't downsample family level average data
+
+args = commandArgs(trailingOnly=TRUE) # get arguments, there are 2
+analysis_name <- args[1]
+slope <- as.numeric(args[2])
+
+# if the analysis is done with the family level tree, update the tree objects
+if(analysis_name %in% c("allometry_body_egg_ON_family")) {
+	tree <- fam_tree; genus_trees <- rep(list(tree),1000); class(genus_trees) <- "multiPhylo"
+}
+
+# function for simulating data and running a PGLS
+run_simulated_pgls <- function(tree){
+	source("analysis/combine_body_egg_datasets.R") # combine ON and other traits, this will randomly shuffle and choose rep. genera each iteration
+
+	if(analysis_name == "allometry_body_egg_ON_family") {	
+		residual_dataset <- body_egg_ON_full %>% 
+			mutate(trait1 = logon, 
+					trait2 = logvol, # egg volume
+					indep = logbody) # body length
+		# calculate reisduals for each trait to body size
+		dataset_to_simulate <- build_resid_pgls(residual_dataset,tree) %>% mutate(X1 = trait1, 
+																				X2 = trait2) # use residuals as traits to simulate
+	}
+	if(analysis_name == "allometry_gilbert_genus") {	
+		residual_dataset <- egg_ON_gilbert_genus %>% mutate(rank = genus, 
+															trait1 = logon, 
+															trait2= logvol, # egg volume
+															indep = logweight) # body weight
+		# calculate reisduals for each trait to body size
+		dataset_to_simulate <- build_resid_pgls(residual_dataset,tree) %>% mutate(X1 = trait1,
+																				X2 = trait2) # use residuals as traits to simulate
+	}
+	
+	# trim the phylogeny to only include tips with both traits
+	dataset_tips <- dataset_to_simulate %>% select(rank,X1,X2) %>% na.omit() %>% pull(rank)
+	sim_tree <- drop.tip(tree,tree$tip.label[which(!(tree$tip.label %in% dataset_tips))])
+
[…]

### 2020-12-06 analysis/run_posterior_pgls_body_size.R

@@ -1,0 +1,77 @@
+# This script was written by SHC in 2020.
+# Its main purpose is to run phylogenetic regressions on ovariole number and egg size / number
+library(phytools)
+source("analysis/pgls_functions.R")
+# Set factor for resampling the body size dataframe
+set.seed(12345)
+downsample_factor <- 0.5
+fam_count_threshold <- 1
+
+# run PGLS over one tree from the posterior distribution
+run_all_allometry_pgls <- function(tree){
+	### build family body size dataset
+	source("analysis/combine_body_egg_datasets.R")
+
+	### PGLS body length vs ON
+	allometry_body_ON_family <- run_all_taxa_pgls(body_egg_ON_full %>% 
+							select(rank,logon,logbody,group) %>% 
+							rename(trait1 = logon, trait2 = logbody),
+								fam_tree)
+
+
+	### Gilbert body weight vs ON, species level
+	allometry_body_gilbert_name <- run_all_taxa_pgls(egg_ON_gilbert_name %>% 
+							select(genus,logon,logweight,group) %>% 
+							mutate(rank = genus, trait1 = logon, trait2 = logweight),
+								tree)
+
+	### Gilbert body weight, genus level
+	allometry_body_gilbert_genus <- run_all_taxa_pgls(egg_ON_gilbert_genus %>% 
+							select(genus,logon,logweight,group) %>% 
+							mutate(rank = genus, trait1 = logon, trait2 = logweight),
+								tree)
+
+	### Curculionoidea body vs ON
+	allometry_body_COL1 <- run_all_taxa_pgls(COL1_body_egg_ON_genus %>% 
+							select(genus,logon,logbody,group) %>% 
+							rename(rank = genus, trait1 = logon, trait2 = logbody),
+								tree)
+
+	### Orthoptera body vs ON
+	allometry_body_ORT <- run_all_taxa_pgls(ORT_body_egg_ON_genus %>% 
+							select(genus,logon,logbody,group) %>% 
+							rename(rank = genus, trait1 = logon, trait2 = logbody),
+								tree)
+
+	### Apoidea body vs ON
+	allometry_body_HYM <- run_all_taxa_pgls(HYM_body_egg_ON_genus %>% 
+							select(genus,logon,logbody,group) %>% 
+							rename(rank = genus, trait1 = logon, trait2 = logbody),
+								tree)
+
+	### Drosophilidae body vs ON
[…]

### 2020-12-06 analysis/run_posterior_pgls_egg_size_ovariole_number.R

@@ -1,0 +1,60 @@
+# This script was written by SHC in 2020.
+# Its main purpose is to run phylogenetic regressions on ovariole number and egg size / number
+library(phytools)
+source("analysis/pgls_functions.R")
+# Set factor for resampling the body size dataframe
+set.seed(12345)
+downsample_factor <- 0.5
+fam_count_threshold <- 1
+
+# run PGLS over one tree from the posterior distribution
+run_all_allometry_pgls <- function(tree){
+	### build family body size dataset
+	source("analysis/combine_body_egg_datasets.R")
+
+	### PGLS on vs egg volume by insect name
+	allometry_egg_ON_name <- run_all_taxa_pgls(egg_ON_name %>% 
+							select(genus,logvol,logon,group) %>% 
+							rename(rank = genus, trait1 = logon, trait2 = logvol),
+								tree)
+
+	### PGLS on vs egg volume by insect genus
+	allometry_egg_ON_genus <- run_all_taxa_pgls(egg_ON_genus %>% 
+							select(genus,logvol,logon,group) %>% 
+							rename(rank = genus, trait1 = logon, trait2 = logvol),
+								tree)
+
+	### PGLS egg vs ON, independent body size
+	allometry_body_egg_ON_family <- run_all_taxa_resid_pgls(body_egg_ON_full %>% 
+							select(rank,logvol,logon,logbody,group) %>% 
+							rename(trait1 = logon, trait2= logvol, indep = logbody),
+								fam_tree)
+
+	### Gilbert egg vs ON, ind body weight name pgls
+	allometry_gilbert_name <- run_all_taxa_resid_pgls(egg_ON_gilbert_name %>% 
+							select(genus,logvol,logon,logweight,group) %>% 
+							mutate(rank = genus, trait1 = logon, trait2= logvol, indep = logweight),
+								tree)
+
+	### Gilbert egg vs ON, ind body weight genus pgls
+	allometry_gilbert_genus <- run_all_taxa_resid_pgls(egg_ON_gilbert_genus %>% 
+							select(genus,logvol,logon,logweight,group) %>% 
+							mutate(rank = genus, trait1 = logon, trait2= logvol, indep = logweight),
+								tree)
+
+	return(list(allometry_egg_ON_name,
+		allometry_egg_ON_genus,
+		allometry_body_egg_ON_family,
+		allometry_gilbert_name,
+		allometry_gilbert_genus))
+}
+
[…]

### 2020-12-06 analysis/run_posterior_pgls_fecundity.R

@@ -1,0 +1,36 @@
+# This script was written by SHC in 2020.
+# Its main purpose is to run phylogenetic regressions on ovariole number and egg size / number
+library(phytools)
+source("analysis/pgls_functions.R")
+# Set factor for resampling the body size dataframe
+set.seed(12345)
+downsample_factor <- 0.5
+fam_count_threshold <- 1
+
+# run PGLS over one tree from the posterior distribution
+run_all_allometry_pgls <- function(tree){
+	### build family body size dataset
+	source("analysis/combine_body_egg_datasets.R")
+
+	### Gilbert fecundity vs ON name
+	allometry_fecundity_name <- run_all_taxa_pgls(ON_gilbert_name %>% 
+							select(genus,logfecundity,logon,group) %>% 
+							mutate(rank = genus, trait1 = logon, trait2= logfecundity),
+								tree)
+
+	### Gilbert fecundity vs ON genus
+	allometry_fecundity_genus <- run_all_taxa_pgls(ON_gilbert_genus %>% 
+							select(genus,logfecundity,logon,group) %>% 
+							mutate(rank = genus, trait1 = logon, trait2= logfecundity),
+								tree)
+
+	return(list(allometry_fecundity_name,
+		allometry_fecundity_genus))
+}
+
+allometry_distribution_raw <- lapply(genus_trees,run_all_allometry_pgls)
+
+allometry_fecundity_name <- lapply(allometry_distribution_raw,function(x){x[[1]]})
+allometry_fecundity_genus <- lapply(allometry_distribution_raw,function(x){x[[2]]})
+
+save.image("analysis/run_posterior_pgls_fecundity.RData")

### 2020-12-06 analysis/run_posterior_pgls_subclade.R

@@ -1,0 +1,52 @@
+# This script was written by SHC in 2020.
+# Its main purpose is to run phylogenetic regressions on ovariole number and egg size / number
+library(phytools)
+source("analysis/pgls_functions.R")
+# Set factor for resampling the body size dataframe
+set.seed(12345)
+downsample_factor <- 0.5
+fam_count_threshold <- 1
+
+# run PGLS over one tree from the posterior distribution
+run_all_allometry_pgls <- function(tree){
+	### build family body size dataset
+	source("analysis/combine_body_egg_datasets.R")
+
+	### Curculionoidea ON, egg, indep body pgls
+	allometry_COL1 <- run_all_taxa_resid_pgls(COL1_body_egg_ON_genus %>% 
+							select(genus,logvol,logon,logbody,group) %>% 
+							rename(rank = genus, trait1 = logon, trait2= logvol, indep = logbody),
+								tree)
+
+	### Orthoptera ON, egg, indep body pgls
+	allometry_ORT <- run_all_taxa_resid_pgls(ORT_body_egg_ON_genus %>% 
+							select(genus,logvol,logon,logbody,group) %>% 
+							rename(rank = genus, trait1 = logon, trait2= logvol, indep = logbody),
+								tree)
+
+	### Apoidea ON, egg, indep body pgls
+	allometry_HYM <- run_all_taxa_resid_pgls(HYM_body_egg_ON_genus %>% 
+							select(genus,logvol,logon,logbody,group) %>% 
+							rename(rank = genus, trait1 = logon, trait2= logvol, indep = logbody),
+								tree)
+
+	### Apoidea ON, egg, indep body pgls
+	allometry_DIP <- run_all_taxa_resid_pgls(DIP_body_egg_ON_name %>% 
+							select(name,logvol,logon,logbody,group) %>% 
+							rename(rank = name, trait1 = logon, trait2= logvol, indep = logbody),
+								dros_tree)
+
+	return(list(allometry_COL1,
+		allometry_ORT,
+		allometry_HYM,
+		allometry_DIP))
+}
+
+allometry_distribution_raw <- lapply(genus_trees,run_all_allometry_pgls)
+
+allometry_COL1 <- lapply(allometry_distribution_raw,function(x){x[[1]]})
+allometry_ORT <- lapply(allometry_distribution_raw,function(x){x[[2]]})
+allometry_HYM <- lapply(allometry_distribution_raw,function(x){x[[3]]})
[…]

### 2020-12-07 analysis/run_pgls_simulated.R
fixing random seed in simulation analyses, and updating results files

@@ -4,6 +4,8 @@ library(geiger)
 library(phylolm)
 downsample_factor <- 1;	fam_count_threshold <- 1 # don't downsample family level average data
 
+set.seed(12345)
+
 args = commandArgs(trailingOnly=TRUE) # get arguments, there are 2
 analysis_name <- args[1]
 slope <- as.numeric(args[2])

### 2020-12-08 manuscript.Rmd
updating Rmd files with new method for calculating and reporting p-values

@@
  \usepackage{float}
+ \usepackage{array}
+ \usepackage{xcolor,colortbl}
  \floatplacement{figure}{H}
@@
 Here we perform a test of the hypothesized trade off between the number and size of offspring using thousands of descriptions of the number of egg-producing compartments in the insect ovary (ovarioles), a common proxy for potential offspring number in insects.
-In contrast to prior claims, we find that ovariole number is not generally negatively correlated with the size of insect eggs, and we highlight several factors that may have contributed to this size-number trade off being strongly asserted in previous studies.
+\textcolor{red}{We find evidence of a negative relationship between egg size and ovariole number when accounting for adult body size, but in contrast to prior claims, we note that this relationship is not generalizable across insect clades, and we highlight several factors that may have contributed to this size-number trade off being stated as a general rule in previous studies}.
 We reconstruct the evolutionary history of the nurse cell arrangement within the ovariole, and show that the diversification of ovariole number and egg size have both been largely independent of nurse cell presence or position within the ovariole.
@@
 
+\textcolor{red}{For two regressions comparing egg size to ovariole number while accounting for adult body size, we tested alternative hypotheses of evolution by simulating new data.
+We considered two hypotheses: no evolutionary correlation with ovariole number, and a strong correlation with ovariole number (slope of -1:1).
+For each trait we simulated 1,000 datasets using evolutionary parameters fit under a Brownian Motion model in the R packages geiger}[@revell2010phylogenetic]\textcolor{red}{, version `r packageVersion('geiger')`, and phylolm}[@tung2014linear]\textcolor{red}{, version `r packageVersion('phylolm')`.}
+
 Ancestral state reconstruction of oogenesis mode was performed with the R package corHMM[@beaulieu2013identifying], version `r packageVersion('corHMM')`, and models of trait evolution were compared using the R package Ouwie[@beaulieu2012modeling], version `r packageVersion('OUwie')`.
@@
 
-The dataset of insect ovariole number is available at Dryad (doi:10.5061/dryad.59zw3r253).
+The dataset of insect ovariole number is available at Dryad, doi:10.5061/dryad.59zw3r253.
+
+## Code availability {-}
+
 The code and phylogenetic trees required to reproduce all the analyses, figures, and generate the manuscript files are provided at 'https://github.com/shchurch/insect_ovariole_number_evolution_2020', commit 6cf446a.
@@
 
-All regressions were performed over a posterior distribution of trees and accounting for phenotypic uncertainty; therefore these analyses produced a range of p-values.
-We considered a relationship significant when the maximum p-value observed from this range was below the threshold 0.01.
-We report the maximum p-value throughout this manuscript, and have included the full range in the supplementary information (see Table \ref{tab:PGLS-posterior}).
+\textcolor{red}{All phylogenetic regressions were performed using both the maximum clade credibility (MCC) tree and over 1000 trees drawn from the posterior distribution.
+We considered a relationship significant when the MCC p-value was below the threshold 0.01.
+To assess the robustness of results to uncertainty, we also report the number of regressions over the posterior distribution of phylogenetic trees that gave a significant result (see Table \ref{tab:PGLS-posterior-egg-ON}).}
+
+\textcolor{red}{For two comparisons, we validated that our tests had sufficient statistical power using the selected threshold by comparing the distribution of p-values from regressions of observed data to regressions of data simulated under alternative hypotheses.
+We used the comparison to results based on simulated data to evaluate the likelihood of false positives (comparing to data simulatd under no correlation) and false negatives (comparing to data simulated with strong correlation).}
 
@@
 b, The range of total adult ovariole number, log~10~ scale, across nine groups of insects, arranged with random jitter on the y-axis within each group.
-Groups are, from top to bottom: Apterygota, Palaeoptera, Polyneoptera, Condylognatha, Psocodea, Amphiesmenoptera, Antliophora, Neuropteroidea, and Hymenoptera.](manuscript_figures/Fig_Diversity_V3-01.png)
+Groups are, from top to bottom: Apterygota, Palaeoptera, Polyneoptera, Condylognatha, Psocodea, Amphiesmenoptera, Antliophora, Neuropteroidea, and Hymenoptera.](manuscript_figures/Fig_Diversity_V4-01.png)
 
@@
 This hypothesis is based on the predictions that (1) female reproduction is resource-limited, therefore egg size should trade off with egg number, and (2) ovariole number can serve as a proxy for egg number[@montague1981reproductive;@smith1974optimal].
-Our data do not support the hypothesis of a negative relationship between egg size and ovariole number across insects (p-value `r all_table[2,]$'pval_max'`; Figs.
-2a, \ref{fig:PGLS-egg-ON-genus}, Table \ref{tab:PGLS-posterior}).
+We did not observe a significant negative relationship when comparing egg size and ovariole number across insect species \textcolor{red}{(Figs.
+2a, Table \ref{tab:PGLS-posterior-egg-ON}, p-value `r round(pgls_results_egg_ON_name$tTable[[8]],3)`, n=`r pgls_results_egg_ON_name$dims$N`).
+To test results with a larger sample size, we also compared across egg size and ovariole number across insect genera and did not observe significant relationship (Fig.
+\ref{fig:PGLS-egg-ON-genus}, p-value `r round(pgls_results_egg_ON_genus$tTable[[8]],3)`, n=`r pgls_results_egg_ON_genus$dims$N`).}
 
-Given that this prediction is often conditioned on taking into account adult body size[@stewart1991reproductive;@starmer2003phylogenetic], we combined data on ovariole number and egg size with several measurements of body size, and applied a size correction that accounts for phylogenetic relationships [@revell2009size].
-Across insects we did not observe a significant relationship between ovariole number and egg size when controlling for body size (p-value `r all_table[6,]$'pval_max'`; Fig.
-2b, \ref{fig:PGLS-Gilbert-name}, \ref{fig:PGLS-egg-ON-family}, Table \ref{tab:PGLS-posterior}), though we do note the presence of a non-significant negative trend, primarily driven by a few extreme values (e.g.
-wasps in the family Trigonalidae, which have many ovarioles and lay minute eggs, Fig.
-\ref{fig:PGLS-egg-ON-family}).
+Given that this predicted relationship is often conditioned on taking into account adult body size[@stewart1991reproductive;@starmer2003phylogenetic], we combined data on ovariole number and egg size with several measurements of body size, and applied a size correction that accounts for phylogenetic relationships [@revell2009size].
+\textcolor{red}{When accounting for adult body weight, we observed a significant negative relationship between egg size and ovariole number (Fig.
+2b, \ref{fig:PGLS-Gilbert-name}, p-value `r gsub(0,"<0.001",round(pgls_results_gilbert_genus$tTable[[8]],3))`, n=`r pgls_results_gilbert_genus$dims$N`).
+To evaluate the robustness of this result, we repeated the analysis 1000 times, taking into account uncertainty in both the phylogeny and trait measurements.
+Out of 1000 regressions, `r sum(sapply(allometry_gilbert_genus,function(x){x$tTable[[8]]}) < 0.01)` indicated a significant negative relationship (Table \ref{tab:PGLS-posterior-egg-ON}).
+We performed the same comparison accounting for adult body length, and likewise observed a significant negative relationship (Fig.
+\ref{fig:PGLS-family-length}, p-value `r gsub(0,"<0.001",round(pgls_results_body_egg_ON_family$tTable[[8]],3))`, n=`r pgls_results_body_egg_ON_family$dims$N`), supported by `r sum(sapply(allometry_body_egg_ON_family,function(x){x$tTable[[8]]}) < 0.01)` of 1000 repeated analyses (Table \ref{tab:PGLS-posterior-egg-ON}).}
 
-We assessed the relationship between these traits for four sublineages of insects and found that across Drosophilidae species, egg size is indeed strongly negatively correlated with ovariole number when accounting for body size (p-value <0.001; Fig.
-2c, Table \ref{tab:PGLS-posterior}).
-However this relationship did not hold for any of the three other lineages we analyzed (e.g.
-Orthoptera p-value `r all_table[9,]$'pval_max'`; Fig.
-2d, \ref{fig:PGLS-Coleoptera-Hymenoptera}, Table \ref{tab:PGLS-posterior}).
+\textcolor{red}{We further explored these results using two methods: First, to evaluate our findings against alternative evolutionary hypotheses, we compared these results to regressions based on simulated data.
+Our results showed that when considering body size, the slope of the regression of egg size and ovariole number is smaller than we would observe by chance, compared to data simulated with no evolutionary correlation (Fig.
+\ref{fig:simulated}).
+When considering adult body length, but not weight, the slope of the regressions on observed data are within the range that would be expected under an strong negative relationship (-1:1 slope in log-log space, Fig.
+\ref{fig:simulated}).
+This suggests that, for at least some measures of body size, there is a strong negative relationship bewteeen egg size and ovariole number when accounting for body size.}
 
-We additionally tested whether ovariole number is positively correlated with adult body size, and in contrast to previous studies[@honvek1993intraspecific], we found no correlation between ovariole number and adult body weight or length across insects (p-values `r body_all_table[3,]$'pval_max'` and `r body_all_table[1,]$'pval_max'`, respectively; Fig.
-\ref{fig:PGLS-body-plot}, Table \ref{tab:PGLS-body-posterior}).
+\textcolor{red}{Second, we assessed the relationship between egg size and ovariole number, accounting for body size, within four subclades of insects.
+We found that across Drosophilidae fly species, egg size is indeed strongly negatively correlated with ovariole number when accounting for body size (Fig.
+2c, Table \ref{tab:PGLS-posterior-clade}, p-value `r gsub(0,"<0.001",round(pgls_results_DIP$tTable[[8]],3))`, n=`r pgls_results_DIP$dims$N`).
+For beetles (Coleoptera), we observed a significant relationship (\ref{fig:PGLS-Coleoptera-Hymenoptera}, p-value `r round(pgls_results_COL1$tTable[[8]],3)`, n=`r pgls_results_COL1$dims$N`), but a substantial fraction of repeat analyses show that these results are not robust to uncertainty (`r sum(sapply(allometry_COL1,function(x){x$tTable[[8]]}) > 0.01)` of 1000 regressions are not significant, Table \ref{tab:PGLS-posterior-clade}).
+For grasshoppers and crickets (Orthoptera) and wasps (Hymenoptera), we observed no significant relationship between ovariole number and egg size, accounting for body size (Fig.
+2d, Table \ref{tab:PGLS-posterior-clade}, \ref{fig:PGLS-Coleoptera-Hymenoptera}, Orthoptera: p-value `r round(pgls_results_ORT$tTable[[8]],3)`, n=`r pgls_results_ORT$dims$N`, Hymenoptera: p-value `r round(pgls_results_HYM$tTable[[8]],3)`, n=`r round(pgls_results_HYM$dims$N)`).
+This indicates that, while a strong negative correlation between egg size and ovariole number exists for some insects, it does not represent a universal pattern across insect clades.}
 
-![**Tests of the hypothesized trade off between egg size and ovariole number.** a, Egg volume (mm^3^) and ovariole number, both log~10~ scale, n=`r nrow(egg_ON_name_pgls)` species.
-See section _Modeling ovariole number evolution_ for discussion of the enrichment of certain low values of ovariole number (points appearing vertically arranged) b, Egg volume and ovariole number, residuals to adult dry body weight, n=`r length(which(egg_ON_gilbert_genus_pgls$rank %in% tree$tip.label))` genera.
-c, Drosophilidae egg volume and ovariole number, residuals to thorax length, n=`r length(which(DIP_pgls$rank %in% dros_tree$tip.label))` species.
-d, Orthoptera egg volume and ovariole number, residuals to body length, n=`r length(which(ORT_pgls$rank %in% tree$tip.label))` genera.](manuscript_figures/Fig_Egg_V2-01.png)
+Finally, we tested whether ovariole number is positively correlated with adult body size, and in contrast to previous studies[@honvek1993intraspecific], we found no correlation between ovariole number and adult body weight or length across insects \textcolor{red}{(Fig.
+\ref{fig:PGLS-body-plot}, Table \ref{tab:PGLS-posterior-body}, body weight: p-value `r round(pgls_results_body_gilbert_name$tTable[[8]],3)`, n=`r pgls_results_body_gilbert_genus$dims$N`, body length: p-value `r round(pgls_results_body_ON_family$tTable[[8]],3)`, n=`r pgls_results_body_ON_family$dims$N`).
+Of the four subclades considered, only insects in the family Drosophilidae and the order Orthoptera indicated a positive relationship between body size and ovariole number (Table \ref{tab:PGLS-posterior-body}, p-value `r gsub(0,"<0.001",round(pgls_results_body_ORT$tTable[[8]],3))`, n=`r pgls_results_body_ORT$dims$N`).}
+
+![**Tests of the hypothesized trade off between egg size and ovariole number.** a, Egg volume (mm^3^) and ovariole number, both log~10~ scale, \textcolor{red}{points represent insect species}.
+See section _Modeling ovariole number evolution_ for discussion of the enrichment of certain low values of ovariole number (points appearing vertically arranged) b, Egg volume and ovariole number, residuals to adult dry body weight, \textcolor{red}{points=genera}.
+c, Drosophilidae egg volume and ovariole number, residuals to thorax length, \textcolor{red}{points=species}.
+d, Orthoptera egg volume and ovariole number, residuals to body length, \textcolor{red}{points=genera}.](manuscript_figures/Fig_Egg_V2-01.png)
 
@@
 
-Using a previously reported dataset of lifetime fecundity measurements across insects[@Gilbert2007], we assessed the relationship between lifetime fecundity and ovariole number, and observed a positive, but not statistically significant, correlation (p-value `r all_table[8,]$'pval_max'`; Figs.
-3, \ref{fig:PGLS-fecundity-name}).
-We interpret our results, in conjunction with those previously reported, to suggest that ovariole number, when variable across insects in a lineage, may be one factor among many influencing the number of eggs produced.
+Using a previously reported dataset of lifetime fecundity measurements across insects[@Gilbert2007], we assessed the relationship between lifetime fecundity and ovariole number, and \textcolor{red}{observed a significant positive relationship (Fig.
+3, p-value `r round(pgls_results_fecundity_name$tTable[[8]],3)`, n=`r pgls_results_fecundity_name$dims$N`).
+However, we note that this relationship is largely defined by the absence of insects with high ovariole number and low fecundity (Fig.
+3, empty bottom right corner).
+For insects with low ovariole number, fecundity varied over more than three orders of magnitude.} We interpret our results, in conjunction with those previously reported, to suggest that ovariole number, when variable across insects in a lineage, may be one factor among many influencing the number of eggs produced.
 However, we caution against using ovariole number as a direct mathematical proxy for offspring number.
@@
 ![**The relationship between lifetime fecundity and ovariole number.** Both values are shown on a log~10~ scale.
-n=`r length(na.omit(fecundity_genus_pgls$fecundity))` genera, colored according to the groups shown in Fig.
+Points \textcolor{red}{represent insect genera} and are colored according to the groups shown in Fig.
 1b.](manuscript_figures/Fig_Fecundity_V1-01.png)
@@
 In insects, egg size and ovariole number are often used as proxies for offspring fitness[@koch2014mother] and number[@david1970nombre;@bouletreau1982reproductive], respectively, and therefore it has been predicted that insects with more ovarioles lay smaller eggs than insects with fewer ovarioles[@hodin2009she;@montague1981reproductive;@berrigan1991allometry;@starmer2003phylogenetic].
-Our results, using a dataset that spans `r nrow(raw_ON)` observations and that takes into account phylogenetic relationships, indicate that a general trade-off between insect egg size and ovariole number does not exist (Fig.
+Our results, using a dataset that spans `r nrow(raw_ON)` observations and that takes into account phylogenetic relationships, indicate that a generalized trade-off between insect egg size and ovariole number does not exist (Fig.
 2).

### 2020-12-08 manuscript_supplement.Rmd
updating commit number
quick correction in manuscript file
updating Rmd files with new method for calculating and reporting p-values

@@
     latex_engine: pdflatex
-    toc: no
     #bookdown::html_document2: default
+toc: true
 bibliography: manuscript_supplement.bib
@@
 ```{r global options, include = FALSE}
-knitr::opts_chunk$set(echo=FALSE, include = FALSE, warning=FALSE, message=FALSE, cache=TRUE)
-```
-
-```{r preliminaries,cache=FALSE}
+knitr::opts_chunk$set(echo=FALSE, include = FALSE, warning=FALSE, message=FALSE, cache=FALSE)
+```
+
+```{r preliminaries}
 library(BAMMtools)
@@
 
-```{r load_data, cache=FALSE}
+```{r load_data}
 # This script reads in and formats the ovariole number dataset
@@
 
-Samuel H Church^1^\*, Bruno A. S. De Medeiros^1,2^, Seth Donoughe^1,3^, Nicole L. Márquez Reyes^4^, Cassandra G. Extavour^1,5^\*
+Samuel H. Church^1^\*, Bruno A. S. De Medeiros^1,2^, Seth Donoughe^1,3^, Nicole L. Márquez Reyes^4^, Cassandra G. Extavour^1,5^\*
 
@@
 Using this approach, we gathered `r nrow(raw_ON)` records for ovariole number from 460 publications.
-A full list of publications is provided in the supplementary files 'ovariole_number_bibliography.pdf'.
+A full list of publications is provided in the supplementary file 'ovariole_number_bibliography.pdf'.
 We matched the scientific names to additional taxonomic information using the software TaxReformer[@TaxReformer] and found additional taxonomic data for `r nrow(matched)` of the `r nrow(raw_ON)` records.
@@
 Average adult body length per insect family[@rainford2016phylogenetic] was matched to the average ovariole number for the corresponding families (Fig.
-\ref{fig:PGLS-egg-ON-family}).
+\ref{fig:PGLS-family-length}).
 The Rainford et al, 2016[@rainford2016phylogenetic] dataset contains a small number of average adult body lengths at the order level (e.g.
@@
 
-```{r PGLS-posterior, include=TRUE, cache=FALSE}
-#source("analysis/run_pgls_over_posterior_distribution.R")
-load("analysis/pgls_results_June2020/run_pgls_over_posterior_distribution.RData")
-
-# analysis names in table 
-pgls_analysis_names <- c("ovariole number vs egg volume",
-  "ovariole number vs egg volume",
-  "ovariole number vs egg volume, residuals to body length",
-  "ovariole number vs egg volume, residuals to body length, excluding Trigonalidae",
-  "ovariole number vs egg volume, residuals to body weight",
-  "ovariole number vs egg volume, residuals to body weight",
-  "ovariole number vs lifetime fecundity",
-  "ovariole number vs lifetime fecundity",
-  "Curculionoidea ovariole number vs egg volume, residual to elytra length",
-  "Orthoptera ovariole number vs egg volume, residuals to body length",
-  "Hymenoptera ovariole number vs egg volume, residuals to mesosoma width",
-  "Drosophilidae ovariole number vs egg volume, residuals to thorax length")
-# level at which each PGLS was performed 
-taxonomic_levels <- c("species","genus","family","family","species","genus","species","genus","genus","genus","genus","species")
-
-sp <- " -- " # separator in tables
-
-# write table
-all_table <- as.data.frame(do.call(rbind, allometry_tables)) %>% round(.,3) %>% mutate(analysis = pgls_analysis_names, "taxonomic level" = taxonomic_levels,slope = paste(sprintf("%.3f",slope_min),sprintf("%.3f",slope_max),sep=sp),
-      intercept = paste(sprintf("%.3f",int_min),sprintf("%.3f",int_max),sep=sp),
-      "p-value" = paste(sprintf("%.3f",pval_min),gsub("0.000","<0.000",sprintf("%.3f",pval_max)),sep=sp),
-      "taxa" = taxa) 
-knitr::kable(all_table  %>% select(c("analysis","taxonomic level","slope","intercept","p-value","taxa")), "latex",caption="Results of PGLS analysis of ovariole number, egg size, and fecundity across a posterior distribution", booktabs = T, linesep = "") %>%
-  kableExtra::kable_styling(latex_options = "striped") %>%
-  kableExtra::column_spec(1, width = "16em") %>%
-  kableExtra::kable_styling(latex_options = "hold_position")  %>%
-  kableExtra::kable_styling(font_size = 8) 
-```
-
-```{r PGLS-body-posterior, include=TRUE, cache=FALSE}
-#source("analysis/run_pgls_body_ovariole_number.R")
-load("analysis/pgls_results_June2020/run_pgls_body_ovariole_number.RData")
-
-# analysis names in table 
-pgls_analysis_names <- c("ovariole number vs body length",
-  "ovariole number vs body weight",
-  "ovariole number vs body weight",
-  "Curculionoidea ovariole number vs elytra length",
-  "Orthoptera ovariole number vs body length",
-  "Hymenoptera ovariole number vs mesosoma width",
-  "Drosophilidae ovariole number vs thorax length")
-# level at which each PGLS was performed 
-taxonomic_levels <- c("family","species","genus","genus","genus","genus","species")
-
-# write table 
-body_all_table <- as.data.frame(do.call(rbind, body_allometry_tables)) %>% round(.,3) %>% mutate(analysis = pgls_analysis_names, "taxonomic level" = taxonomic_levels,slope = paste(sprintf("%.3f",slope_min),sprintf("%.3f",slope_max),sep=sp),
-       intercept = paste(sprintf("%.3f",int_min),sprintf("%.3f",int_max),sep=sp),
-       "p-value" = paste(sprintf("%.3f",pval_min),gsub("0.000","<0.000",sprintf("%.3f",pval_max)),sep=sp),
-       "taxa" = taxa)
-knitr::kable(body_all_table %>% select(c("analysis","taxonomic level","slope","intercept","p-value","taxa")), "latex",caption="Results of PGLS analysis of ovariole number and body size across a posterior distribution", booktabs = T, linesep = "") %>%
-  kableExtra::kable_styling(latex_options = "striped") %>%
-  kableExtra::column_spec(1, width = "16em") %>%
-  kableExtra::kable_styling(latex_options = "hold_position")  %>%
-  kableExtra::kable_styling(font_size = 8) 
-```
-
-For each regression, we repeated a Phylogenetic Generalized Least Squares (PGLS) analysis 1000 times, accounting for phylogenetic and phenotypic uncertainty, using the R packages ape (version `r packageVersion('ape')`)[@paradis2004ape] and nlme (version `r packageVersion('nlme')`)[@pinheiro2014r].
-In these analyses we used a Brownian Motion based covariance matrix for traits.
-
-For regressions at the species and genus level, we reshuffled and matched records at each iteration to account for variation across records for the same taxon.
-For regressions at the family level we recalculated the average ovariole number per insect family, downsampling the representation for each family by half.
-No posterior distribution was made available with the published family level phylogeny[@rainford2014phylogenetic].
-
-To account for body size, we calculated the phylogenetic residuals[@revell2010phylogenetic] of each trait to body size, and then compared the evolution of these residuals using a PGLS regression.
-
 ```{r PGLS-setup}
-# build body size dataset
-# set factor for resampling the body size dataframe, here no downsampling used
-downsample_factor <- 1.0
+source("analysis/pgls_functions.R")
+
+# combine datasets, using all data and randomly shuffled database entries
+downsample_factor <- 1
 fam_count_threshold <- 1
+
+set.seed(12345)  # set seed for cobmination
 source("analysis/combine_body_egg_datasets.R")
-source("analysis/pgls_functions.R") # get functions for plotting PGLS results
-```
-
-```{r PGLS-egg-ON-name}
+
+# use the maximum clade credibility tree
+tree <- genus_mcc_tree
+```
+
+```{r PGLS-egg-ON}
 ### PGLS egg volume vs ON, species level pgls
-egg_ON_name_pgls <- egg_ON_name %>% select(genus,vol,on,group) %>% rename(rank = genus, trait1 = on, trait2 = vol)
+pgls_results_egg_ON_name <- run_all_taxa_pgls(egg_ON_name %>% 
+    select(genus,logvol,logon,group) %>% 
+    rename(rank = genus, trait1 = logon, trait2 = logvol) ,
+    tree)
+
+egg_ON_name_plot <- egg_ON_name %>% 
+    select(genus,vol,on,group) %>% 
+    rename(rank = genus, trait1 = on, trait2 = vol)
 
@@
 pdf("manuscript_figures/panels/FigEgg_ON_name.pdf",height=5,width=7.5,useDingbats = F)
-  g1 <- ggplot(egg_ON_name_pgls,aes(x = trait1, y = trait2, color = group)) +
+g1 <- ggplot(egg_ON_name_plot,aes(x = trait1, y = trait2, color = group)) +
     geom_point() +
@@
 invisible(dev.off())
-
-### Gilbert egg volume vs ON, independent body weight, genus level pgls
-egg_ON_gilbert_genus_pgls <- egg_ON_gilbert_genus %>% select(genus,logvol,logon,logweight,group) %>% mutate(rank = genus, trait1 = logon, trait2= logvol, indep = logweight)
-pdf("manuscript_figures/panels/FigEgg_gilbert_weight_residuals.pdf",height=2.5,width=2.5,useDingbats = F)
-run_plotting_resid_pgls(egg_ON_gilbert_genus_pgls,tree,"gilbert_genus") + xlab("residual of ovariole number") +
-  ylab("residual of egg volume")
-invisible(dev.off())
-
-### Drosophilidae pgls
-DIP_pgls <- DIP_body_egg_ON_name %>% select(name,logvol,logon,logbody,group) %>% rename(rank = name, trait1 = logon, trait2= logvol, indep = logbody)
-pdf("manuscript_figures/panels/FigEgg_Drosophila_residuals.pdf",height=2.5,width=2.5,useDingbats = F)
-run_plotting_resid_pgls(DIP_pgls,dros_tree,"DIP") + xlab("residual of ovariole number") +
-  ylab("residual of egg volume")
-invisible(dev.off())
-
-### Orthoptera pgls
-ORT_pgls <- ORT_body_egg_ON_genus %>% select(genus,logvol,logon,logbody,group) %>% rename(rank = genus, trait1 = logon, trait2= logvol, indep = logbody)
-pdf("manuscript_figures/panels/FigEgg_Orthoptera_residuals.pdf",height=2.5,width=2.5,useDingbats = F)
-run_plotting_resid_pgls(ORT_pgls,tree,"ORT") + xlab("residual of ovariole number") +
-  ylab("residual of egg volume")
-invisible(dev.off())
-```
-
-\clearpage
+```
 
 ```{r PGLS-egg-ON-genus, fig.cap="**Egg volume vs ovariole number, matching records at the genus level.**", fig.width=7.5, fig.height=5, include=TRUE}
-### PGLS egg volume vs ON, species level pgls
-egg_ON_genus_pgls <- egg_ON_genus %>% select(genus,logvol,logon,group) %>% rename(rank = genus, trait1 = logon, trait2 = logvol)
+### PGLS on vs egg volume by insect genus
+egg_ON_genus_pgls <- egg_ON_genus %>% 
+    select(genus,logvol,logon,group) %>% 
+    rename(rank = genus, trait1 = logon, trait2 = logvol)
+
+pgls_results_egg_ON_genus <- run_all_taxa_pgls(egg_ON_genus_pgls,tree)
+
 run_plotting_pgls(egg_ON_genus_pgls,tree,"egg_ON_genus") +
@@
 
+```{r PGLS-egg-ON-weight-genus}
+### Gilbert egg vs ON, ind body weight genus pgls
+gilbert_genus_pgls <- egg_ON_gilbert_genus %>% 
+    select(genus,logvol,logon,logweight,group) %>% 
+    mutate(rank = genus, trait1 = logon, trait2= logvol, indep = logweight)
+
+pgls_results_gilbert_genus <- run_all_taxa_resid_pgls(gilbert_genus_pgls,tree)
+
+### Gilbert egg volume vs ON, independent body weight, genus level pgls
+pdf("manuscript_figures/panels/FigEgg_gilbert_weight_residuals.pdf",height=2.5,width=2.5,useDingbats = F)
+run_plotting_resid_pgls(gilbert_genus_pgls,tree,"gilbert_genus") + 
+    xlab("residual of ovariole number") + 
+    ylab("residual of egg volume")
+invisible(dev.off())
+```
+
 ```{r PGLS-Gilbert-name, fig.cap="**Egg volume vs ovariole number, phylogenetic residuals to dry adult body weight, matching records at the species level.**", fig.width=3, fig.height=3, include=TRUE}
+### Gilbert egg vs ON, ind body weight name pgls
+egg_ON_gilbert_name_pgls <- egg_ON_gilbert_name %>% 
+    select(genus,logvol,logon,logweight,group) %>% 
+    mutate(rank = genus, trait1 = logon, trait2= logvol, indep = logweight)
+
+pgls_results_gilbert_name <- run_all_taxa_resid_pgls(egg_ON_gilbert_name_pgls,tree)
+
 ### Gilbert body weight, species level pgls
-egg_ON_gilbert_name_pgls <- egg_ON_gilbert_name %>% select(genus,logvol,logon,logweight,group) %>% mutate(rank = genus, trait1 = logon, trait2= logvol, indep = logweight)
-run_plotting_resid_pgls(egg_ON_gilbert_name_pgls,tree,"gilbert_name") + xlab("residual of ovariole number") +
+run_plotting_resid_pgls(egg_ON_gilbert_name_pgls,tree,"gilbert_name") + 
+    xlab("residual of ovariole number") + 
     ylab("residual of egg volume")
@@
 
-```{r PGLS-egg-ON-family, fig.cap="**Family average egg volume vs ovariole number, phylogenetic residuals to adult body length.** Point outlined in black = Hymenoptera: Trigonalidae, an outlier that drives the relationship.",  fig.width=3, fig.height=3, include=TRUE}
-### PGLS egg volume vs ON, family averages
-body_egg_ON_family_pgls <- body_egg_ON_full %>% select(rank,logvol,logon,logbody,group) %>% rename(trait1 = logon, trait2= logvol, indep = logbody)
-egfam1 <- run_plotting_resid_pgls(body_egg_ON_family_pgls,fam_tree,"body_egg_ON_family")
-
-egfam1 + xlab("residual of average ovariole number") +
-  ylab("residual of average egg volume") +
-  geom_point(data = (egfam1$data %>% filter(rank == "Trigonalidae")),aes(fill=group),color="black",stroke=1,pch=21)
+```{r PGLS-family-length, fig.cap="**Family average egg volume vs ovariole number, phylogenetic residuals to adult body length.**",  fig.width=3, fig.height=3, include=TRUE}
+### PGLS egg vs ON, independent body size
+body_egg_ON_family_pgls <- body_egg_ON_full %>% 
+    select(rank,logvol,logon,logbody,group) %>% 
+    rename(trait1 = logon, trait2= logvol, indep = logbody)
+
+pgls_results_body_egg_ON_family <- run_all_taxa_resid_pgls(body_egg_ON_family_pgls,fam_tree)
+
+run_plotting_resid_pgls(body_egg_ON_family_pgls,fam_tree,"body_egg_ON_family") + 
+    xlab("residual of average ovariole number") +
+    ylab("residual of average egg volume")
+```
+
+\textcolor{red}{We ran each Phylogenetic Generalized Least Squares regression over the Maximum Clade Credibilty (MCC) tree.
+For each regression, we also repeated each PGLS analysis 1000 times,} accounting for phylogenetic and phenotypic uncertainty, using the R packages ape (version `r packageVersion('ape')`)[@paradis2004ape] and nlme (version `r packageVersion('nlme')`)[@pinheiro2014r].
+In these analyses we used a Brownian Motion based covariance matrix for traits.
+
+For regressions at the species and genus level, we reshuffled and matched records at each iteration to account for variation across records for the same taxon.
+For regressions at the family level we recalculated the average ovariole number per insect family, downsampling the representation for each family by half.
+No posterior distribution was available with the previously published family level phylogeny[@rainford2014phylogenetic].
+
+To account for body size, we calculated the phylogenetic residuals[@revell2010phylogenetic] of each trait to body size, and then compared the evolution of these residuals using a PGLS regression.
+
+\textcolor{red}{For regressions of egg size and ovariole number when accounting for adult body size, we compared the results of our regression analyses to distributions estimated using simulated data under alternative hypotheses.
+We fit a Brownian motion model to the phylogenetic residuals of egg size and body size (R package geiger}[@harmon2007geiger]\textcolor{red}{, version `r packageVersion('geiger')`), and then used the parameters of this fitted model to simulate new datasets (R package phylolm}[@tung2014linear]\textcolor{red}{, version `r packageVersion('phylolm')`).
+We performed this resimulation using the datasets of egg size and body length at the family level, and egg size and body weight at the genus level.
+We simulated 1000 datasets each under two hypotheses: no correlation (slope=0) and a strong negative correlation (slope=-1:1).
+We performed the regressions as described above and compared the distribution of p-values and slopes to values from regressions on observed data.}
+
+```{r simulated, fig.cap="**\\textcolor{red}{Using simulated data to test alternative hypotheses of evolutionary relationships}** Top row, distributions of p-values over 1000 replicate regressions, dashed black line indicates threshold of 0.01.
+Bottom row, distribution of estimated slopes between egg size and ovariole number, dashed lines indicate 95% interval of simulated distributions.
+Left, comparing egg size and ovariole number, accounting for body length at the family level.
+Right, comparing egg size and ovariole number, accounting for body weight at the genus level.
+Red=observed values, gray=simulated with no correlation, black=simulated with a -1:1 correlation.
+n=1000 regressions.", fig.width=7.5, fig.height=5, include=TRUE}
+get_tstat <- function(sum,n) {
+    tstat <- as.numeric(unname(sum[[n]]$coefficients[[2]]))
+    return(tstat)
+}
+
+get_pval <- function(sum,n) {
+    pval <- as.numeric(unname(sum[[n]]$tTable[[8]]))
+    return(pval)
+}
+
+load("analysis/pgls_results_Fall2020/run_pgls_simulated_allometry_body_egg_ON_family_-1_image.RData")
+sim_corr_tstat_dist <- sapply(allometry_distribution_raw,get_tstat,n=1)
+sim_corr_pval_dist <- sapply(allometry_distribution_raw,get_pval,n=1)
+load("analysis/pgls_results_Fall2020/run_pgls_simulated_allometry_body_egg_ON_family_0_image.RData")
+sim_nocorr_tstat_dist <- sapply(allometry_distribution_raw,get_tstat,n=1)
+sim_nocorr_pval_dist <- sapply(allometry_distribution_raw,get_pval,n=1)
+load("analysis/pgls_results_Fall2020/run_posterior_pgls_egg_size_ovariole_number.RData")
+obs_tstat_dist <- sapply(allometry_distribution_raw,get_tstat,n=3)
+obs_pval_dist <- sapply(allometry_distribution_raw,get_pval,n=3)
+
+body_egg_ON_family_tstat_dist <- rbind(tibble(type = "simulated",
+    relationship = "-1:1 correlation",
+    tstat=sim_corr_tstat_dist),
+tibble(type = "simulated",
+    relationship = "no correlation",
+    tstat=sim_nocorr_tstat_dist),
+tibble(type = "observed",
+    relationship = "observed",
+    tstat=obs_tstat_dist))
+
+body_egg_ON_family_pval_dist <- rbind(tibble(type = "simulated",
+    relationship = "-1:1 correlation",
+    pval=sim_corr_pval_dist),
+tibble(type = "simulated",
+    relationship = "no correlation",
+    pval=sim_nocorr_pval_dist),
+tibble(type = "observed",
+    relationship = "observed",
+    pval=obs_pval_dist))
+
+load("analysis/pgls_results_Fall2020/run_pgls_simulated_allometry_gilbert_genus_-1_image.RData")
+sim_corr_tstat_dist <- sapply(allometry_distribution_raw,get_tstat,n=1)
+sim_corr_pval_dist <- sapply(allometry_distribution_raw,get_pval,n=1)
+load("analysis/pgls_results_Fall2020/run_pgls_simulated_allometry_gilbert_genus_0_image.RData")
+sim_nocorr_tstat_dist <- sapply(allometry_distribution_raw,get_tstat,n=1)
+sim_nocorr_pval_dist <- sapply(allometry_distribution_raw,get_pval,n=1)
+load("analysis/pgls_results_Fall2020/run_posterior_pgls_egg_size_ovariole_number.RData")
+obs_tstat_dist <- sapply(allometry_distribution_raw,get_tstat,n=5)
+obs_pval_dist <- sapply(allometry_distribution_raw,get_pval,n=5)
+
+gilbert_genus_pval_dist <- rbind(tibble(type = "simulated",
+    relationship = "-1:1 correlation",
+    pval=sim_corr_pval_dist),
+tibble(type = "simulated",
+    relationship = "no correlation",
+    pval=sim_nocorr_pval_dist),
+tibble(type = "observed",
+    relationship = "observed",
+    pval=obs_pval_dist))
+
+gilbert_genus_tstat_dist <- rbind(tibble(type = "simulated",
+    relationship = "-1:1 correlation",
+    tstat=sim_corr_tstat_dist),
+tibble(type = "simulated",
+    relationship = "no correlation",
+    tstat=sim_nocorr_tstat_dist),
+tibble(type = "observed",
+    relationship = "observed",
+    tstat=obs_tstat_dist))
+
+t1 <- ggplot(body_egg_ON_family_tstat_dist,aes(x=tstat,y=relationship,color=relationship)) + geom_jitter(size=0.2)  + 
+geom_vline(xintercept=quantile(body_egg_ON_family_tstat_dist %>% filter(relationship == "-1:1 correlation") %>% pull(tstat),0.05),colour="black",linetype="longdash") + 
+geom_vline(xintercept=quantile(body_egg_ON_family_tstat_dist %>% filter(relationship == "-1:1 correlation") %>% pull(tstat),0.95),colour="black",linetype="longdash") + 
+geom_vline(xintercept=quantile(body_egg_ON_family_tstat_dist %>% filter(relationship == "no correlation") %>% pull(tstat),0.05),colour="dark gray",linetype="longdash") + 
+geom_vline(xintercept=quantile(body_egg_ON_family_tstat_dist %>% filter(relationship == "no correlation") %>% pull(tstat),0.95),colour="dark gray",linetype="longdash") + 
+#ggtitle("size vs number, body length, family")  + 
+xlab("test statistic, slope") +
+#ylab("type") + 
+xlim(c(-2,2)) +
+scale_color_manual(values = c("black","dark gray","red")) +
+theme(legend.position="none",
+    axis.title.y=element_blank())
+
+t2 <- ggplot(gilbert_genus_tstat_dist,aes(x=tstat,y=relationship,color=relationship)) + geom_jitter(size=0.2)  + 
+geom_vline(xintercept=quantile(gilbert_genus_tstat_dist %>% filter(relationship == "-1:1 correlation") %>% pull(tstat),0.05),colour="black",linetype="longdash") + 
+geom_vline(xintercept=quantile(gilbert_genus_tstat_dist %>% filter(relationship == "-1:1 correlation") %>% pull(tstat),0.95),colour="black",linetype="longdash") + 
+geom_vline(xintercept=quantile(gilbert_genus_tstat_dist %>% filter(relationship == "no correlation") %>% pull(tstat),0.05),colour="dark gray",linetype="longdash") + 
+geom_vline(xintercept=quantile(gilbert_genus_tstat_dist %>% filter(relationship == "no correlation") %>% pull(tstat),0.95),colour="dark gray",linetype="longdash") + 
+#ggtitle("size vs number, body weight, genus") + 
+xlab("test statistic, slope") +
+#ylab("type") + 
+xlim(c(-2,2)) +
+scale_color_manual(values = c("black","dark gray","red")) +
+theme(legend.position="none",
+    axis.title.y=element_blank(),
+axis.text.y=element_blank(),
+axis.ticks.y=element_blank())
+
+p1 <- ggplot(body_egg_ON_family_pval_dist,aes(x=pval,y=relationship,color=relationship)) + geom_jitter(size=0.2)  + 
+geom_vline(xintercept=0.01,colour="black",linetype="dashed") + 
+#geom_vline(xintercept=0.05,colour="black",linetype="longdash") + 
+ggtitle("egg size vs ov. number, body length")  + 
+xlab("p-value") +
+#ylab("type") + 
+xlim(c(0,1)) +
+scale_color_manual(values = c("black","dark gray","red")) +
+theme(legend.position="none",
+    axis.title.y=element_blank())
+
+p2 <- ggplot(gilbert_genus_pval_dist,aes(x=pval,y=relationship,color=relationship)) + geom_jitter(size=0.2)  + 
+geom_vline(xintercept=0.01,colour="black",linetype="dashed") + 
+#geom_vline(xintercept=0.05,colour="black",linetype="longdash") + 
+ggtitle("egg size vs ov. number, body weight") + 
+xlab("p-value") +
+#ylab("type") + 
+xlim(c(0,1)) +
+scale_color_manual(values = c("black","dark gray","red")) +
+theme(legend.position="none",
+    axis.title.y=element_blank(),
+axis.text.y=element_blank(),
+axis.ticks.y=element_blank())
+
+grid.arrange(p1,p2,t1,t2,ncol=2,widths=5:4)
+```
+
+```{r PGLS-sublineage}
+### Apoidea ON, egg, indep body pgls
+DIP_body_egg_ON_name_pgls <- DIP_body_egg_ON_name %>% 
+    select(name,logvol,logon,logbody,group) %>% 
+    rename(rank = name, trait1 = logon, trait2= logvol, indep = logbody)
+
+pgls_results_DIP <- run_all_taxa_resid_pgls(DIP_body_egg_ON_name_pgls,dros_tree)
+
+pdf("manuscript_figures/panels/FigEgg_Drosophila_residuals.pdf",height=2.5,width=2.5,useDingbats = F)
+run_plotting_resid_pgls(DIP_body_egg_ON_name_pgls,dros_tree,"DIP") + 
+    xlab("residual of ovariole number") +
+    ylab("residual of egg volume")
+invisible(dev.off())
+
+### Orthoptera ON, egg, indep body pgls
+ORT_body_egg_ON_genus_pgls <- ORT_body_egg_ON_genus %>% 
+    select(genus,logvol,logon,logbody,group) %>% 
+    rename(rank = genus, trait1 = logon, trait2= logvol, indep = logbody)
+
+pgls_results_ORT <- run_all_taxa_resid_pgls(ORT_body_egg_ON_genus_pgls,tree)
+
+pdf("manuscript_figures/panels/FigEgg_Orthoptera_residuals.pdf",height=2.5,width=2.5,useDingbats = F)
+run_plotting_resid_pgls(ORT_body_egg_ON_genus_pgls,tree,"ORT") + 
+    xlab("residual of ovariole number") +
+    ylab("residual of egg volume")
+invisible(dev.off())
 ```
@@
 ```{r PGLS-Coleoptera-Hymenoptera, fig.cap="**Additional lineage-specific comparisons of egg volume vs ovariole number, phylogenetic residuals to body size, matching records at the genus level.** Weevils (Curculionoidea, left) were measured using elytra length and wasps (Hymenoptera, right) were measured using mesosoma width.", fig.width=6, fig.height=3, include=TRUE}
-### Curculionoidea pgls
-COL1_pgls <- COL1_body_egg_ON_genus %>% select(genus,logvol,logon,logbody,group) %>% rename(rank = genus, trait1 = logon, trait2 = logvol, indep = logbody)
+### Curculionoidea ON, egg, indep body pgls
+COL1_body_egg_ON_genus_pgls <- COL1_body_egg_ON_genus %>% 
+    select(genus,logvol,logon,logbody,group) %>% 
+    rename(rank = genus, trait1 = logon, trait2 = logvol, indep = logbody)
+
+pgls_results_COL1 <- run_all_taxa_resid_pgls(COL1_body_egg_ON_genus_pgls,tree)
+
 theme_update(plot.title = element_text(hjust = 0.5))
-col1 <- run_plotting_resid_pgls(COL1_pgls,tree,"COL1") + xlab("residual of ovariole number") +
-  ylab("residual of egg volume") + ggtitle("Coleoptera: Curculionoidea")
-
-### Apoidea pgls
-HYM_pgls <- HYM_body_egg_ON_genus %>% select(genus,logvol,logon,logbody,group) %>% rename(rank = genus, trait1 = logon, trait2 = logvol, indep = logbody)
-hym1 <- run_plotting_resid_pgls(HYM_pgls,tree,"HYM") + xlab("residual of ovariole number") +
+col1 <- run_plotting_resid_pgls(COL1_body_egg_ON_genus_pgls,tree,"COL1") + 
+    xlab("residual of ovariole number") +
+    ylab("residual of egg volume") + 
+    ggtitle("Coleoptera: Curculionoidea")
+
+### Apoidea ON, egg, indep body pgls
+HYM_body_egg_ON_genus_pgls <- HYM_body_egg_ON_genus %>% 
+    select(genus,logvol,logon,logbody,group) %>% 
+    rename(rank = genus, trait1 = logon, trait2 = logvol, indep = logbody)
+
+pgls_results_HYM <- run_all_taxa_resid_pgls(HYM_body_egg_ON_genus_pgls,tree)
+
+hym1 <- run_plotting_resid_pgls(HYM_body_egg_ON_genus_pgls,tree,"HYM") + xlab("residual of ovariole number") +
 ylab(element_blank()) + ggtitle("Hymenoptera")
@@
 ```{r PGLS-body-plot, fig.cap="**Adult body size vs ovariole number.** Adult body weight compared at the genus level (left), and average adult body length compared at the family level (right).", fig.width=6, fig.height=3, include=TRUE}
-### Gilbert body weight vs ON pgls
-ON_body_gilbert_genus_pgls <- egg_ON_gilbert_genus %>% select(genus,logon,logweight,group) %>% mutate(rank = genus, trait1 = logon, trait2 = logweight)
+### body length vs ON, family level
+ON_body_family_pgls <- body_egg_ON_full %>% 
+    select(rank,logon,logbody,group) %>% 
+    rename(trait1 = logon, trait2 = logbody)
+
+pgls_results_body_ON_family <- run_all_taxa_pgls(ON_body_family_pgls,fam_tree)
+
+family_body <- run_plotting_pgls(ON_body_family_pgls,fam_tree,"body_egg_ON_family") + 
+    xlab(bquote('average ovariole number,'~log[10])) +
+    ylab(bquote('average adult body length (mm),'~log[10]))
+
+### Gilbert body weight vs ON, species level
+pgls_results_body_gilbert_name <- run_all_taxa_pgls(egg_ON_gilbert_name %>% 
+    select(genus,logon,logweight,group) %>% 
+    mutate(rank = genus, trait1 = logon, trait2 = logweight) ,
+    tree)
+
+### Gilbert body weight, genus level
+ON_body_gilbert_genus_pgls <- egg_ON_gilbert_genus %>% 
+    select(genus,logon,logweight,group) %>% 
+    mutate(rank = genus, trait1 = logon, trait2 = logweight)
+
+pgls_results_body_gilbert_genus <- run_all_taxa_pgls(ON_body_gilbert_genus_pgls,tree)
+
 gilbert_body <- run_plotting_pgls(ON_body_gilbert_genus_pgls,tree,"body_gilbert") + xlab(bquote('ovariole number,'~log[10])) +
@@
 
-### PGLS body length vs ON, family averages
-ON_body_family_pgls <- body_egg_ON_full %>% select(rank,logon,logbody,group) %>% rename(trait1 = logon, trait2 = logbody)
-family_body <- run_plotting_pgls(ON_body_family_pgls,fam_tree,"body_egg_ON_family") + xlab(bquote('average ovariole number,'~log[10])) +
-  ylab(bquote('average adult body length (mm),'~log[10]))
+### Curculionoidea body vs ON
+pgls_results_body_COL1 <- run_all_taxa_pgls(COL1_body_egg_ON_genus %>% 
+    select(genus,logon,logbody,group) %>% 
+    rename(rank = genus, trait1 = logon, trait2 = logbody) ,
+    tree)
+
+### Orthoptera body vs ON
+pgls_results_body_ORT <- run_all_taxa_pgls(ORT_body_egg_ON_genus %>% 
+    select(genus,logon,logbody,group) %>% 
+    rename(rank = genus, trait1 = logon, trait2 = logbody) ,
+    tree)
+
+### Apoidea body vs ON
+pgls_results_body_HYM <- run_all_taxa_pgls(HYM_body_egg_ON_genus %>% 
+    select(genus,logon,logbody,group) %>% 
+    rename(rank = genus, trait1 = logon, trait2 = logbody) ,
+    tree)
+
+### Drosophilidae body vs ON
+pgls_results_body_DIP <- run_all_taxa_pgls (DIP_body_egg_ON_name %>% 
+    select(name,logon,logbody,group) %>% 
+    rename(rank = name, trait1 = logon, trait2 = logbody) ,
+    dros_tree)
 
@@
 
+```{r PGLS-fecundity-name, fig.cap="**Lifetime fecundity vs ovariole number, matching records at the species level.**", fig.width=3, fig.height=3, include=TRUE}
+### Gilbert fecundity vs ON name
+fecundity_name_pgls <-  ON_gilbert_name %>% 
+    select(genus,logfecundity,logon,group) %>% 
+    mutate(rank = genus, trait1 = logon, trait2= logfecundity)
+
+pgls_results_fecundity_name <- run_all_taxa_pgls(fecundity_name_pgls,tree)
+
+run_plotting_pgls(fecundity_name_pgls,tree,"fecundity_name") + 
+    xlab(bquote("ovariole number,"~log[10])) +
+    ylab(bquote("lifetime fecundity,"~log[10]))
+```
+
 ```{r PGLS-fecundity-genus}
-### Gilbert fecundity vs ON, genus level
-fecundity_genus_pgls <-  ON_gilbert_genus %>% select(genus,fecundity,on,group) %>% mutate(rank = genus, trait1 = on, trait2= fecundity) 
+### Gilbert fecundity vs ON genus
+fecundity_genus_pgls <-  ON_gilbert_genus %>% 
+    select(genus,fecundity,on,group) %>% 
+    mutate(rank = genus, trait1 = on, trait2= fecundity)
+
+pgls_results_fecundity_genus <- run_all_taxa_pgls(fecundity_genus_pgls,tree)
+
 pdf("manuscript_figures/panels/FigFecundity.pdf",height=4,width=4,useDingbats = F)
@@
 
-```{r PGLS-fecundity-name, fig.cap="**Lifetime fecundity vs ovariole number, matching records at the species level.**", fig.width=3, fig.height=3, include=TRUE}
-### Gilbert fecundity vs ON, species level
-fecundity_name_pgls <-  ON_gilbert_name %>% select(genus,logfecundity,logon,group) %>% mutate(rank = genus, trait1 = logon, trait2= logfecundity)
-run_plotting_pgls(fecundity_name_pgls,tree,"fecundity_name") + xlab(bquote("ovariole number,"~log[10])) +
-  ylab(bquote("lifetime fecundity,"~log[10]))
-```
+```{r PGLS-posterior-egg-ON, include=TRUE}
+# retrieve results from over the posterior distribution
+get_pgls_table <- function(dist) {
+    table <- round(data.frame(
+    slope_min = min(sapply(dist,function(x) {x$coefficients[[2]]})),
+    slope_max = max(sapply(dist,function(x) {x$coefficients[[2]]})),
+    pval_sig = sum(sapply(dist,function(x) {x$tTable[[8]] < 0.01})),
+    taxa = min(sapply(dist,function(x) {x$dims$N}))),
+    4)
+    return(table)
+}
+
+sp <- " -- " # separator in tables
+
+### PGLS TABLE EGG SIZE OVARIOLE NUMBER
+load("analysis/pgls_results_Fall2020/run_posterior_pgls_egg_size_ovariole_number.RData")
+
+egg_ON_posterior_results <- list(allometry_egg_ON_name,
+    allometry_egg_ON_genus,
+    allometry_gilbert_name,
+    allometry_gilbert_genus,
+    allometry_body_egg_ON_family)
+
+egg_ON_pgls_results <- list(pgls_results_egg_ON_name,
+    pgls_results_egg_ON_genus,
+    pgls_results_gilbert_name,
+    pgls_results_gilbert_genus,
+    pgls_results_body_egg_ON_family)
+
+egg_ON_pgls_analysis_names <- c("ovariole number vs egg volume",
+    "ovariole number vs egg volume",
+    "ovariole number vs egg volume, residuals to body weight",
+    "ovariole number vs egg volume, residuals to body weight",
+    "ovariole number vs egg volume, residuals to body length")
+
+egg_ON_taxonomic_levels <- c("species","genus","species","genus","family")
+
+egg_ON_table <- as.data.frame(bind_rows(lapply(egg_ON_posterior_results,get_pgls_table))) %>% mutate(analysis = egg_ON_pgls_analysis_names,
+        "taxonomic level" = egg_ON_taxonomic_levels,
+        "slope" = paste(sprintf("%.3f",slope_min),sprintf("%.3f",slope_max),sep=sp),
+        "MCC p-value" = gsub("0.000","<0.001",sprintf("%.3f",sapply(egg_ON_pgls_results,function(x){x$tTable[[8]]}))),
+        "num. sig. / 1000" = pval_sig,
+        "taxa" = taxa)
+knitr::kable(egg_ON_table %>% select(c("analysis","taxonomic level","slope","MCC p-value","num. sig. / 1000","taxa")),
+        "latex",
+        caption="Results of PGLS analysis of ovariole number and egg size across a posterior distribution", booktabs = T, linesep = "") %>% 
+    kableExtra::kable_styling(latex_options = "striped") %>%
+    kableExtra::column_spec(1, width = "16em") %>%
+    kableExtra::kable_styling(latex_options = "hold_position")  %>%
+    kableExtra::kable_styling(font_size = 8) 
+```
+
+```{r PGLS-posterior-clade, include=TRUE}
+### PGLS TABLE SUBCLADE
+load("analysis/pgls_results_Fall2020/run_posterior_pgls_subclade.RData")
+
+clade_posterior_results <- list(allometry_DIP,
+    allometry_ORT,
+    allometry_COL1,
+    allometry_HYM)
+
+clade_pgls_results <- list(pgls_results_DIP,
+    pgls_results_ORT,
+    pgls_results_COL1,
+    pgls_results_HYM)
+
+clade_pgls_analysis_names <- c("Drosophilidae ovariole number vs egg volume, residuals to thorax length",
+    "Orthoptera ovariole number vs egg volume, residuals to body length",
+    "Curculionoidea ovariole number vs egg volume, residual to elytra length",
+    "Hymenoptera ovariole number vs egg volume, residuals to mesosoma width")
+
+clade_taxonomic_levels <- c("species","genus","genus","genus")
+
+clade_table <- as.data.frame(bind_rows(lapply(clade_posterior_results,get_pgls_table))) %>% mutate(analysis = clade_pgls_analysis_names,
+        "taxonomic level" = clade_taxonomic_levels,
+        "slope" = paste(sprintf("%.3f",slope_min),sprintf("%.3f",slope_max),sep=sp),
+        "MCC p-value" = gsub("0.000","<0.001",sprintf("%.3f",sapply(clade_pgls_results,function(x){x$tTable[[8]]}))),
+        "num. sig. / 1000" = pval_sig,
+        "taxa" = taxa)
+knitr::kable(clade_table %>% select(c("analysis","taxonomic level","slope","MCC p-value","num. sig. / 1000","taxa")),
+        "latex",
+        caption="Results of PGLS analysis of ovariole number and egg size across a posterior distribution", booktabs = T, linesep = "") %>% 
+    kableExtra::kable_styling(latex_options = "striped") %>%
+    kableExtra::column_spec(1, width = "16em") %>%
+    kableExtra::kable_styling(latex_options = "hold_position")  %>%
+    kableExtra::kable_styling(font_size = 8) 
+```
+
+```{r PGLS-posterior-body, include=TRUE}
+### PGLS TABLE BODY SIZE
+load("analysis/pgls_results_Fall2020/run_posterior_pgls_body_size.RData")
+
+body_posterior_results <- list(allometry_body_gilbert_name,
+    allometry_body_gilbert_genus,
+    allometry_body_ON_family,
+    allometry_body_DIP,
+    allometry_body_ORT,
+    allometry_body_COL1,
+    allometry_body_HYM)
+
+body_pgls_results <- list(pgls_results_body_gilbert_name,
+    pgls_results_body_gilbert_genus,
+    pgls_results_body_ON_family,
+    pgls_results_body_DIP,
+    pgls_results_body_ORT,
+    pgls_results_body_COL1,
+    pgls_results_body_HYM)
+
+body_pgls_analysis_names <- c("ovariole number vs body length",
+    "ovariole number vs body weight",
+    "ovariole number vs body weight",
+    "Drosophilidae ovariole number vs thorax length",
+    "Orthoptera ovariole number vs body length",
+    "Curculionoidea ovariole number vs elytra length",
+    "Hymenoptera ovariole number vs mesosoma width")
+
+body_taxonomic_levels <- c("species","genus","family","species","genus","genus","genus")
+
+body_table <- as.data.frame(bind_rows(lapply(body_posterior_results,get_pgls_table))) %>% mutate(analysis = body_pgls_analysis_names,
+        "taxonomic level" = body_taxonomic_levels,
+        "slope" = paste(sprintf("%.3f",slope_min),sprintf("%.3f",slope_max),sep=sp),
+        "MCC p-value" = gsub("0.000","<0.001",sprintf("%.3f",sapply(body_pgls_results,function(x){x$tTable[[8]]}))),
+        "num. sig. / 1000" = pval_sig,
+        "taxa" = taxa)
+knitr::kable(body_table %>% select(c("analysis","taxonomic level","slope","MCC p-value","num. sig. / 1000","taxa")),
+        "latex",
+        caption="Results of PGLS analysis of ovariole number and body size across a posterior distribution", booktabs = T, linesep = "") %>% 
+    kableExtra::kable_styling(latex_options = "striped") %>%
+    kableExtra::column_spec(1, width = "16em") %>%
+    kableExtra::kable_styling(latex_options = "hold_position")  %>%
+    kableExtra::kable_styling(font_size = 8) 
+```
+
+```{r PGLS-posterior-fecundity, include=TRUE}
+### PGLS TABLE FECUNDITY
+load("analysis/pgls_results_Fall2020/run_posterior_pgls_fecundity.RData")
+
+fecundity_posterior_results <- list(allometry_fecundity_name,
+    allometry_fecundity_genus)
+
+fecundity_pgls_results <- list(pgls_results_fecundity_name,
+    pgls_results_fecundity_genus)
+
+fecundity_pgls_analysis_names <- c("ovariole number vs lifetime fecundity",
+    "ovariole number vs lifetime fecundity")
+
+fecundity_taxonomic_levels <- c("species","genus")
+
+fecundity_table <- as.data.frame(bind_rows(lapply(fecundity_posterior_results,get_pgls_table))) %>% mutate(analysis = fecundity_pgls_analysis_names,
+        "taxonomic level" = fecundity_taxonomic_levels,
+        "slope" = paste(sprintf("%.3f",slope_min),sprintf("%.3f",slope_max),sep=sp),
+        "MCC p-value" = gsub("0.000","<0.001",sprintf("%.3f",sapply(fecundity_pgls_results,function(x){x$tTable[[8]]}))),
+        "num. sig. / 1000" = pval_sig,
+        "taxa" = taxa)
+knitr::kable(fecundity_table %>% select(c("analysis","taxonomic level","slope","MCC p-value","num. sig. / 1000","taxa")),
+        "latex",
+        caption="Results of PGLS analysis of ovariole number and fecundity across a posterior distribution", booktabs = T, linesep = "") %>% 
+    kableExtra::kable_styling(latex_options = "striped") %>%
+    kableExtra::column_spec(1, width = "16em") %>%
+    kableExtra::kable_styling(latex_options = "hold_position")  %>%
+    kableExtra::kable_styling(font_size = 8)
+
+```
+
+\clearpage
 
@@
 
-```{r save, cache=FALSE}
+```{r save}
 save.image("manuscript_supplement.RData")

### 2021-01-18 manuscript.Rmd
adding conda environment; rerunning analyses; updating manuscript and README

--- 2020-07-07 manuscript.Rmd
@@
 Here we perform a test of the hypothesized trade off between the number and size of offspring using thousands of descriptions of the number of egg-producing compartments in the insect ovary (ovarioles), a common proxy for potential offspring number in insects.
-In contrast to prior claims, we find that ovariole number is not generally negatively correlated with the size of insect eggs, and we highlight several factors that may have contributed to this size-number trade off being strongly asserted in previous studies.
+We find evidence of a negative relationship between egg size and ovariole number when accounting for adult body size.
+However in contrast to prior claims, we note that this relationship is not generalizable across all insect clades, and we highlight several factors that may have contributed to this size-number trade off being stated as a general rule in previous studies.
 We reconstruct the evolutionary history of the nurse cell arrangement within the ovariole, and show that the diversification of ovariole number and egg size have both been largely independent of nurse cell presence or position within the ovariole.
@@
 
-We combined the data we collected on total ovariole number with existing datasets of egg size and shape[@church2019dataset], insect lifetime fecundity and dry adult body weight[@Gilbert2007], average adult body length per insect family[@rainford2016phylogenetic], several lineage-specific measures of adult body size[@iwata1966large;@Iwata1966a;@waloff1954number;@starmer2003phylogenetic;@reinhardt2005low], and the mode of oogensis[@buning1994insect].
+We combined the data we collected on total ovariole number with existing datasets of egg size and shape[@church2019dataset], insect lifetime fecundity and dry adult body mass[@Gilbert2007], average adult body length per insect family[@rainford2016phylogenetic], several lineage-specific measures of adult body size[@iwata1966large;@Iwata1966a;@waloff1954number;@starmer2003phylogenetic;@reinhardt2005low], and the mode of oogenesis[@buning1994insect].
 See supplementary methods section \ref{sec:combining} for details.
@@
 
-Ancestral state reconstruction of oogenesis mode was performed with the R package corHMM[@beaulieu2013identifying], version `r packageVersion('corHMM')`, and models of trait evolution were compared using the R package Ouwie[@beaulieu2012modeling], version `r packageVersion('OUwie')`.
+For two regressions comparing egg size to ovariole number while accounting for adult body size, we tested alternative hypotheses of evolution by simulating new data.
+We considered two such hypotheses: no evolutionary correlation with ovariole number, and a strong correlation with ovariole number (slope of -1).
+For each trait we simulated 1,000 datasets using evolutionary parameters fit under a Brownian Motion model in the R packages geiger (version `r packageVersion('geiger')`)[@revell2010phylogenetic], and phylolm (version `r packageVersion('phylolm')`)[@tung2014linear].
+
+Ancestral state reconstruction of oogenesis mode was performed with the R package corHMM (version 1.22)[@beaulieu2013identifying], and models of trait evolution were compared using the R package Ouwie (version 1.57)[@beaulieu2012modeling].
 Ancestral state reconstruction and model comparison were repeated 100 times over a posterior distribution of trees and resampling data to account for variation across records for the same taxon.
@@
 
-Other comparisons of model fit were performed using the R package geiger[@harmon2007geiger], version `r packageVersion('geiger')`, and validated using a parametric bootstrap with the R package arbutus[@pennell2015model], version `r packageVersion('arbutus')`.
+Other comparisons of model fit were performed using the R package geiger(version `r packageVersion('geiger')`)[@harmon2007geiger] and validated using a parametric bootstrap with the R package arbutus (version `r packageVersion('arbutus')`)[@pennell2015model].
 See supplementary methods section \ref{sec:bootstrap}.
 
-Analyses of evolutionary rate were performed using BAMM[@rabosky2014automatic], version 2.5.0.
+Analyses of evolutionary rate were performed using BAMM (version 2.5.0)[@rabosky2014automatic].
 For this analysis, we calculated the average ovariole number (log~10~ transformed) for each genus present in the phylogeny (507 taxa).
@@
 \ref{fig:BAMM-analysis-convergence}) and numerically.
-Running BAMM for the maximum possible number of generations and selecting the optimum burn-in\ref{fig:BAMM-analysis-burnin} resulted in an effective size for the number of shifts of `r round(es_nshifts,2)`, and for log-likelihood of `r round(es_logl,2)`.
+Running BAMM for the maximum possible number of generations and selecting the optimum burn-in (Fig.
+\ref{fig:BAMM-analysis-burnin}) resulted in an effective size for the number of shifts of `r round(es_nshifts,2)`, and for log-likelihood of `r round(es_logl,2)`.
 Repeated BAMM analyses showed similar distributions of high and low rate regimes, indicating the implications for ovariole number evolution are robust to uncertainty in rate estimates.
@@
 
-The dataset of insect ovariole number is available at Dryad (doi:10.5061/dryad.59zw3r253).
+The dataset of insect ovariole number is available at Dryad, doi:10.5061/dryad.59zw3r253.
+
+## Code availability {-}
+
 The code and phylogenetic trees required to reproduce all the analyses, figures, and generate the manuscript files are provided at 'https://github.com/shchurch/insect_ovariole_number_evolution_2020', commit 6cf446a.
+All analyses performed in R (version 4.0.3) were done so in a clean environment, built with conda (version 4.9.2), and instructions for rebuilding this environment are provided in the same repository.
 
@@
 
-All regressions were performed over a posterior distribution of trees and accounting for phenotypic uncertainty; therefore these analyses produced a range of p-values.
-We considered a relationship significant when the maximum p-value observed from this range was below the threshold 0.01.
-We report the maximum p-value throughout this manuscript, and have included the full range in the supplementary information (see Table \ref{tab:PGLS-posterior}).
+All phylogenetic regressions were performed using both the maximum clade credibility (MCC) tree.
+We considered a relationship significant when the MCC p-value was below the threshold 0.01.
+To assess the robustness of results to uncertainty in phylogenetic relationships, we also report the number of regressions over the posterior distribution of phylogenetic trees that gave a significant result (see Table \ref{tab:PGLS-posterior-egg-ON}).
+
+For two comparisons, we validated that our tests had sufficient statistical power using the selected threshold by comparing the distribution of p-values from regressions of observed data to regressions of data simulated under alternative hypotheses.
+We compared the results of analyses of our observed to those based on simulated data to evaluate the likelihood of false positives (comparing to data simulatd under no correlation) and false negatives (comparing to data simulated with strong correlation).
 
@@
 This hypothesis is based on the predictions that (1) female reproduction is resource-limited, therefore egg size should trade off with egg number, and (2) ovariole number can serve as a proxy for egg number[@montague1981reproductive;@smith1974optimal].
-Our data do not support the hypothesis of a negative relationship between egg size and ovariole number across insects (p-value `r all_table[2,]$'pval_max'`; Figs.
-2a, \ref{fig:PGLS-egg-ON-genus}, Table \ref{tab:PGLS-posterior}).
+We did not observe a significant negative relationship when comparing egg size and ovariole number across insect species (Fig.
+2a, Table \ref{tab:PGLS-posterior-egg-ON}, p-value `r round(pgls_results_egg_ON_name$tTable[[8]],3)`, n=`r pgls_results_egg_ON_name$dims$N`).
+We also compared egg size and ovariole number across insect genera and did not observe a significant relationship (Fig.
+\ref{fig:PGLS-egg-ON-genus}, p-value `r round(pgls_results_egg_ON_genus$tTable[[8]],3)`, n=`r pgls_results_egg_ON_genus$dims$N`).
 
-Given that this prediction is often conditioned on taking into account adult body size[@stewart1991reproductive;@starmer2003phylogenetic], we combined data on ovariole number and egg size with several measurements of body size, and applied a size correction that accounts for phylogenetic relationships [@revell2009size].
-Across insects we did not observe a significant relationship between ovariole number and egg size when controlling for body size (p-value `r all_table[6,]$'pval_max'`; Fig.
-2b, \ref{fig:PGLS-Gilbert-name}, \ref{fig:PGLS-egg-ON-family}, Table \ref{tab:PGLS-posterior}), though we do note the presence of a non-significant negative trend, primarily driven by a few extreme values (e.g.
-wasps in the family Trigonalidae, which have many ovarioles and lay minute eggs, Fig.
-\ref{fig:PGLS-egg-ON-family}).
+Given that this predicted relationship is often conditioned on taking into account adult body size[@stewart1991reproductive;@starmer2003phylogenetic], we combined data on ovariole number and egg size with several measurements of body size, and applied a size correction that accounts for phylogenetic relationships[@revell2009size].
+When accounting for adult body mass, we observed a significant negative relationship between egg size and ovariole number (Fig.
+2b, \ref{fig:PGLS-Gilbert-name}, p-value `r round(pgls_results_gilbert_genus$tTable[[8]],3)`, n=`r pgls_results_gilbert_genus$dims$N`).
+To evaluate the robustness of this result, we repeated the analysis 1000 times, taking into account uncertainty in both the phylogeny and trait measurements.
+Out of 1000 regressions, `r sum(sapply(allometry_gilbert_genus,function(x){x$tTable[[8]]}) < 0.01)` indicated a significant negative relationship (Table \ref{tab:PGLS-posterior-egg-ON}).
+We performed the same comparison accounting for adult body length, and likewise observed a significant negative relationship (Fig.
+\ref{fig:PGLS-family-length}, p-value `r gsub(0,"<0.001",round(pgls_results_body_egg_ON_family$tTable[[8]],3))`, n=`r pgls_results_body_egg_ON_family$dims$N`), supported by `r sum(sapply(allometry_body_egg_ON_family,function(x){x$tTable[[8]]}) < 0.01)` of 1000 repeated analyses (Table \ref{tab:PGLS-posterior-egg-ON}).
 
-We assessed the relationship between these traits for four sublineages of insects and found that across Drosophilidae species, egg size is indeed strongly negatively correlated with ovariole number when accounting for body size (p-value <0.001; Fig.
-2c, Table \ref{tab:PGLS-posterior}).
-However this relationship did not hold for any of the three other lineages we analyzed (e.g.
-Orthoptera p-value `r all_table[9,]$'pval_max'`; Fig.
-2d, \ref{fig:PGLS-Coleoptera-Hymenoptera}, Table \ref{tab:PGLS-posterior}).
+We further explored these results using two methods: First, to evaluate our findings against alternative evolutionary hypotheses, we compared these results to regressions based on simulated data.
+Our results showed that when considering body size, the slope of the regression of egg size and ovariole number is lower than we would expect to observe by chance, as assessed by comparing to data simulated with no evolutionary correlation (Fig.
+\ref{fig:simulated}).
+However, for both adult body length and dry mass, the slope of the regressions on observed data are not within the range that would be expected under an strong negative correlation (slope of -1 in log-log space, Fig.
+\ref{fig:simulated}).
+This suggests the presence of a weak evolutionary relationship bewteeen ovariole number and egg size, when accounting for body size.
 
-We additionally tested whether ovariole number is positively correlated with adult body size, and in contrast to previous studies[@honvek1993intraspecific], we found no correlation between ovariole number and adult body weight or length across insects (p-values `r body_all_table[3,]$'pval_max'` and `r body_all_table[1,]$'pval_max'`, respectively; Fig.
-\ref{fig:PGLS-body-plot}, Table \ref{tab:PGLS-body-posterior}).
+Second, we assessed the relationship between egg size and ovariole number, accounting for body size, within four subclades of insects.
+We found that across Drosophilidae fly species, egg size is indeed strongly negatively correlated with ovariole number when accounting for body size (Fig.
+2c, Table \ref{tab:PGLS-posterior-clade}, p-value `r gsub(0,"<0.001",round(pgls_results_DIP$tTable[[8]],3))`, n=`r pgls_results_DIP$dims$N`).
+For grasshoppers and crickets (Orthoptera), beetles (Coleoptera), and wasps (Hymenoptera), we observed no significant relationship between ovariole number and egg size, even when accounting for body size (Fig.
+2d, Table \ref{tab:PGLS-posterior-clade}, \ref{fig:PGLS-Coleoptera-Hymenoptera}, Orthoptera: p-value `r round(pgls_results_ORT$tTable[[8]],3)`, n=`r pgls_results_ORT$dims$N`, Coleoptera: p-value `r round(pgls_results_COL1$tTable[[8]],3)`, n=`r pgls_results_COL1$dims$N`, Hymenoptera: p-value `r round(pgls_results_HYM$tTable[[8]],3)`, n=`r round(pgls_results_HYM$dims$N)`).
+This indicates that, while a strong negative correlation between egg size and ovariole number exists for some insects, it does not represent a universal pattern across insect clades.
 
-![**Tests of the hypothesized trade off between egg size and ovariole number.** a, Egg volume (mm^3^) and ovariole number, both log~10~ scale, n=`r nrow(egg_ON_name_pgls)` species.
-See section _Modeling ovariole number evolution_ for discussion of the enrichment of certain low values of ovariole number (points appearing vertically arranged) b, Egg volume and ovariole number, residuals to adult dry body weight, n=`r length(which(egg_ON_gilbert_genus_pgls$rank %in% tree$tip.label))` genera.
-c, Drosophilidae egg volume and ovariole number, residuals to thorax length, n=`r length(which(DIP_pgls$rank %in% dros_tree$tip.label))` species.
-d, Orthoptera egg volume and ovariole number, residuals to body length, n=`r length(which(ORT_pgls$rank %in% tree$tip.label))` genera.](manuscript_figures/Fig_Egg_V2-01.png)
+Finally, we tested whether ovariole number is positively correlated with adult body size, and in contrast to previous studies[@honvek1993intraspecific], we found no correlation between ovariole number and adult body mass or length across insects (Fig.
+\ref{fig:PGLS-body-plot}, Table \ref{tab:PGLS-posterior-body}, body mass: p-value `r round(pgls_results_body_gilbert_name$tTable[[8]],3)`, n=`r pgls_results_body_gilbert_genus$dims$N`, body length: p-value `r round(pgls_results_body_ON_family$tTable[[8]],3)`, n=`r pgls_results_body_ON_family$dims$N`).
+Of the four subclades considered, only insects in the order Orthoptera had a positive relationship between body size and ovariole number (Table \ref{tab:PGLS-posterior-body}, p-value `r round(pgls_results_body_ORT$tTable[[8]],3)`, n=`r pgls_results_body_ORT$dims$N`).
+
+![**Tests of the hypothesized trade off between egg size and ovariole number.** a, Egg volume (mm^3^) and ovariole number, both log~10~ scale; points represent insect species.
+See section _Modeling ovariole number evolution_ for discussion of the enrichment of certain low values of ovariole number (points appearing vertically arranged) b, Egg volume and ovariole number, residuals to dry adult body mass, points represent genera.
+c, Drosophilidae egg volume and ovariole number, residuals to thorax length, points represent species.
+d, Orthoptera egg volume and ovariole number, residuals to body length, points represent genera.](manuscript_figures/Fig_Egg_V2-01.png)
 
@@
 
-Using a previously reported dataset of lifetime fecundity measurements across insects[@Gilbert2007], we assessed the relationship between lifetime fecundity and ovariole number, and observed a positive, but not statistically significant, correlation (p-value `r all_table[8,]$'pval_max'`; Figs.
-3, \ref{fig:PGLS-fecundity-name}).
+Using a previously reported dataset of lifetime fecundity measurements across insects[@Gilbert2007], we assessed the relationship between lifetime fecundity and ovariole number.
+We observed a significant positive relationship (Fig.
+3, p-value `r round(pgls_results_fecundity_genus$tTable[[8]],3)`, n=`r pgls_results_fecundity_genus$dims$N`), however, a substantial fraction of repeat analyses show these results are not robust to uncertainty (`r sum(sapply(allometry_fecundity_genus,function(x){x$tTable[[8]]}) > 0.01)` of 1000 regressions are not significant, Table \ref{tab:PGLS-posterior-fecundity}).
+We note that this relationship is largely defined by the absence of insects with high ovariole number and low fecundity (Fig.
+3, empty bottom right corner), while for insects with low ovariole number, fecundity varied over more than three orders of magnitude.
 We interpret our results, in conjunction with those previously reported, to suggest that ovariole number, when variable across insects in a lineage, may be one factor among many influencing the number of eggs produced.
@@
 ![**The relationship between lifetime fecundity and ovariole number.** Both values are shown on a log~10~ scale.
-n=`r length(na.omit(fecundity_genus_pgls$fecundity))` genera, colored according to the groups shown in Fig.
+Points represent insect genera and are colored according to the groups shown in Fig.
 1b.](manuscript_figures/Fig_Fecundity_V1-01.png)
@@
 4a).
-The ovary of the well-studied species _D.
+Egg formation in the well-studied species _D.
 melanogaster_ is an example of a meroistic oogenesis mode, meaning that its ovarioles contain nurse cells of germ line origin that are connected to developing oocytes via cytoplasmic bridges[@KingRobertC1970OdiD].
@@
 An alternative arrangement is seen in telotrophic meroistic ovaries, where oocytes in each ovariole are connected to a common pool of nurse cells located in the germarium[@buning1994insect].
-Meroistic ovaries are thought to have evolved from a panoistic ancestral state, meaning they lack nurse cells[@buning1994insect].
+Meroistic ovaries are thought to have evolved from an ancestral panoistic mode, meaning they lack nurse cells[@buning1994insect].
 Using a previously published set of descriptions of these oogenesis modes across insects[@buning1994insect], we reconstructed the evolutionary transitions between these states.
-Consistent with previous analyses[@buning1994insect], we found that the ancestral insect likely had panoistic ovaries (without nurse cells), with several independent shifts to both telotrophic and polytrophic meroistic modes, and at least two reversals from meroistic back to panoistic (Figs.
+Consistent with previous analyses[@buning1994insect], we found that the ancestral insect likely had panoistic ovaries (lacking nurse cells), with several independent shifts to both telotrophic and polytrophic meroistic modes, and at least two reversals from meroistic back to panoistic (Figs.
 4b, \ref{fig:reconstruct-ovary}).
@@
 5b, cyan), from an ancestral variable state.
-These invariant lineages were objectively identified by finding regions of the phylogeny that experience extremely low rates of ovariole number diversification (Figs.
+These invariant lineages were identified by finding regions of the phylogeny that experience extremely low rates of ovariole number diversification (Figs.
 \ref{fig:BAMM-analysis-results}, \ref{fig:BAMM-anlaysis-mean-rates}).
@@
 
-We find that the rate of evolutionary change in ovariole number is correlated with the number of ovarioles: lineages with relatively low ovariole number also experience relatively low degrees of change (Fig.
+We find that the rate of evolutionary change in ovariole number is correlated with the number of ovarioles: lineages with relatively low ovariole number also experience relatively low degrees of ovariole number change (Fig.
 \ref{fig:arbutus}).
@@
 This would again be a departure from known mechanisms in _Drosophila_, in which a variable number of precursor cells are gathered into terminal filaments until the population is depleted[@godt1995mechanisms;@sarikaya2012roles].
-The determining factor for partitioning the precursor pool could be, for example, a spatially variable morphogen emanating from adjacent tissues [@lawrence1996morphogens] or a reaction-diffusion patterning process [@kondo2010reaction] within the developing ovary, as has been shown to generate fixed numbers of multicellular structures in other developmental contexts[@salazar2008tooth;@hatini2001divide;@clark2019arthropod].
+The determining factor for partitioning the precursor pool could be, for example, a spatially variable morphogen emanating from adjacent tissues [@lawrence1996morphogens] or a reaction-diffusion patterning process [@kondo2010reaction] within the developing ovary, as these have been shown to generate fixed numbers of multicellular structures in other developmental contexts[@salazar2008tooth;@hatini2001divide;@clark2019arthropod].
 These predictions could be tested by characterizing the dynamics of cell number and position across invariant lineages, and making comparisons to corresponding data from their variable relatives.

### 2021-01-18 manuscript_supplement.Rmd
adding conda environment; rerunning analyses; updating manuscript and README

@@
 title: Repeated loss of variation in insect ovary morphology highlights the role of developmental constraint in life-history evolution - Supplementary Methods
+author: 
+    - Samuel H. Church^1^\*, 
+    - Bruno A. S. De Medeiros^1,2^, 
+    - Seth Donoughe^1,3^, 
+    - Nicole L. Márquez Reyes^4^, 
+    - Cassandra G. Extavour^1,5^\*
 output:
@@
 
-```{r load_data}
+```{r load_data, cache=FALSE}
 # This script reads in and formats the ovariole number dataset
@@
 ```
-
-Samuel H. Church^1^\*, Bruno A. S. De Medeiros^1,2^, Seth Donoughe^1,3^, Nicole L. Márquez Reyes^4^, Cassandra G. Extavour^1,5^\*
 
@@
 
-Ovariole number was recorded as either an average with deviations, a range, or a single total value, with priority for recording given in that order when multiple data types were available.
+Ovariole number was recorded as either an average with deviations, a range, or a single total value.
+When multiple types of data were available from a single publication, we recorded only a single type, with priority given to averages over ranges, and to both over single total values.
 Ovariole number was recorded as the total number of ovarioles per female, summing over both the left and right adult ovaries.
 When authors reported ovariole number from a single ovary, the total value was calculated by doubling the reported value.
-When authors described differences between the two ovaries, this information was recorded in an additional column.
+When authors described differences between the two ovaries, this information was recorded in an additional notes column.
 
@@
 
-We combined the data we collected on total ovariole number with existing datasets of egg size and shape[@church2019dataset], insect lifetime fecundity and adult dry body weight[@Gilbert2007], average adult body length per insect family[@rainford2016phylogenetic], and several lineage-specific measures of adult body size[@iwata1966large;@Iwata1966a;@waloff1954number;@starmer2003phylogenetic;@reinhardt2005low].
+We combined the data we collected on total ovariole number with existing datasets of egg size and shape[@church2019dataset], insect lifetime fecundity and dry adult body mass[@Gilbert2007], average adult body length per insect family[@rainford2016phylogenetic], and several lineage-specific measures of adult body size[@iwata1966large;@Iwata1966a;@waloff1954number;@starmer2003phylogenetic;@reinhardt2005low].
 
@@
 This variation across records for the same species was accounted for in regressions by reshuffling and matching records at each iteration of the analysis.
-In order to test whether results were robust with a larger sample size, we also matched records at the genus level following the same reshuffling method (Fig.
+We also matched records at the genus level following the same reshuffling method, which allowed us to test whether results were robust with a larger sample size when an exact species match was not available (Fig.
 \ref{fig:PGLS-egg-ON-genus}).
@@
 
-Ovariole number, egg volume, lifetime fecundity, and adult body weight[@Gilbert2007] were combined by matching records at the species level and genus level, using the same method as described above (Figs.
+Ovariole number, egg volume, lifetime fecundity, and adult body mass[@Gilbert2007] were combined by matching records at the species level and genus level, using the same method as described above (Figs.
 2b, 3, and \ref{fig:PGLS-Gilbert-name}).
@@
 
-set.seed(12345)  # set seed for cobmination
+set.seed(12345)  # set seed for combination
 source("analysis/combine_body_egg_datasets.R")
@@
 
-```{r PGLS-Gilbert-name, fig.cap="**Egg volume vs ovariole number, phylogenetic residuals to dry adult body weight, matching records at the species level.**", fig.width=3, fig.height=3, include=TRUE}
+```{r PGLS-Gilbert-name, fig.cap="**Egg volume vs ovariole number, phylogenetic residuals to dry adult body mass, matching records at the species level.**", fig.width=3, fig.height=3, include=TRUE}
 ### Gilbert egg vs ON, ind body weight name pgls
@@
 
-\textcolor{red}{We ran each Phylogenetic Generalized Least Squares regression over the Maximum Clade Credibilty (MCC) tree.
-For each regression, we also repeated each PGLS analysis 1000 times,} accounting for phylogenetic and phenotypic uncertainty, using the R packages ape (version `r packageVersion('ape')`)[@paradis2004ape] and nlme (version `r packageVersion('nlme')`)[@pinheiro2014r].
+We ran each Phylogenetic Generalized Least Squares (PGLS) regression over the Maximum Clade Credibilty (MCC) tree.
+For each regression, we also repeated each PGLS analysis 1000 times, accounting for phylogenetic and phenotypic uncertainty, using the R packages ape (version `r packageVersion('ape')`)[@paradis2004ape] and nlme (version `r packageVersion('nlme')`)[@pinheiro2014r].
 In these analyses we used a Brownian Motion based covariance matrix for traits.
@@
 
-\textcolor{red}{For regressions of egg size and ovariole number when accounting for adult body size, we compared the results of our regression analyses to distributions estimated using simulated data under alternative hypotheses.
-We fit a Brownian motion model to the phylogenetic residuals of egg size and body size (R package geiger}[@harmon2007geiger]\textcolor{red}{, version `r packageVersion('geiger')`), and then used the parameters of this fitted model to simulate new datasets (R package phylolm}[@tung2014linear]\textcolor{red}{, version `r packageVersion('phylolm')`).
-We performed this resimulation using the datasets of egg size and body length at the family level, and egg size and body weight at the genus level.
-We simulated 1000 datasets each under two hypotheses: no correlation (slope=0) and a strong negative correlation (slope=-1:1).
-We performed the regressions as described above and compared the distribution of p-values and slopes to values from regressions on observed data.}
-
-```{r simulated, fig.cap="**\\textcolor{red}{Using simulated data to test alternative hypotheses of evolutionary relationships}** Top row, distributions of p-values over 1000 replicate regressions, dashed black line indicates threshold of 0.01.
+For regressions of egg size and ovariole number when accounting for adult body size, we compared the results of our regression analyses to distributions estimated using simulated data under alternative hypotheses.
+We fit a Brownian motion model to the phylogenetic residuals of egg size and body size (R package geiger, version `r packageVersion('geiger')`)[@harmon2007geiger], and then used the parameters of this fitted model to simulate new datasets (R package phylolm, version `r packageVersion('phylolm')`)[@tung2014linear].
+We performed this resimulation using the datasets of egg size and body length at the family level, and egg size and body mass at the genus level.
+We simulated 1000 datasets each under two hypotheses: no correlation (slope=0) and a strong negative correlation (slope=-1).
+We performed the regressions as described above and compared the distribution of p-values and slopes to values from regressions on observed data.
+
+```{r simulated, fig.cap="**Using simulated data to test alternative hypotheses of evolutionary relationships.** Top row, distributions of p-values over 1000 replicate regressions, dashed black line indicates threshold of 0.01.
 Bottom row, distribution of estimated slopes between egg size and ovariole number, dashed lines indicate 95% interval of simulated distributions.
 Left, comparing egg size and ovariole number, accounting for body length at the family level.
-Right, comparing egg size and ovariole number, accounting for body weight at the genus level.
+Right, comparing egg size and ovariole number, accounting for body mass at the genus level.
 Red=observed values, gray=simulated with no correlation, black=simulated with a -1:1 correlation.
@@
 
-load("analysis/pgls_results_Fall2020/run_pgls_simulated_allometry_body_egg_ON_family_-1_image.RData")
+load("analysis/pgls_results_Jan2021/run_pgls_simulated_allometry_body_egg_ON_family_-1_image.RData")
 sim_corr_tstat_dist <- sapply(allometry_distribution_raw,get_tstat,n=1)
 sim_corr_pval_dist <- sapply(allometry_distribution_raw,get_pval,n=1)
-load("analysis/pgls_results_Fall2020/run_pgls_simulated_allometry_body_egg_ON_family_0_image.RData")
+load("analysis/pgls_results_Jan2021/run_pgls_simulated_allometry_body_egg_ON_family_0_image.RData")
 sim_nocorr_tstat_dist <- sapply(allometry_distribution_raw,get_tstat,n=1)
 sim_nocorr_pval_dist <- sapply(allometry_distribution_raw,get_pval,n=1)
-load("analysis/pgls_results_Fall2020/run_posterior_pgls_egg_size_ovariole_number.RData")
+load("analysis/pgls_results_Jan2021/run_posterior_pgls_egg_size_ovariole_number.RData")
 obs_tstat_dist <- sapply(allometry_distribution_raw,get_tstat,n=3)
@@
 
-load("analysis/pgls_results_Fall2020/run_pgls_simulated_allometry_gilbert_genus_-1_image.RData")
+load("analysis/pgls_results_Jan2021/run_pgls_simulated_allometry_gilbert_genus_-1_image.RData")
 sim_corr_tstat_dist <- sapply(allometry_distribution_raw,get_tstat,n=1)
 sim_corr_pval_dist <- sapply(allometry_distribution_raw,get_pval,n=1)
-load("analysis/pgls_results_Fall2020/run_pgls_simulated_allometry_gilbert_genus_0_image.RData")
+load("analysis/pgls_results_Jan2021/run_pgls_simulated_allometry_gilbert_genus_0_image.RData")
 sim_nocorr_tstat_dist <- sapply(allometry_distribution_raw,get_tstat,n=1)
 sim_nocorr_pval_dist <- sapply(allometry_distribution_raw,get_pval,n=1)
-load("analysis/pgls_results_Fall2020/run_posterior_pgls_egg_size_ovariole_number.RData")
+load("analysis/pgls_results_Jan2021/run_posterior_pgls_egg_size_ovariole_number.RData")
 obs_tstat_dist <- sapply(allometry_distribution_raw,get_tstat,n=5)
@@
 #geom_vline(xintercept=0.05,colour="black",linetype="longdash") + 
-ggtitle("egg size vs ov. number, body weight") + 
+ggtitle("egg size vs ov. number, body mass") + 
 xlab("p-value") +
@@
 
-```{r PGLS-body-plot, fig.cap="**Adult body size vs ovariole number.** Adult body weight compared at the genus level (left), and average adult body length compared at the family level (right).", fig.width=6, fig.height=3, include=TRUE}
+```{r PGLS-body-plot, fig.cap="**Adult body size vs ovariole number.** Adult body mass compared at the genus level (left), and average adult body length compared at the family level (right).", fig.width=6, fig.height=3, include=TRUE}
 ### body length vs ON, family level
@@
 gilbert_body <- run_plotting_pgls(ON_body_gilbert_genus_pgls,tree,"body_gilbert") + xlab(bquote('ovariole number,'~log[10])) +
-ylab(bquote('adult dry body weight (g),'~log[10]))
+ylab(bquote('dry adult body mass (g),'~log[10]))
 
@@
 ### PGLS TABLE EGG SIZE OVARIOLE NUMBER
-load("analysis/pgls_results_Fall2020/run_posterior_pgls_egg_size_ovariole_number.RData")
+load("analysis/pgls_results_Jan2021/run_posterior_pgls_egg_size_ovariole_number.RData")
 
@@
     "ovariole number vs egg volume",
-    "ovariole number vs egg volume, residuals to body weight",
-    "ovariole number vs egg volume, residuals to body weight",
+    "ovariole number vs egg volume, residuals to body mass",
+    "ovariole number vs egg volume, residuals to body mass",
     "ovariole number vs egg volume, residuals to body length")
@@
         "latex",
+        align=rep('l', 6),
         caption="Results of PGLS analysis of ovariole number and egg size across a posterior distribution", booktabs = T, linesep = "") %>% 
@@
 ### PGLS TABLE SUBCLADE
-load("analysis/pgls_results_Fall2020/run_posterior_pgls_subclade.RData")
+load("analysis/pgls_results_Jan2021/run_posterior_pgls_subclade.RData")
 
@@
 ### PGLS TABLE BODY SIZE
-load("analysis/pgls_results_Fall2020/run_posterior_pgls_body_size.RData")
+load("analysis/pgls_results_Jan2021/run_posterior_pgls_body_size.RData")
 
@@
 body_pgls_analysis_names <- c("ovariole number vs body length",
-    "ovariole number vs body weight",
-    "ovariole number vs body weight",
+    "ovariole number vs body mass",
+    "ovariole number vs body mass",
     "Drosophilidae ovariole number vs thorax length",
@@
         "latex",
+        align=rep('l', 6),
         caption="Results of PGLS analysis of ovariole number and body size across a posterior distribution", booktabs = T, linesep = "") %>% 
@@
 ### PGLS TABLE FECUNDITY
-load("analysis/pgls_results_Fall2020/run_posterior_pgls_fecundity.RData")
+load("analysis/pgls_results_Jan2021/run_posterior_pgls_fecundity.RData")
 
@@
 Using these 106 records for ovary type, we reconstructed the ancestral state at each node of the published phylogeny of `r length(pp$phy$tip.label)` insect genera using an equal-rates model that allows for missing data (Fig.
-\ref{fig:reconstruct-ovary}, R package corHMM,[@beaulieu2013identifying] version `r packageVersion('corHMM')`, function rayDISC, node.states = 'marginal').
+\ref{fig:reconstruct-ovary}, R package corHMM, version 1.22[@beaulieu2013identifying] function rayDISC, node.states = 'marginal').
 
@@
 We then compared the fit of models of trait evolution that take into account these shifts in oogenesis mode against those that do not.
-These comparisons were performed with the R package OUwie[@beaulieu2012modeling]  (version `r packageVersion('OUwie')`).
+These comparisons were performed with the R package OUwie (version 1.57)[@beaulieu2012modeling].
 
@@
 
-We evaluated the fit of a Brownian Motion (BM) model for ovariole number evolution using the R package arbutus[@pennell2015model], version `r packageVersion('arbutus')`  (Fig.
+We evaluated the fit of a Brownian Motion (BM) model for ovariole number evolution using the R package arbutus (version `r packageVersion('arbutus')`[@pennell2015model], Fig.
 \ref{fig:arbutus}).
-In this approach, a BM model is fit to the data (R package geiger[@harmon2007geiger], version `r packageVersion('geiger')`), and the resulting parameters of the model are used to simulate 1000 new datasets.
+In this approach, a BM model is fit to the data (R package geiger, version `r packageVersion('geiger')`)[@harmon2007geiger], and the resulting parameters of the model are used to simulate 1000 new datasets.
 Six statistical parameters are used to compare the phylogenetic contrasts of the observed data to the simulated data, and their interpretations are as follows[@pennell2015model]:
@@
 ```{r BAMM-analysis-setup}
+#set.seed(12345)  # set seed for BAMM
+
 ### Select a continuous trait to analyze
@@
 
-Given the result that our dataset contains substantial rate heterogeneity, we identified regions of the tree with high and low rates of ovariole number evolution using the software BAMM[@rabosky2014automatic], version 2.5.0.
+Given the result that our dataset contains substantial rate heterogeneity, we identified regions of the tree with high and low rates of ovariole number evolution using the software BAMM (version 2.5.0)[@rabosky2014automatic].
 For this analysis, we calculated the average ovariole number for each genus in the insect phylogeny[@church2019insect].
@@
 ### Assess convergence
-mcmcout <- read.csv("analysis/BAMM_results_June2020/bamm_mcmc_out.txt", header=T)
+mcmcout <- read.csv("analysis/BAMM_results_Jan2021/bamm_mcmc_out.txt", header=T)
 
@@
 \ref{fig:BAMM-analysis-burnin}).
-Running BAMM for the maximum possible number of generations and selecting the optimum burn-in\ref{fig:BAMM-analysis-burnin} resulted in an effective size for the number of shifts of `r round(es_nshifts,2)`, and for log-likelihood of `r round(es_logl,2)`.
+Running BAMM for the maximum possible number of generations and selecting the optimum burn-in (Fig.
+\ref{fig:BAMM-analysis-burnin}) resulted in an effective size for the number of shifts of `r round(es_nshifts,2)`, and for log-likelihood of `r round(es_logl,2)`.
 Repeated BAMM analyses showed similar distributions of high and low rate regimes, indicating the implications for ovariole number evolution are robust to uncertainty in rate estimates.
@@
 ### Read in the BAMM results and plot the shifts on the tree
-bamm_tree <- ladderize(read.tree(file="analysis/BAMM_results_June2020/bamm_tree.tre"))
+bamm_tree <- ladderize(read.tree(file="analysis/BAMM_results_Jan2021/bamm_tree.tre"))
 
 # get event data
-edata <- getEventData(bamm_tree, eventdata = paste("analysis/BAMM_results_June2020/bamm_event_data.txt",sep=""), burnin=burnin, type="trait")
+edata <- getEventData(bamm_tree, eventdata = paste("analysis/BAMM_results_Jan2021/bamm_event_data.txt",sep=""), burnin=burnin, type="trait")
 ```
@@
 \ref{fig:BAMM-analysis-results}).
-Visualizing the distribution of mean rates along branches revealed a discontinuous distribution, with one peak at a moderate rate of evolution and another at an extremely low rate, separated from the first peak by over six orders of magnitude (Fig.
+Visualizing the distribution of mean rates along branches revealed a discontinuous distribution, with one peak at a moderate rate of evolution and several clusters at extremely low rates, separated from the first peak by over six orders of magnitude (Fig.
 \ref{fig:BAMM-anlaysis-mean-rates}).
@@
 ```{r BAMM-egg-setup}
+#set.seed(12345)  # set seed for egg BAMM
+
 tree <- genus_mcc_tree
@@
 ```{r BAMM-egg-read}
-egg_mcmcout <- read.csv("analysis/BAMM_results_June2020/egg_bamm_mcmc_out.txt", header=T)
+egg_mcmcout <- read.csv("analysis/BAMM_results_Jan2021/egg_bamm_mcmc_out.txt", header=T)
 
@@
 
-egg_bamm_tree <- ladderize(read.tree(file="analysis/BAMM_results_June2020/egg_bamm_tree.tre"))
-egg_edata <- getEventData(egg_bamm_tree, eventdata = paste("analysis/BAMM_results_June2020/egg_bamm_event_data.txt",sep=""), burnin=egg_burnin, type="trait")
+egg_bamm_tree <- ladderize(read.tree(file="analysis/BAMM_results_Jan2021/egg_bamm_tree.tre"))
+egg_edata <- getEventData(egg_bamm_tree, eventdata = paste("analysis/BAMM_results_Jan2021/egg_bamm_event_data.txt",sep=""), burnin=egg_burnin, type="trait")
 ```
@@
 \ref{fig:BAMM-egg-plot}).
-This analysis converged in $20*10^8$ generations (the effective size for the number of shifts and log-likelihood were >200, `r round(egg_es_nshifts,2)` and `r round(egg_es_logl,2)` respectively).
+This analysis converged in $7*10^9$ generations (the effective size for the number of shifts and log-likelihood were >200, `r round(egg_es_nshifts,2)` and `r round(egg_es_logl,2)` respectively).
 We compared the correlation between the rates of trait diversification for ovariole number and egg volume by matching the mean rate predicted for each insect genus (the tips of the phylogeny from the BAMM analyses, Fig.

### 2021-01-18 analysis/reconstruct_ancestral_ovary_morphology_type.R

@@ -4,8 +4,9 @@ library(OUwie)
 library(corHMM)
 library(parallel)
 library(geiger)
-source("analysis/get_ovariole_number_data")
-egg_database <- read.delim("analysis/datafiles/egg_database_final_values_July17.tsv",header=T,stringsAsFactors=F)
+set.seed(12345)
+source("analysis/get_ovariole_number_data.R")
+egg_database <- read.delim("analysis/datafiles/egg_database_final_values_July2019.tsv",header=T,stringsAsFactors=F)
 
 ### These commands are used to estimate the evolutionary shifts in ovary type 
 # read in table of taxonomic groups, labeled by ovary type, as recorded in Buning 1994
@@ -66,6 +67,8 @@ pp_pruned <- drop.tip(tree,setdiff(tree$tip.label,new_tree_dataset$species))
 
 # Reconstruct ancestral ovary type (with missing data), Equal Rates model
 pp <- rayDISC(pp_pruned,na.omit(as.data.frame(new_tree_dataset[,c(1,2)])),model="ER",node.states="marginal")
+
+save.image("reconstruct_ancestral_ovary_morphology_type.RData")
 
 
 

### 2021-01-18 analysis/run_mode_of_oogenesis_model_comparison.R

@@ -4,7 +4,7 @@ library(OUwie)
 library(corHMM)
 library(parallel)
 library(geiger)
-source("analysis/get_ovariole_number_data")
+source("analysis/get_ovariole_number_data.R")
 g_trees <- genus_trees[1:100] # select 100 trees from the posterior distribution
 downsample_factor <- 1 # do not downsample the data
 fam_count_threshold <- 1 # use all families
@@ -114,8 +114,8 @@ run_model_comparison <- function(tree) {
 
 	# Extract estimated tip states
 	est_tips <- pp$tip.states %>% as.data.frame() %>%  
-				mutate(discrete = ifelse(`1` == pmax(`1`,`2`,`3`),1,
-						ifelse(`2` == pmax(`1`,`2`,`3`),2,3)),
+				mutate(discrete = ifelse(V1 == pmax(V1,V2,V3),1,
+						ifelse(V2 == pmax(V1,V2,V3),2,3)),
 						species = pp$phy$tip.label) %>% 
 				select(species,discrete)
 

### 2021-01-18 set_up_conda_environment.sh

@@ -1,0 +1,94 @@
+### This code was written by SHC in Jan 2021
+### This gives the commands used to build the R environment from which
+###    all analyses were performed
+
+# build an R environment with base R
+conda create -n r-4.0.3 -c conda-forge r-base=4.0.3
+
+# activate the environment
+conda activate r-4.0.3
+
+# install R packages
+conda install -c conda-forge r-devtools
+R
+install.packages("ape")
+install.packages("dplyr")
+install.packages("plyr")
+install.packages("nlme")
+install.packages("phytools")
+install.packages("ggplot2")
+install.packages("BAMMtools")
+install.packages("coda")
+install.packages("OUwie")
+#conda install -c conda-forge r-rmpfr
+#conda install -c anaconda gmp
+install.packages("corHMM")
+install.packages("grid")
+install.packages("gridExtra")
+install.packages("gtable")
+install.packages("geiger")
+install.packages("RColorBrewer")
+install.packages("rmarkdown")
+install.packages("bookdown")
+install.packages("kableExtra")
+install.packages("ggplotify")
+
+if (!requireNamespace("BiocManager", quietly = TRUE))
+    install.packages("BiocManager")
+BiocManager::install("ggtree")
+
+devtools::install_github("mwpennell/arbutus")
+
+sessionInfo()
+#	R version 4.0.3 (2020-10-10)
+#	Platform: x86_64-apple-darwin13.4.0 (64-bit)
+#	Running under: macOS Catalina 10.15.7
+#	
+#	Matrix products: default
+#	BLAS/LAPACK: /opt/anaconda3/envs/r-4.0.3/lib/libopenblasp-r0.3.12.dylib
+#	
+#	locale:
+#	[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8
+#	
+#	attached base packages:
+#	[1] grid      stats     graphics  grDevices utils     datasets  methods  
+#	[8] base     
+#	
+#	other attached packages:
+#	 [1] arbutus_0.1        ggtree_2.4.1       ggplotify_0.0.5    kableExtra_1.3.1  
+#	 [5] bookdown_0.21      rmarkdown_2.6      gtable_0.3.0       gridExtra_2.3     
+#	 [9] corHMM_2.6         GenSA_1.1.7        OUwie_2.5          RColorBrewer_1.1-2
[…]

### 2021-01-19 manuscript.Rmd
updating Rmd and manuscript files in association with submission

@@
 output:
+  #word_document
   bookdown::pdf_document2:
@@
 For regressions at the species and genus level, we reshuffled and matched records for each iteration to account for variation across records for the same taxon.
-For regressions at the family level we recalculated the average ovariole number per insect family, randomly downsampling the representation for each family by half.
+For regressions at the family level, we recalculated the average ovariole number per insect family, randomly downsampling the representation for each family by half.
 To weight traits by body size, we calculated the phylogenetic residuals[@revell2010phylogenetic] of each trait to body size, and then compared the evolution of these residuals using a PGLS regression.
@@
 
-Other comparisons of model fit were performed using the R package geiger(version `r packageVersion('geiger')`)[@harmon2007geiger] and validated using a parametric bootstrap with the R package arbutus (version `r packageVersion('arbutus')`)[@pennell2015model].
+Other comparisons of model fit were performed using the R package geiger (version `r packageVersion('geiger')`)[@harmon2007geiger] and validated using a parametric bootstrap with the R package arbutus (version `r packageVersion('arbutus')`)[@pennell2015model].
 See supplementary methods section \ref{sec:bootstrap}.
@@
 For this analysis, we calculated the average ovariole number (log~10~ transformed) for each genus present in the phylogeny (507 taxa).
-We used the R package BAMMtools (version `r packageVersion('BAMMtools')`)[@rabosky2014bamm] to select appropriate priors, and ran BAMM for the maximum number of generations ($2*10^{-9}$), sampling every 10^6^ generations.
+We used the R package BAMMtools (version `r packageVersion('BAMMtools')`)[@rabosky2014bamm] to select appropriate priors, and ran BAMM for the maximum number of generations ($2*10^{9}$), sampling every 10^6^ generations.
 Convergence was evaluated both visually (Fig.
@@
 
-The code and phylogenetic trees required to reproduce all the analyses, figures, and generate the manuscript files are provided at 'https://github.com/shchurch/insect_ovariole_number_evolution_2020', commit 6cf446a.
+The code and phylogenetic trees required to reproduce all the analyses, figures, and generate the manuscript files are provided at 'https://github.com/shchurch/insect_ovariole_number_evolution_2020', commit 9876387.
 All analyses performed in R (version 4.0.3) were done so in a clean environment, built with conda (version 4.9.2), and instructions for rebuilding this environment are provided in the same repository.
@@
 \ref{fig:simulated}).
-However, for both adult body length and dry mass, the slope of the regressions on observed data are not within the range that would be expected under an strong negative correlation (slope of -1 in log-log space, Fig.
+However, for both adult body length and dry mass, the slope of the regressions on observed data are not within the range that would be expected under a strong negative correlation (slope of -1 in log-log space, Fig.
 \ref{fig:simulated}).
-This suggests the presence of a weak evolutionary relationship bewteeen ovariole number and egg size, when accounting for body size.
+This suggests the presence of a weak evolutionary relationship between ovariole number and egg size, when accounting for body size.
 
@@
 \ref{fig:BAMM-analysis-results}), meaning that for some lineages ovariole number has evolved rapidly where in others, ovariole number has evolved very slowly or not at all.
-The most striking example of this are the multiple lineages which have independently evolved invariant or near-invariant ovariole number across taxa (e.g.
-nearly all Lepidoptera have exactly eight ovarioles, Fig.
+The most striking example of this are the multiple lineages which have independently evolved invariant or near-invariant ovariole number across taxa (e.g., nearly all Lepidoptera have exactly eight ovarioles, Fig.
 5b, cyan), from an ancestral variable state.
@@
 NM performed the initial literature search, data collection, and analysis.
-BASdM assembled all phylogenies for analysis.
-SD, BASdM, and CGE contributed to data analysis, visualization, discussion, and writing.
+BASM assembled all phylogenies for analysis.
+SD, BASM, and CGE contributed to data analysis, visualization, discussion, and writing.
 

### 2021-03-20 manuscript.Rmd
changes to Rmd files during revision process

--- 2021-01-18 manuscript.Rmd
@@
 ---
-title: Repeated loss of variation in insect ovary morphology highlights the role of developmental constraint in life-history evolution
+title: Repeated loss of variation in insect ovary morphology highlights the role of development in life-history evolution
 output:
   bookdown::pdf_document2:
+  #bookdown::word_document2: default
   keep_tex: yes
@@
 The number of offspring an organism can produce is a key component of its evolutionary fitness and life-history.
-Here we perform a test of the hypothesized trade off between the number and size of offspring using thousands of descriptions of the number of egg-producing compartments in the insect ovary (ovarioles), a common proxy for potential offspring number in insects.
+Here we perform a test of the hypothesized trade-off between the number and size of offspring using thousands of descriptions of the number of egg-producing compartments in the insect ovary (ovarioles), a common proxy for potential offspring number in insects.
 We find evidence of a negative relationship between egg size and ovariole number when accounting for adult body size.
-However in contrast to prior claims, we note that this relationship is not generalizable across all insect clades, and we highlight several factors that may have contributed to this size-number trade off being stated as a general rule in previous studies.
+However in contrast to prior claims, we note that this relationship is not generalizable across all insect clades, and we highlight several factors that may have contributed to this size-number trade-off being stated as a general rule in previous studies.
 We reconstruct the evolutionary history of the nurse cell arrangement within the ovariole, and show that the diversification of ovariole number and egg size have both been largely independent of nurse cell presence or position within the ovariole.
@@
 Offspring number is a fundamental parameter in the study of life-history[@stearns1992evolution].
-This number differs widely between organisms[@stearns1992evolution], and its variation is the foundation for several hypotheses about life-history evolution, including the prediction that there is an evolutionary trade off between the number of offspring and their size (e.g.
+This number differs widely between organisms[@stearns1992evolution], and its variation is the foundation for several hypotheses about life-history evolution, including the prediction that there is an evolutionary trade-off between the number of offspring and their size (e.g.
 egg size)[@smith1974optimal;@lack1947significance;@stearns1992evolution].
@@
 However, without an understanding of the phylogenetic distribution of ovariole number, this hypothesized relationship cannot be assessed across insects.
-Here we tested for the presence of a general trade off between ovariole number and egg size by collecting thousands of records of ovariole number from the published literature, placing them in a phylogenetic context, and comparing them to other datasets of insect reproductive morphology.
+Here we tested for the presence of a general trade-off between ovariole number and egg size by collecting thousands of records of ovariole number from the published literature, placing them in a phylogenetic context, and comparing them to other datasets of insect reproductive morphology.
 
@@
 Here we compiled `r nrow(raw_ON)` records of ovariole number from across `r ON %>% distinct(order) %>% nrow()` orders, `r ON %>% distinct(family) %>% nrow()` families, and `r ON %>% filter(species != "") %>% distinct(genus,species) %>% nrow()` species of insects.
-We combined these data with published datasets of egg size[@church2019dataset], fecundity[@Gilbert2007], body size[@rainford2016phylogenetic], and mode of oogenesis[@buning1994insect] to test hypotheses about reproductive evolution.
-In these analyses we used an existing phylogeny of insects[@church2019insect] to analyze evolutionary patterns in ovariole number, and found that hypotheses about reproductive trait evolution do not hold generally true across insects.
-Instead we observe that the phylogenetic distribution of ovariole number suggests a model where the developmental mechanisms that govern ovariole number have shifted between variable and invariant states several times over the course of insect evolution.
-We propose that the developmental mechanisms used to establish ovariole number in well-studied insects such as _Drosophila melanogaster_ are unlikely to regulate ovariole number in all insects.
+We combined these data with published datasets of egg size[@church2019dataset], fecundity[@Gilbert2007;@gilbert2010parental], and body size[@rainford2016phylogenetic], to test hypotheses about \textcolor{red}{the evolutionary trade-off between offspring size and number}.
+In these analyses we used an existing phylogeny of insects[@church2019insect] to analyze evolutionary patterns in ovariole number, and found that hypotheses about \textcolor{red}{life-history} evolution do not hold generally true across insects.
+\textcolor{red}{We then combined these data with published observations of the mode of oogenesis}[@buning1994insect]\textcolor{red}{ and reconstructed the evolutionary history of the presence and position of nurse cells that contribute to the oocyte during oogenesis.
+We tested whether patterns in the distribution of ovariole number, egg size, or egg shape were driven by the evolution of nurse cells, and found no significant results.} Instead we observe that the phylogenetic distribution of ovariole number suggests a model where the developmental mechanisms that govern ovariole number have shifted between variable and invariant states several times over the course of insect evolution.
+Based on this finding, we propose that the developmental mechanisms used to establish ovariole number in well-studied insects such as _Drosophila melanogaster_ are unlikely to regulate ovariole number in all insects.
 
@@
 We searched the published literature for references to insect ovariole number using a predetermined set of 131 search terms, entered into Google Scholar (scholar.google.com) between June and October of 2019.
-Each search term consisted of an insect taxonomic group and the words "ovariole number".
-This list was created to include all insect orders, many large insect families, and groups well-represented in the insect egg dataset[@church2019dataset].
-For each search term, we evaluated all publications in the first page of results (ten publications).
+Each search term was \textcolor{red}{comprised of} an insect taxonomic group and the words "ovariole number".
+\textcolor{red}{The taxonomic groups used to search included} all insect orders, many large insect families, and taxonomic groups that are well-represented in the insect egg dataset[@church2019dataset].
+\textcolor{red}{For each Google Scholar search}, we evaluated all publications in the first page of results (ten publications).
 For 61 search terms that had a large number of informative hits, significant representation in the egg dataset, or that corresponded to very speciose groups, we evaluated an additional 20 publications.
@@
 
-We combined the data we collected on total ovariole number with existing datasets of egg size and shape[@church2019dataset], insect lifetime fecundity and dry adult body mass[@Gilbert2007], average adult body length per insect family[@rainford2016phylogenetic], several lineage-specific measures of adult body size[@iwata1966large;@Iwata1966a;@waloff1954number;@starmer2003phylogenetic;@reinhardt2005low], and the mode of oogenesis[@buning1994insect].
+\textcolor{red}{For records of ovariole number that reported intraspecific variation in ovariole number, we calculated the percent difference as follows: if ovariole number was reported as a range, percent difference was calculated as the $100 * ((max - min) / median)$; if ovariole number was reported as an average with deviations, percent difference was calculated as $100 * ((2 * deviation) / mean)$.
+When independent observations of ovariole number for a given species were available across multiple published records, we calculated the percent difference as the $100 * ((max - min) / median)$.}
+
+We combined the data we collected on total ovariole number with existing datasets of egg size and shape[@church2019dataset], insect lifetime fecundity and dry adult body mass[@Gilbert2007;@gilbert2010parental;@gilbert2011insect], average adult body length per insect family[@rainford2016phylogenetic], several lineage-specific measures of adult body size[@iwata1966large;@Iwata1966a;@waloff1954number;@starmer2003phylogenetic;@reinhardt2005low], and the mode of oogenesis[@buning1994insect].
 See supplementary methods section \ref{sec:combining} for details.
@@
 The analyses in this manuscript were performed using the insect phylogeny published in Church et al., 2019[@church2019insect], unless otherwise specified.
-Analyses of insect family-level ovariole number, egg size, and body size were performed using the insect phylogeny published in Rainford et.
+\textcolor{red}{For regressions involving body size data that were reported as insect family-level averages,} we used the insect phylogeny published in Rainford et.
 al, 2014[@rainford2014phylogenetic].
@@
 For regressions at the family level we recalculated the average ovariole number per insect family, randomly downsampling the representation for each family by half.
-To weight traits by body size, we calculated the phylogenetic residuals[@revell2010phylogenetic] of each trait to body size, and then compared the evolution of these residuals using a PGLS regression.
+To weight traits by body size, we calculated the phylogenetic residuals[@revell2009size] of each trait to body size, and then compared the evolution of these residuals using a PGLS regression.
 See supplementary methods section \ref{sec:PGLS} for details.
@@
 We considered two such hypotheses: no evolutionary correlation with ovariole number, and a strong correlation with ovariole number (slope of -1).
-For each trait we simulated 1,000 datasets using evolutionary parameters fit under a Brownian Motion model in the R packages geiger (version `r packageVersion('geiger')`)[@revell2010phylogenetic], and phylolm (version `r packageVersion('phylolm')`)[@tung2014linear].
+For each trait we simulated 1,000 datasets using evolutionary parameters fit under a Brownian Motion model in the R packages geiger (version `r packageVersion('geiger')`)[@harmon2007geiger], and phylolm (version `r packageVersion('phylolm')`)[@tung2014linear].
 
@@
 
-All phylogenetic regressions were performed using both the maximum clade credibility (MCC) tree.
-We considered a relationship significant when the MCC p-value was below the threshold 0.01.
-To assess the robustness of results to uncertainty in phylogenetic relationships, we also report the number of regressions over the posterior distribution of phylogenetic trees that gave a significant result (see Table \ref{tab:PGLS-posterior-egg-ON}).
+All phylogenetic regressions were performed using the maximum clade credibility (MCC) tree \textcolor{red}{(the tree with highest credibility score from the posterior distribution of the Bayesian analysis)}.
+We considered a relationship significant when the p-value was below the threshold 0.01.
+To assess the robustness of results to uncertainty in phylogenetic relationships, we also \textcolor{red}{repeated these analyses over the posterior distribution of phylogenetic trees} and report the number of regressions that gave a significant result (see Table \ref{tab:PGLS-posterior-egg-ON}).
 
 For two comparisons, we validated that our tests had sufficient statistical power using the selected threshold by comparing the distribution of p-values from regressions of observed data to regressions of data simulated under alternative hypotheses.
-We compared the results of analyses of our observed to those based on simulated data to evaluate the likelihood of false positives (comparing to data simulatd under no correlation) and false negatives (comparing to data simulated with strong correlation).
+We compared the results of analyses of our observed to those based on simulated data to evaluate the likelihood of false positives (comparing to data \textcolor{red}{simulated} under no correlation) and false negatives (comparing to data simulated with strong correlation).
 
@@
 
-Ovariole number varies by at least four orders of magnitude across insects (Fig.
+Ovariole number varies by at least four orders of magnitude across \textcolor{red}{insect species} (Fig.
 1b).
@@
 In these insects one of the two ovaries presumably established during embryogenesis is reported to atrophy during development[@pollock1996viviparous;@pluot1979evolution], resulting in an asymmetric adult reproductive system.
+\textcolor{red}{We also evaluated intraspecific variation in ovariole number, and found that, for species for which it has been reported, the average percent difference number within species is between 10\% and 100\% of the median value (Fig.
+\ref{fig:intraspecific-variation}).}
 
@@
 2a, Table \ref{tab:PGLS-posterior-egg-ON}, p-value `r round(pgls_results_egg_ON_name$tTable[[8]],3)`, n=`r pgls_results_egg_ON_name$dims$N`).
-We also compared egg size and ovariole number across insect genera and did not observe a significant relationship (Fig.
+We also compared egg size and ovariole number, \textcolor{red}{combining data from species within the same genus to increase sample size,} and again did not observe a significant relationship (Fig.
 \ref{fig:PGLS-egg-ON-genus}, p-value `r round(pgls_results_egg_ON_genus$tTable[[8]],3)`, n=`r pgls_results_egg_ON_genus$dims$N`).
+\textcolor{red}{To verify this finding was not driven by the high ovariole numbers seen in the queens of some eusocial insects, we repeated this comparison excluding insects from families with eusocial representatives, with the same result (Fig.
+\ref{fig:PGLS-egg-ON-genus-noneusocial}, p-value `r round(pgls_results_egg_ON_genus_noneusocial$tTable[[8]],3)`, n=`r pgls_results_egg_ON_genus_noneusocial$dims$N`).}
 
-Given that this predicted relationship is often conditioned on taking into account adult body size[@stewart1991reproductive;@starmer2003phylogenetic], we combined data on ovariole number and egg size with several measurements of body size, and applied a size correction that accounts for phylogenetic relationships[@revell2009size].
+Given that this predicted relationship is often conditioned on body size, \textcolor{red}{which is predicted to limit total potential reproductive investment}[@stewart1991reproductive;@starmer2003phylogenetic], we combined data on ovariole number and egg size with data on insect adult body mass[@Gilbert2007;@gilbert2010parental;@gilbert2011insect] and length[@rainford2016phylogenetic].
 When accounting for adult body mass, we observed a significant negative relationship between egg size and ovariole number (Fig.
-2b, \ref{fig:PGLS-Gilbert-name}, p-value `r round(pgls_results_gilbert_genus$tTable[[8]],3)`, n=`r pgls_results_gilbert_genus$dims$N`).
+2b, \ref{fig:PGLS-Gilbert-name}, p-value `r round(pgls_results_gilbert_genus$tTable[[8]],3)`, slope `r round(pgls_results_gilbert_genus$coefficients[[2]],3)`, n=`r pgls_results_gilbert_genus$dims$N`).
 To evaluate the robustness of this result, we repeated the analysis 1000 times, taking into account uncertainty in both the phylogeny and trait measurements.
@@
 We performed the same comparison accounting for adult body length, and likewise observed a significant negative relationship (Fig.
-\ref{fig:PGLS-family-length}, p-value `r gsub(0,"<0.001",round(pgls_results_body_egg_ON_family$tTable[[8]],3))`, n=`r pgls_results_body_egg_ON_family$dims$N`), supported by `r sum(sapply(allometry_body_egg_ON_family,function(x){x$tTable[[8]]}) < 0.01)` of 1000 repeated analyses (Table \ref{tab:PGLS-posterior-egg-ON}).
+\ref{fig:PGLS-family-length}, p-value `r gsub(0,"<0.001",round(pgls_results_body_egg_ON_family$tTable[[8]],3))`, slope `r round(pgls_results_body_egg_ON_family$coefficients[[2]],3)`, n=`r pgls_results_body_egg_ON_family$dims$N`), supported by `r sum(sapply(allometry_body_egg_ON_family,function(x){x$tTable[[8]]}) < 0.01)` of 1000 repeated analyses (Table \ref{tab:PGLS-posterior-egg-ON}).
 
 We further explored these results using two methods: First, to evaluate our findings against alternative evolutionary hypotheses, we compared these results to regressions based on simulated data.
-Our results showed that when considering body size, the slope of the regression of egg size and ovariole number is lower than we would expect to observe by chance, as assessed by comparing to data simulated with no evolutionary correlation (Fig.
+Our results showed that when considering body size, the slope of the regression of egg size and ovariole number is \textcolor{red}{more negative} than we would expect to observe by chance, as assessed by comparing to data simulated with no evolutionary correlation (Fig.
 \ref{fig:simulated}).
-However, for both adult body length and dry mass, the slope of the regressions on observed data are not within the range that would be expected under an strong negative correlation (slope of -1 in log-log space, Fig.
+However, for both adult body length and dry mass, the slope of the regressions on observed data are not within the range that would be expected under a strong negative correlation (slope of -1 in log-log space, Fig.
 \ref{fig:simulated}).
-This suggests the presence of a weak evolutionary relationship bewteeen ovariole number and egg size, when accounting for body size.
+This suggests the presence of a weak evolutionary relationship between ovariole number and egg size, when accounting for body size.
 
@@
 We found that across Drosophilidae fly species, egg size is indeed strongly negatively correlated with ovariole number when accounting for body size (Fig.
-2c, Table \ref{tab:PGLS-posterior-clade}, p-value `r gsub(0,"<0.001",round(pgls_results_DIP$tTable[[8]],3))`, n=`r pgls_results_DIP$dims$N`).
+2c, Table \ref{tab:PGLS-posterior-clade}, p-value `r gsub(0,"<0.001",round(pgls_results_DIP$tTable[[8]],3))`, slope `r round(pgls_results_DIP$coefficients[[2]],3)`, n=`r pgls_results_DIP$dims$N`).
 For grasshoppers and crickets (Orthoptera), beetles (Coleoptera), and wasps (Hymenoptera), we observed no significant relationship between ovariole number and egg size, even when accounting for body size (Fig.
-2d, Table \ref{tab:PGLS-posterior-clade}, \ref{fig:PGLS-Coleoptera-Hymenoptera}, Orthoptera: p-value `r round(pgls_results_ORT$tTable[[8]],3)`, n=`r pgls_results_ORT$dims$N`, Coleoptera: p-value `r round(pgls_results_COL1$tTable[[8]],3)`, n=`r pgls_results_COL1$dims$N`, Hymenoptera: p-value `r round(pgls_results_HYM$tTable[[8]],3)`, n=`r round(pgls_results_HYM$dims$N)`).
+2d, \ref{fig:PGLS-Coleoptera-Hymenoptera}, Table \ref{tab:PGLS-posterior-clade}, Orthoptera: p-value `r round(pgls_results_ORT$tTable[[8]],3)`, n=`r pgls_results_ORT$dims$N`, Coleoptera: p-value `r round(pgls_results_COL1$tTable[[8]],3)`, n=`r pgls_results_COL1$dims$N`, Hymenoptera: p-value `r round(pgls_results_HYM$tTable[[8]],3)`, n=`r round(pgls_results_HYM$dims$N)`).
 This indicates that, while a strong negative correlation between egg size and ovariole number exists for some insects, it does not represent a universal pattern across insect clades.
@@
 \ref{fig:PGLS-body-plot}, Table \ref{tab:PGLS-posterior-body}, body mass: p-value `r round(pgls_results_body_gilbert_name$tTable[[8]],3)`, n=`r pgls_results_body_gilbert_genus$dims$N`, body length: p-value `r round(pgls_results_body_ON_family$tTable[[8]],3)`, n=`r pgls_results_body_ON_family$dims$N`).
-Of the four subclades considered, only insects in the order Orthoptera had a positive relationship between body size and ovariole number (Table \ref{tab:PGLS-posterior-body}, p-value `r round(pgls_results_body_ORT$tTable[[8]],3)`, n=`r pgls_results_body_ORT$dims$N`).
+Of the four subclades considered, only insects in the order Orthoptera had a positive relationship between body size and ovariole number (Table \ref{tab:PGLS-posterior-body}, p-value `r round(pgls_results_body_ORT$tTable[[8]],3)`, slope `r round(pgls_results_body_ORT$coefficients[[2]],3)`, n=`r pgls_results_body_ORT$dims$N`).
 
-![**Tests of the hypothesized trade off between egg size and ovariole number.** a, Egg volume (mm^3^) and ovariole number, both log~10~ scale; points represent insect species.
+![**Tests of the hypothesized trade-off between egg size and ovariole number.** a, Egg volume (mm^3^) and ovariole number, both log~10~ scale; points represent insect species.
 See section _Modeling ovariole number evolution_ for discussion of the enrichment of certain low values of ovariole number (points appearing vertically arranged) b, Egg volume and ovariole number, residuals to dry adult body mass, points represent genera.
@@
 
-If the hypothesized trade off between the number and size of offspring is true for insects, then one explanation for the lack of a consistent negative relationship between ovariole number and egg size is that ovariole number may not be a reasonable proxy for offspring number.
+If the hypothesized trade-off between the number and size of offspring is true for insects, then one explanation for the lack of a consistent negative relationship between ovariole number and egg size is that ovariole number may not be a reasonable proxy for offspring number.
 Previous research has shown that, across individuals within the same species, ovariole number is correlated with certain measurements of fecundity, such as maximum daily rate of egg production for _Drosophila_ [@david1970nombre;@bouletreau1982reproductive], but not others, such as lifetime fecundity[@schmidt2005geographic] or fitness in competition assays[@wayne1997quantitative].
@@
 
-Using a previously reported dataset of lifetime fecundity measurements across insects[@Gilbert2007], we assessed the relationship between lifetime fecundity and ovariole number.
+Using a previously reported dataset of lifetime fecundity measurements across insects[@Gilbert2007;@gilbert2010parental], we assessed the relationship between lifetime fecundity and ovariole number.
 We observed a significant positive relationship (Fig.
-3, p-value `r round(pgls_results_fecundity_genus$tTable[[8]],3)`, n=`r pgls_results_fecundity_genus$dims$N`), however, a substantial fraction of repeat analyses show these results are not robust to uncertainty (`r sum(sapply(allometry_fecundity_genus,function(x){x$tTable[[8]]}) > 0.01)` of 1000 regressions are not significant, Table \ref{tab:PGLS-posterior-fecundity}).
+3, p-value `r round(pgls_results_fecundity_genus$tTable[[8]],3)`, slope `r round(pgls_results_fecundity_genus$coefficients[[2]],3)`, n=`r pgls_results_fecundity_genus$dims$N`), however, a substantial fraction of repeat analyses show these results are not robust to uncertainty (`r sum(sapply(allometry_fecundity_genus,function(x){x$tTable[[8]]}) > 0.01)` of 1000 regressions are not significant, Table \ref{tab:PGLS-posterior-fecundity}).
 We note that this relationship is largely defined by the absence of insects with high ovariole number and low fecundity (Fig.
@@
 nearly all Lepidoptera have exactly eight ovarioles, Fig.
-5b, cyan), from an ancestral variable state.
+5b, \textcolor{red}{Lepidoptera are part of Amphiesmenoptera}, in cyan), from an ancestral variable state.
 These invariant lineages were identified by finding regions of the phylogeny that experience extremely low rates of ovariole number diversification (Figs.
@@
 
-Additionally, the distribution of ovariole numbers across insects is enriched for even numbers of total ovarioles (Fig.
+The distribution of ovariole numbers across insects is enriched for even numbers of total ovarioles (Fig.
 5c).
 While many insects show asymmetries in the number of ovarioles between the left and right ovaries, all of the invariant lineages are symmetric (at 4, 6, 8, 10, 12, and 14 total ovarioles).
-Therefore, invariant lineages have near-zero variation when comparing between species, between individuals within a species, and between the left and right ovary within an individual.
+\textcolor{red}{Additionally, for the insects identified as part of invariant lineages, none have any reported intraspecific variation in ovariole number.} Therefore, invariant lineages have near-zero variation when comparing between species, between individuals within a species, and between the left and right ovary within an individual.
 
@@
 Here we demonstrate that a multi-rate Brownian motion model far outperforms a single rate model in fitting the data ($\Delta$ AICc `r round(BM1$AICc - BMS$AICc,2)`).
+\textcolor{red}{In addition, using a parametric bootstrap to evaluate model fit, we find evidence that processes beyond Brownian Motion processes are likely at play (Fig.
+\ref{fig:arbutus})}[@pennell2015model]\textcolor{red}{.
+We suggest that as researchers continue to develop non-Gaussian models for continuous trait evolution}[@blomberg2020beyond]\textcolor{red}{, those models will be useful for describing the evolution of ovariole number.}
 
@@
 We suggest that considering the evolution of developmental processes that govern ovariole number specification may be more useful in explaining patterns of diversity than predictions based on metabolic trade-offs.
-The regulation of ovariole number is best understood from research on _Drosophila melanogaster_, where the number of ovarioles in the adult is determined by cell proliferation and rearrangement during larval development[@godt1995mechanisms;@king1968development].
-In this species, the number of ovarioles can vary between the left and right ovaries within an individual, as well as across individuals within a population[@telonis2005genetic;@bergland2008quantitative].
-This variation is derived primarily from variation in the number of “terminal filament precursor cells”[@green2012convergent;@sarikaya2019reproductive], as well as from variation in the number of those precursor cells that group together to form the structure that initiates ovariole formation, known as a “terminal filament”[@sarikaya2012roles].
+\textcolor{red}{As evidence of this, we point to the fact that invariant lineages appear to have near-zero variation not only across species, but also within species, and between the left and right ovary within individuals.
+This suggests that the mechanism that determines ovariole number has become canalized in these groups.
+In contrast,} our previous understanding of how ovariole number is regulated comes from research on _Drosophila melanogaster_, where the number of ovarioles can vary between the left and right ovaries within an individual, as well as across individuals within a population[@telonis2005genetic;@bergland2008quantitative].
+In this species, adult ovariole number is determined by cell proliferation and rearrangement during larval development[@godt1995mechanisms;@king1968development].
+Variation in adult number is derived primarily from variation in the number of “terminal filament precursor cells”[@green2012convergent;@sarikaya2019reproductive], as well as from variation in the number of those precursor cells that group together to form the structure that initiates ovariole formation, known as a “terminal filament”[@sarikaya2012roles].
 Across species of _Drosophila_, variation in average adult ovariole number results primarily from variation in the average number of terminal filament precursor cells[@sarikaya2019reproductive].

### 2021-03-20 manuscript_supplement.Rmd
changes to Rmd files during revision process

@@
 ---
-title: Repeated loss of variation in insect ovary morphology highlights the role of developmental constraint in life-history evolution - Supplementary Methods
+title: Repeated loss of variation in insect ovary morphology highlights the role of development in life-history evolution - Supplementary Methods
 author: 
@@
 
-Using this approach, we gathered `r nrow(raw_ON)` records for ovariole number from 460 publications.
+For records of ovariole number that reported intraspecific variation in ovariole number, we calculated the percent difference as follows: if ovariole number was reported as a range, percent difference was calculated as the $100 * ((max - min) / median)$; if ovariole number was reported as an average with deviations, percent difference was calculated as $100 * ((2 * deviation) / mean)$.
+When independent observations of ovariole number for a given species were available across multiple published records, we calculated the percent difference as the $100 * ((max - min) / median)$ (Fig.
+\ref{fig:intraspecific-variation}).
+
+Using this approach, we gathered `r nrow(raw_ON)` records for ovariole number from `r raw_ON %>% distinct(publication) %>% nrow()` publications.
 A full list of publications is provided in the supplementary file 'ovariole_number_bibliography.pdf'.
@@
 
+```{r intraspecific-variation, fig.cap="**Intraspecific variation in ovariole number.**", fig.width=7.5, fig.height=3, include=TRUE}
+# calculate the percent difference reported for each especies
+percdiff_within_record <- ON %>% filter(!is.na(deviation_on)) %>% 
+    filter(!is.na(species)) %>% #check that species name is present
+    mutate(percdiff = ifelse(!is.na(deviation_on),100*((2*deviation_on)/mean_on), #if ovariole number is recorded as an average and deviation
+        ifelse(!is.na(median_on),100*((maximum_on - minimum_on)/median_on),NA))) %>% #if it is recorded as a range
+    filter(!is.na(percdiff) & percdiff>0)
+
+# get the names of species that have more than one description
+ON_name_count <- ON %>% 
+    filter(species != "") %>% #check that species name is present
+    count(name) %>% 
+    filter(n > 1)
+
+percdiff_across_record <- ON %>% 
+    filter(name %in% ON_name_count$name) %>% #get records for multiple species
+    group_by(name) %>%  
+    summarize(max_on = max(on), min_on = min(on),genus=genus) %>%
+    ungroup() %>%
+    mutate(percdiff = 100*((max_on - min_on)/(min_on + ((max_on - min_on)/2)))) %>% #calculated percent difference
+    filter(!is.na(percdiff) & percdiff>0)
+
+pd1 <- ggplot(percdiff_within_record,aes(x=percdiff)) + geom_histogram(bins=25) + scale_x_log10() + ggtitle("reported variation within observation") + xlab("percent difference")
+pd2 <- ggplot(percdiff_across_record,aes(x=percdiff)) + geom_histogram(bins=25) + scale_x_log10() + ggtitle("variation across repeated observations") + xlab("percent difference")
+grid.arrange(pd1,pd2,ncol=2)
+```
+
 # Phylogenetic trees {#sec:trees}
@@
 
-Analyses of insect family-level ovariole number, egg size, and body size were performed using the insect phylogeny published in Rainford et. al, 2014[@rainford2014phylogenetic].
+For regressions involving body size data that were reported as insect family-level averages, we used the insect phylogeny published in Rainford et. al, 2014[@rainford2014phylogenetic].
 
@@
 
-We combined the data we collected on total ovariole number with existing datasets of egg size and shape[@church2019dataset], insect lifetime fecundity and dry adult body mass[@Gilbert2007], average adult body length per insect family[@rainford2016phylogenetic], and several lineage-specific measures of adult body size[@iwata1966large;@Iwata1966a;@waloff1954number;@starmer2003phylogenetic;@reinhardt2005low].
-
-Ovariole number and egg size[@church2019insect] data were combined by matching records at the species level (Fig 2a).
+We combined the data we collected on total ovariole number with existing datasets of egg size and shape[@church2019dataset], insect lifetime fecundity and dry adult body mass[@Gilbert2007;@gilbert2010parental;@gilbert2011insect], average adult body length per insect family[@rainford2016phylogenetic], and several lineage-specific measures of adult body size[@iwata1966large;@Iwata1966a;@waloff1954number;@starmer2003phylogenetic;@reinhardt2005low].
+
+Ovariole number and egg size[@church2019insect] data were combined by matching records across datasets for the same insect species (Fig 2a).
 When multiple records existed for a given species, the dataset was randomly shuffled and a single matching record was selected.
 This variation across records for the same species was accounted for in regressions by reshuffling and matching records at each iteration of the analysis.
-We also matched records at the genus level following the same reshuffling method, which allowed us to test whether results were robust with a larger sample size when an exact species match was not available (Fig.
+We also matched records for insects in the same genus following the same reshuffling method, which allowed us to test whether results were robust with a larger sample size when an exact species match was not available (Fig.
 \ref{fig:PGLS-egg-ON-genus}).
@@
 
-Ovariole number, egg volume, lifetime fecundity, and adult body mass[@Gilbert2007] were combined by matching records at the species level and genus level, using the same method as described above (Figs.
+Ovariole number, egg volume, lifetime fecundity, and adult body mass[@Gilbert2007;@gilbert2010parental;@gilbert2011insect] were combined by matching records at the species level and genus level, using the same method as described above (Figs.
 2b, 3, and \ref{fig:PGLS-Gilbert-name}).
@@
 
+```{r PGLS-egg-ON-genus-noneusocial, fig.cap="**Egg volume vs ovariole number, matching records at the genus level, excluding records from families with eusocial insects.**", fig.width=7.5, fig.height=5, include=TRUE}
+### PGLS on vs egg volume by insect genus, excluding eusocial
+eusocial_family <- ON %>% filter(grepl("queen",condition)) %>% pull(family) %>% unique()
+
+egg_ON_genus_noneusocial_pgls <- egg_ON_genus %>% 
+    filter(!(family %in% eusocial_family)) %>% 
+    select(genus,logvol,logon,group) %>% 
+    rename(rank = genus, trait1 = logon, trait2 = logvol)
+
+pgls_results_egg_ON_genus_noneusocial <- run_all_taxa_pgls(egg_ON_genus_noneusocial_pgls,tree)
+
+run_plotting_pgls(egg_ON_genus_noneusocial_pgls,tree,"egg_ON_genus_noneusocial") +
+    xlab(bquote('ovariole number,'~log[10])) +
+    ylab(bquote('egg volume,'~mm^3~','~log[10]))
+```
+
 ```{r PGLS-egg-ON-weight-genus}
@@
 
-For regressions at the species and genus level, we reshuffled and matched records at each iteration to account for variation across records for the same taxon.
-For regressions at the family level we recalculated the average ovariole number per insect family, downsampling the representation for each family by half.
+For regressions using data matched across species or genera, we reshuffled and matched records at each iteration to account for variation across records for the same taxon.
+For regressions on family-level average data, we recalculated the average ovariole number per insect family, downsampling the representation for each family by half.
 No posterior distribution was available with the previously published family level phylogeny[@rainford2014phylogenetic].
 
-To account for body size, we calculated the phylogenetic residuals[@revell2010phylogenetic] of each trait to body size, and then compared the evolution of these residuals using a PGLS regression.
+To account for body size, we calculated the phylogenetic residuals[@revell2009size] of each trait to body size, and then compared the evolution of these residuals using a PGLS regression.
 
@@
 Right, comparing egg size and ovariole number, accounting for body mass at the genus level.
+The slope of regressions on observed egg size and ovariole number (red) is more negative than we would expect to observe by chance, as assessed by comparing to data simulated with no evolutionary correlation (gray), but also not within the range that would be expected under an strong negative correlation (black).
 Red=observed values, gray=simulated with no correlation, black=simulated with a -1:1 correlation.
@@
 
-```{r PGLS-body-plot, fig.cap="**Adult body size vs ovariole number.** Adult body mass compared at the genus level (left), and average adult body length compared at the family level (right).", fig.width=6, fig.height=3, include=TRUE}
+```{r PGLS-body-plot, fig.cap="**Adult body size vs ovariole number.** Adult body mass, matching records at the genus level (left), and family-level average adult body length (right).", fig.width=6, fig.height=3, include=TRUE}
 ### body length vs ON, family level
@@
 
-egg_ON_taxonomic_levels <- c("species","genus","species","genus","family")
+egg_ON_taxonomic_levels <- c("species","genus","species","genus","family average")
 
 egg_ON_table <- as.data.frame(bind_rows(lapply(egg_ON_posterior_results,get_pgls_table))) %>% mutate(analysis = egg_ON_pgls_analysis_names,
-        "taxonomic level" = egg_ON_taxonomic_levels,
+        "data match" = egg_ON_taxonomic_levels,
         "slope" = paste(sprintf("%.3f",slope_min),sprintf("%.3f",slope_max),sep=sp),
@@
         "taxa" = taxa)
-knitr::kable(egg_ON_table %>% select(c("analysis","taxonomic level","slope","MCC p-value","num. sig. / 1000","taxa")),
+knitr::kable(egg_ON_table %>% select(c("analysis","data match","slope","MCC p-value","num. sig. / 1000","taxa")),
         "latex",
         align=rep('l', 6),
-        caption="Results of PGLS analysis of ovariole number and egg size across a posterior distribution", booktabs = T, linesep = "") %>% 
+        caption="Results of PGLS analysis of ovariole number and egg size across a posterior distribution.
+The 'data match' columns describes how observations were matched across datasets, e.g.
+matching records for the same species, genera, or using family-level averages.", booktabs = T, linesep = "") %>%
     kableExtra::kable_styling(latex_options = "striped") %>%
@@
 clade_table <- as.data.frame(bind_rows(lapply(clade_posterior_results,get_pgls_table))) %>% mutate(analysis = clade_pgls_analysis_names,
-        "taxonomic level" = clade_taxonomic_levels,
+        "data match" = clade_taxonomic_levels,
         "slope" = paste(sprintf("%.3f",slope_min),sprintf("%.3f",slope_max),sep=sp),
@@
         "taxa" = taxa)
-knitr::kable(clade_table %>% select(c("analysis","taxonomic level","slope","MCC p-value","num. sig. / 1000","taxa")),
+knitr::kable(clade_table %>% select(c("analysis","data match","slope","MCC p-value","num. sig. / 1000","taxa")),
         "latex",
         align=rep('l', 6),
-        caption="Results of PGLS analysis of ovariole number and egg size across a posterior distribution", booktabs = T, linesep = "") %>% 
+        caption="Results of PGLS analysis of ovariole number and egg size across a posterior distribution.", booktabs = T, linesep = "") %>% 
     kableExtra::kable_styling(latex_options = "striped") %>%
@@
 
-body_taxonomic_levels <- c("species","genus","family","species","genus","genus","genus")
+body_taxonomic_levels <- c("species","genus","family average","species","genus","genus","genus")
 
 body_table <- as.data.frame(bind_rows(lapply(body_posterior_results,get_pgls_table))) %>% mutate(analysis = body_pgls_analysis_names,
-        "taxonomic level" = body_taxonomic_levels,
+        "data match" = body_taxonomic_levels,
         "slope" = paste(sprintf("%.3f",slope_min),sprintf("%.3f",slope_max),sep=sp),
@@
         "taxa" = taxa)
-knitr::kable(body_table %>% select(c("analysis","taxonomic level","slope","MCC p-value","num. sig. / 1000","taxa")),
+knitr::kable(body_table %>% select(c("analysis","data match","slope","MCC p-value","num. sig. / 1000","taxa")),
         "latex",
         align=rep('l', 6),
-        caption="Results of PGLS analysis of ovariole number and body size across a posterior distribution", booktabs = T, linesep = "") %>% 
+        caption="Results of PGLS analysis of ovariole number and body size across a posterior distribution.", booktabs = T, linesep = "") %>% 
     kableExtra::kable_styling(latex_options = "striped") %>%
@@
 fecundity_table <- as.data.frame(bind_rows(lapply(fecundity_posterior_results,get_pgls_table))) %>% mutate(analysis = fecundity_pgls_analysis_names,
-        "taxonomic level" = fecundity_taxonomic_levels,
+        "data match" = fecundity_taxonomic_levels,
         "slope" = paste(sprintf("%.3f",slope_min),sprintf("%.3f",slope_max),sep=sp),
@@
         "taxa" = taxa)
-knitr::kable(fecundity_table %>% select(c("analysis","taxonomic level","slope","MCC p-value","num. sig. / 1000","taxa")),
+knitr::kable(fecundity_table %>% select(c("analysis","data match","slope","MCC p-value","num. sig. / 1000","taxa")),
         "latex",
-        caption="Results of PGLS analysis of ovariole number and fecundity across a posterior distribution", booktabs = T, linesep = "") %>% 
+        caption="Results of PGLS analysis of ovariole number and fecundity across a posterior distribution.", booktabs = T, linesep = "") %>% 
     kableExtra::kable_styling(latex_options = "striped") %>%

### 2021-03-20 set_up_conda_environment.sh

@@ -20,6 +20,7 @@ install.packages("ggplot2")
 install.packages("BAMMtools")
 install.packages("coda")
 install.packages("OUwie")
+# on cluster, outside of R
 #conda install -c conda-forge r-rmpfr
 #conda install -c anaconda gmp
 install.packages("corHMM")

### 2021-04-02 manuscript.Rmd
proof read revisions April 2021

@@
 However in contrast to prior claims, we note that this relationship is not generalizable across all insect clades, and we highlight several factors that may have contributed to this size-number trade-off being stated as a general rule in previous studies.
-We reconstruct the evolutionary history of the nurse cell arrangement within the ovariole, and show that the diversification of ovariole number and egg size have both been largely independent of nurse cell presence or position within the ovariole.
+We reconstruct the evolution of the arrangement of cells that contribute nutrients and patterning information during oogenesis (nurse cells), and show that the diversification of ovariole number and egg size have both been largely independent of their presence or position within the ovariole.
 Instead we show that ovariole number evolution has been shaped by a series of transitions between variable and invariant states, with multiple independent lineages evolving to have almost no variation in ovariole number.
@@
 Therefore total ovariole number may be an even or odd integer for an individual female insect.
-In addition to variation in the number of ovarioles, the tissue morphology within ovarioles varies across insects, and has been classified into several modes of oogenesis based on the presence and position of special nutritive cells called nurse cells[@buning1994insect].
+In addition to variation in the number of ovarioles, the tissue morphology within ovarioles varies across insects, and has been classified into several modes of oogenesis based on the presence and position of special oocyte-associated  cells called nurse cells[@buning1994insect].
 
 Here we compiled `r nrow(raw_ON)` records of ovariole number from across `r ON %>% distinct(order) %>% nrow()` orders, `r ON %>% distinct(family) %>% nrow()` families, and `r ON %>% filter(species != "") %>% distinct(genus,species) %>% nrow()` species of insects.
-We combined these data with published datasets of egg size[@church2019dataset], fecundity[@Gilbert2007;@gilbert2010parental], and body size[@rainford2016phylogenetic], to test hypotheses about \textcolor{red}{the evolutionary trade-off between offspring size and number}.
-In these analyses we used an existing phylogeny of insects[@church2019insect] to analyze evolutionary patterns in ovariole number, and found that hypotheses about \textcolor{red}{life-history} evolution do not hold generally true across insects.
-\textcolor{red}{We then combined these data with published observations of the mode of oogenesis}[@buning1994insect]\textcolor{red}{ and reconstructed the evolutionary history of the presence and position of nurse cells that contribute to the oocyte during oogenesis.
-We tested whether patterns in the distribution of ovariole number, egg size, or egg shape were driven by the evolution of nurse cells, and found no significant results.} Instead we observe that the phylogenetic distribution of ovariole number suggests a model where the developmental mechanisms that govern ovariole number have shifted between variable and invariant states several times over the course of insect evolution.
+We combined these data with published datasets of egg size[@church2019dataset], fecundity[@Gilbert2007;@gilbert2010parental], and body size[@rainford2016phylogenetic], to test hypotheses about the evolutionary trade-off between offspring size and number.
+In these analyses we used an existing phylogeny of insects[@church2019insect] to analyze evolutionary patterns in ovariole number, and found that hypotheses about life-history evolution do not hold generally true across insects.
+We then combined these data with published observations of the mode of oogenesis[@buning1994insect] and reconstructed the evolutionary history of the presence and position of nurse cells that contribute to the oocyte during oogenesis.
+We tested whether patterns in the distribution of ovariole number, egg size, or egg shape were driven by the evolution of nurse cells, and found no significant results.
+Instead we observe that the phylogenetic distribution of ovariole number suggests a model where the developmental mechanisms that govern ovariole number have shifted between variable and invariant states several times over the course of insect evolution.
 Based on this finding, we propose that the developmental mechanisms used to establish ovariole number in well-studied insects such as _Drosophila melanogaster_ are unlikely to regulate ovariole number in all insects.
@@
 We searched the published literature for references to insect ovariole number using a predetermined set of 131 search terms, entered into Google Scholar (scholar.google.com) between June and October of 2019.
-Each search term was \textcolor{red}{comprised of} an insect taxonomic group and the words "ovariole number".
-\textcolor{red}{The taxonomic groups used to search included} all insect orders, many large insect families, and taxonomic groups that are well-represented in the insect egg dataset[@church2019dataset].
-\textcolor{red}{For each Google Scholar search}, we evaluated all publications in the first page of results (ten publications).
+Each search term was comprised of an insect taxonomic group and the words "ovariole number".
+The taxonomic groups used in the search process included all insect orders, many large insect families, and taxonomic groups that are well-represented in the insect egg dataset[@church2019dataset].
+For each Google Scholar search, we evaluated the first ten pages in the search results.
 For 61 search terms that had a large number of informative hits, significant representation in the egg dataset, or that corresponded to very speciose groups, we evaluated an additional 20 publications.
@@
 
-Using this approach, we gathered `r nrow(raw_ON)` records for ovariole number from `r raw_ON %>% distinct(publication) %>% nrow()` publications, listed in the supplementary file 'ovariole_number_bibliography.pdf'.
+Using this approach, we gathered `r nrow(raw_ON)` records for ovariole number for  `r ON %>% distinct(order) %>% nrow()` insect orders, `r ON %>% distinct(family) %>% nrow()` families, and `r ON %>% filter(species != "") %>% distinct(genus,species) %>% nrow()` species, using `r raw_ON %>% distinct(publication) %>% nrow()` publications that are listed in the supplementary file 'ovariole_number_bibliography.pdf'.
 We matched these records to additional taxonomic information using the software TaxReformer[@TaxReformer].
-For all subsequent analyses, we excluded observations made in non-reproductive individuals from eusocial species (workers), as well as two observations that represented significant outliers and could not be validated using additional sources[@su2015testicular;@hernandez2017biology].
+For all subsequent analyses, we excluded observations made in non-reproductive individuals from eusocial species (e.g.
+workers), as well as two observations that represented significant outliers and could not be validated using additional sources[@su2015testicular;@hernandez2017biology].
 See supplementary methods section \ref{sec:gathering} for details.
 
-\textcolor{red}{For records of ovariole number that reported intraspecific variation in ovariole number, we calculated the percent difference as follows: if ovariole number was reported as a range, percent difference was calculated as the $100 * ((max - min) / median)$; if ovariole number was reported as an average with deviations, percent difference was calculated as $100 * ((2 * deviation) / mean)$.
-When independent observations of ovariole number for a given species were available across multiple published records, we calculated the percent difference as the $100 * ((max - min) / median)$.}
+For records of ovariole number that reported intraspecific variation in ovariole number, we calculated the percent difference as follows: if ovariole number was reported as a range, percent difference was calculated as the $100 * ((max - min) / median)$; if ovariole number was reported as an average with deviations, percent difference was calculated as $100 * ((2 * deviation) / mean)$.
+When independent observations of ovariole number for a given species were available across multiple published records, we calculated the percent difference as the $100 * ((max - min) / median)$.
 
@@
 The analyses in this manuscript were performed using the insect phylogeny published in Church et al., 2019[@church2019insect], unless otherwise specified.
-\textcolor{red}{For regressions involving body size data that were reported as insect family-level averages,} we used the insect phylogeny published in Rainford et.
+For regressions involving body size data that were reported as insect family-level averages, we used the insect phylogeny published in Rainford et.
 al, 2014[@rainford2014phylogenetic].
@@
 Running BAMM for the maximum possible number of generations and selecting the optimum burn-in (Fig.
-\ref{fig:BAMM-analysis-burnin}) resulted in an effective size for the number of shifts of `r round(es_nshifts,2)`, and for log-likelihood of `r round(es_logl,2)`.
+\ref{fig:BAMM-analysis-burnin}) resulted in an effective size for the number of shifts of `r round(es_nshifts,2)`, and for the log-likelihood of `r round(es_logl,2)`.
 Repeated BAMM analyses showed similar distributions of high and low rate regimes, indicating the implications for ovariole number evolution are robust to uncertainty in rate estimates.
@@
 
-All phylogenetic regressions were performed using the maximum clade credibility (MCC) tree \textcolor{red}{(the tree with highest credibility score from the posterior distribution of the Bayesian analysis)}.
+All phylogenetic regressions were performed using the maximum clade credibility (MCC) tree (the tree with highest credibility score from the posterior distribution of the Bayesian analysis).
 We considered a relationship significant when the p-value was below the threshold 0.01.
-To assess the robustness of results to uncertainty in phylogenetic relationships, we also \textcolor{red}{repeated these analyses over the posterior distribution of phylogenetic trees} and report the number of regressions that gave a significant result (see Table \ref{tab:PGLS-posterior-egg-ON}).
+To assess the robustness of results to uncertainty in phylogenetic relationships, we also repeated these analyses over the posterior distribution of phylogenetic trees and report the number of regressions that gave a significant result (see Table \ref{tab:PGLS-posterior-egg-ON}).
 
 For two comparisons, we validated that our tests had sufficient statistical power using the selected threshold by comparing the distribution of p-values from regressions of observed data to regressions of data simulated under alternative hypotheses.
-We compared the results of analyses of our observed to those based on simulated data to evaluate the likelihood of false positives (comparing to data \textcolor{red}{simulated} under no correlation) and false negatives (comparing to data simulated with strong correlation).
+We compared the results of analyses of our observed to those based on simulated data to evaluate the likelihood of false positives (comparing to data simulated under no correlation) and false negatives (comparing to data simulated with strong correlation).
 
@@
 
-Ovariole number varies by at least four orders of magnitude across \textcolor{red}{insect species} (Fig.
+Ovariole number varies by at least four orders of magnitude across insect species (Fig.
 1b).
@@
 In these insects one of the two ovaries presumably established during embryogenesis is reported to atrophy during development[@pollock1996viviparous;@pluot1979evolution], resulting in an asymmetric adult reproductive system.
-\textcolor{red}{We also evaluated intraspecific variation in ovariole number, and found that, for species for which it has been reported, the average percent difference number within species is between 10\% and 100\% of the median value (Fig.
-\ref{fig:intraspecific-variation}).}
+We also evaluated intraspecific variation in ovariole number, and found that, for species for which it has been reported, the average percent difference number within species is between 10\% and 100\% of the median value (Fig.
+\ref{fig:intraspecific-variation}).
 
@@
 2a, Table \ref{tab:PGLS-posterior-egg-ON}, p-value `r round(pgls_results_egg_ON_name$tTable[[8]],3)`, n=`r pgls_results_egg_ON_name$dims$N`).
-We also compared egg size and ovariole number, \textcolor{red}{combining data from species within the same genus to increase sample size,} and again did not observe a significant relationship (Fig.
+We also compared egg size and ovariole number, combining data from species within the same genus to increase sample size, and again did not observe a significant relationship (Fig.
 \ref{fig:PGLS-egg-ON-genus}, p-value `r round(pgls_results_egg_ON_genus$tTable[[8]],3)`, n=`r pgls_results_egg_ON_genus$dims$N`).
-\textcolor{red}{To verify this finding was not driven by the high ovariole numbers seen in the queens of some eusocial insects, we repeated this comparison excluding insects from families with eusocial representatives, with the same result (Fig.
-\ref{fig:PGLS-egg-ON-genus-noneusocial}, p-value `r round(pgls_results_egg_ON_genus_noneusocial$tTable[[8]],3)`, n=`r pgls_results_egg_ON_genus_noneusocial$dims$N`).}
+To verify this finding was not driven by the high ovariole numbers seen in the queens of some eusocial insects, we repeated this comparison excluding insects from families with eusocial representatives, with the same result (Fig.
+\ref{fig:PGLS-egg-ON-genus-noneusocial}, p-value `r round(pgls_results_egg_ON_genus_noneusocial$tTable[[8]],3)`, n=`r pgls_results_egg_ON_genus_noneusocial$dims$N`).
 
-Given that this predicted relationship is often conditioned on body size, \textcolor{red}{which is predicted to limit total potential reproductive investment}[@stewart1991reproductive;@starmer2003phylogenetic], we combined data on ovariole number and egg size with data on insect adult body mass[@Gilbert2007;@gilbert2010parental;@gilbert2011insect] and length[@rainford2016phylogenetic].
-When accounting for adult body mass, we observed a significant negative relationship between egg size and ovariole number (Fig.
+Given that this predicted relationship is often conditioned on body size, which is predicted to limit total potential reproductive investment[@stewart1991reproductive;@starmer2003phylogenetic], we combined data on ovariole number and egg size with data on insect adult body mass[@Gilbert2007;@gilbert2010parental;@gilbert2011insect] and length[@rainford2016phylogenetic].
+When accounting for adult body mass, we observed a significant negative relationship between egg size and ovariole number across genera (Fig.
 2b, \ref{fig:PGLS-Gilbert-name}, p-value `r round(pgls_results_gilbert_genus$tTable[[8]],3)`, slope `r round(pgls_results_gilbert_genus$coefficients[[2]],3)`, n=`r pgls_results_gilbert_genus$dims$N`).
@@
 We further explored these results using two methods: First, to evaluate our findings against alternative evolutionary hypotheses, we compared these results to regressions based on simulated data.
-Our results showed that when considering body size, the slope of the regression of egg size and ovariole number is \textcolor{red}{more negative} than we would expect to observe by chance, as assessed by comparing to data simulated with no evolutionary correlation (Fig.
+Our results showed that when considering body size, the slope of the regression of egg size and ovariole number is more negative than we would expect to observe by chance, as assessed by comparing to data simulated with no evolutionary correlation (Fig.
 \ref{fig:simulated}).
@@
 
-In addition to the number of ovarioles, insect ovary morphology has been classified into several modes of oogenesis based on the presence and position of nutritive cells called nurse cells[@buning1994insect], (Fig.
+In addition to the number of ovarioles, insect ovary morphology has been classified into several modes of oogenesis based on the presence and position of cells that provide nutritive and patterning molecules to the oocyte, which are called nurse cells[@buning1994insect] (Fig.
 4a).
@@
 melanogaster_ is an example of a meroistic oogenesis mode, meaning that its ovarioles contain nurse cells of germ line origin that are connected to developing oocytes via cytoplasmic bridges[@KingRobertC1970OdiD].
-In insects with a polytrophic arrangement, these nurse cells are clonally related and immediately adjacent to each oocyte.
+In insects with a polytrophic meroistic arrangement, these nurse cells are clonally related and immediately adjacent to each oocyte.
 An alternative arrangement is seen in telotrophic meroistic ovaries, where oocytes in each ovariole are connected to a common pool of nurse cells located in the germarium[@buning1994insect].
@@
 nearly all Lepidoptera have exactly eight ovarioles, Fig.
-5b, \textcolor{red}{Lepidoptera are part of Amphiesmenoptera}, in cyan), from an ancestral variable state.
+5b, Lepidoptera are part of Amphiesmenoptera, in cyan), from an ancestral variable state.
 These invariant lineages were identified by finding regions of the phylogeny that experience extremely low rates of ovariole number diversification (Figs.
@@
 While many insects show asymmetries in the number of ovarioles between the left and right ovaries, all of the invariant lineages are symmetric (at 4, 6, 8, 10, 12, and 14 total ovarioles).
-\textcolor{red}{Additionally, for the insects identified as part of invariant lineages, none have any reported intraspecific variation in ovariole number.} Therefore, invariant lineages have near-zero variation when comparing between species, between individuals within a species, and between the left and right ovary within an individual.
+Additionally, for the insects identified as part of invariant lineages, none have any reported intraspecific variation in ovariole number.
+Therefore, invariant lineages have near-zero variation when comparing between species, between individuals within a species, and between the left and right ovary within an individual.
 
@@
 Here we demonstrate that a multi-rate Brownian motion model far outperforms a single rate model in fitting the data ($\Delta$ AICc `r round(BM1$AICc - BMS$AICc,2)`).
-\textcolor{red}{In addition, using a parametric bootstrap to evaluate model fit, we find evidence that processes beyond Brownian Motion processes are likely at play (Fig.
-\ref{fig:arbutus})}[@pennell2015model]\textcolor{red}{.
-We suggest that as researchers continue to develop non-Gaussian models for continuous trait evolution}[@blomberg2020beyond]\textcolor{red}{, those models will be useful for describing the evolution of ovariole number.}
+In addition, using a parametric bootstrap to evaluate model fit, we find evidence that processes beyond Brownian Motion processes are likely at play (Fig.
+\ref{fig:arbutus})[@pennell2015model].
+We suggest that as researchers continue to develop non-Gaussian models for continuous trait evolution[@blomberg2020beyond], those models will be useful for describing the evolution of ovariole number.
 
@@
 In insects, egg size and ovariole number are often used as proxies for offspring fitness[@koch2014mother] and number[@david1970nombre;@bouletreau1982reproductive], respectively, and therefore it has been predicted that insects with more ovarioles lay smaller eggs than insects with fewer ovarioles[@hodin2009she;@montague1981reproductive;@berrigan1991allometry;@starmer2003phylogenetic].
-Our results, using a dataset that spans `r nrow(raw_ON)` observations and that takes into account phylogenetic relationships, indicate that a generalized trade-off between insect egg size and ovariole number does not exist (Fig.
+Our results, using a dataset that spans `r nrow(raw_ON)` observations across `r ON %>% filter(species != "") %>% distinct(genus,species) %>% nrow()` species, and that takes into account phylogenetic relationships, indicate that a generalized trade-off between insect egg size and ovariole number does not exist (Fig.
 2).
@@
 We suggest that considering the evolution of developmental processes that govern ovariole number specification may be more useful in explaining patterns of diversity than predictions based on metabolic trade-offs.
-\textcolor{red}{As evidence of this, we point to the fact that invariant lineages appear to have near-zero variation not only across species, but also within species, and between the left and right ovary within individuals.
+As evidence of this, we point to the fact that invariant lineages appear to have near-zero variation not only across species, but also within species, and between the left and right ovary within individuals.
 This suggests that the mechanism that determines ovariole number has become canalized in these groups.
-In contrast,} our previous understanding of how ovariole number is regulated comes from research on _Drosophila melanogaster_, where the number of ovarioles can vary between the left and right ovaries within an individual, as well as across individuals within a population[@telonis2005genetic;@bergland2008quantitative].
+In contrast, our previous understanding of how ovariole number is regulated comes from research on _Drosophila melanogaster_, where the number of ovarioles can vary between the left and right ovaries within an individual, as well as across individuals within a population[@telonis2005genetic;@bergland2008quantitative].
 In this species, adult ovariole number is determined by cell proliferation and rearrangement during larval development[@godt1995mechanisms;@king1968development].
